---
title: "Figure 2 UMAP Making"
author: "Jonathan Chin Cheong"
date: "2024-10-21"
output: html_document
---

```{r}
set.seed(123)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1-2")
```

```{r pressure, echo=FALSE}
#Load Libraries
library(Seurat)
library(SeuratData)
library(ggplot2)
library(patchwork)
library(dplyr)
library(Rfast2)
library(viridis)
library(scCustomize)
library(harmony)
library(khroma)
```

Load Data Sets

```{r}
#Newer Data
#alldata = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Preprocessing/MultimodalAtlas_Integrated_Annotated_012725.rds")

alldata = readRDS("")
```

```{r}
# 1. Define base colors for the metaclusters
metacluster_colors <- c(
  "IFE" = "royalblue2",              # Light Red for IFE
  "HF" = "turquoise2",               # Red for HF
  "Sebocytes" = "deepskyblue2",        # Orange for Sebocytes
  "Fibroblasts" = "red2",      # Green for Fibroblasts
  "Immune cells" = "springgreen2",     # Light Blue for Immune cells
  "Endothelial cells" = "gold2",# Lighter Blue for Endothelial cells
  "Pericytes" = "mediumpurple2",        # Yellow for Pericytes
  "Muscle cells" = "orange2",     # Purple for Muscle cells
  "Schwann cells" = "darkseagreen2",    # Olive for Schwann cells
  "Melanocytes" = "orchid2",      # Pink for Melanocytes
  "Adipocytes" = "magenta2",       # Light Green for Adipocytes
  "RBCs" = "tomato3"              # Dark Red for RBCs
)

# 2. Create finer colors within each metacluster group
# You can use functions like colorRampPalette to generate gradients or variations for each fine cell type.
fine_colors <- list()

for (meta in names(metacluster_groups)) {
  num_celltypes <- length(metacluster_groups[[meta]])
  pal <- colorway::build_palette(central_color = metacluster_colors[[meta]], n_colors = num_celltypes)
  fine_colors[[meta]] <- pal$palette
}

# 3. Unlist the fine colors into a single vector, matching the new_order
color_palette <- unname(unlist(fine_colors))

# 4. Assign the custom color palette to your Seurat object
# Assuming `new_order` holds the correctly ordered cell type identities
#Idents(alldata) <- factor(Idents(alldata), levels = new_order)

# Map the palette to the factor levels in the Seurat object
names(color_palette) <- new_order

# 5. Now, use this palette in your plots
DimPlot(alldata, cols = color_palette, reduction = "umap.rpca") 
ggsave("UMAPwithLedgend.png", height = 11, width = 11)
DimPlot(alldata, cols = color_palette, reduction = "umap.rpca") & NoLegend()
DimPlot(alldata, cols = color_palette, reduction = "umap.rpca", label = T, repel = T, label.size = 3) & NoLegend()

alldata <- subset(alldata, subset = orig.ident %in% c("UW_Han", "UW_Joost_1", "UW_Joost_2", "UW_Joost_3"), invert = T)
```

MakeUMAPS for Figure 1

```{r}
DimPlot(alldata, cols = color_palette, reduction = "umap.rpca", raster =F) & NoLegend() + NoAxes()
ggsave("UMAPofFineCellTypeAnnotationsNoLabels.svg", width = 6, height = 6)
ggsave("UMAPofFineCellTypeAnnotationsNoLabels.png", width = 6, height = 6)
```

```{r}
VlnPlot(alldata, "percent_mito", group.by = modality)
```

```{r}
library(dplyr)
#Add new modality group 
alldata$modality_group <- ifelse(alldata$modality %in% c("scRNAseq", "CITEseq"), "whole cell", "nuclei")

# Add a new metadata column called 'condition' based on 'timepoint'
alldata@meta.data <- alldata@meta.data %>%
  mutate(condition = case_when(
    timepoint == "Unwounded" ~ "Unwounded",
    timepoint %in% c("Wounded_D1PW", "Wounded_D2PW", "Wounded_D4PW", "Wounded_D7PW", "Wounded_D15PW", "Wounded_D30PW") ~ "Wounded",
    TRUE ~ as.character(timepoint)  # Keeps any other potential values unchanged
  ))
```

```{r}
cols = c("springgreen2", "orange2", "steelblue2")

p <- VlnPlot(alldata, 
             features = c("nFeature_RNA", "percent_mito", "percent_ribo", "TF Percentage"), 
             cols = cols, 
             group.by = "modality", 
             pt.size = 0)

# Customize the font
p <- p & theme(
  text = element_text(family = "Arial", face = "bold", size = 11),
  axis.text = element_text(family = "Arial", face = "bold", size = 11),
  axis.title = element_text(family = "Arial", face = "bold", size = 11),
  legend.text = element_text(family = "Arial", face = "bold", size = 11),
  legend.title = element_text(family = "Arial", face = "bold", size = 11)
)

# Display the plot
p

ggsave("QualityControlofModalities.png", height = 6, width = 7)


DimPlot(alldata, group.by = "orig.ident", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByBatch.png", height = 6, width = 7)

DimPlot(alldata, group.by = "timepoint", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByTimepoint.png", height = 6, width = 7)

DimPlot(alldata, group.by = "modality", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByModality.png", height = 6, width = 7)

DimPlot(alldata, group.by = "chemistry", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByChemistry.png", height = 6, width = 7)

DimPlot(alldata, group.by = "sex", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupBySex.png", height = 6, width = 7)


DimPlot(alldata, group.by = "orig.ident", reduction = "umap.rpca", raster = F,shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByBatchNoLegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "timepoint", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByTimepointNolegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "modality", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByModalityNoLegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "chemistry", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByChemistryNoLegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "sex", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() +  NoLegend()
ggsave("RPCAUMAP_GroupBySexNoLegend.png", height = 6, width = 6)



```

Making Graphs for CSCI WIP Progress

```{r}
name = "AllCells"
cols = c("royalblue2", "turquoise2", "deepskyblue2", "red2", "springgreen2", "gold2", "mediumpurple2", "orange2", "darkseagreen2", "orchid2", "magenta2", "tomato3")


alldata$celltype <- Idents(alldata)


cells <- as.data.frame(table(alldata$metaclusters,alldata$timepoint))

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4


#ggsave(filename = paste0(name, "Subset_RatioOvertimeVer2WithNewRuns.tiff"))

alldata.1 = SplitObject(alldata, split.by = "modality")
cells <- as.data.frame(table(alldata.1$scRNAseq$metaclusters,alldata.1$scRNAseq$timepoint))

cells <- cells %>% filter(Freq != 0)

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4

ggsave("scRANSeqMetaclusterBarPlotOverTimepointNewTimepoints.png")


cells <- as.data.frame(table(alldata.1$snRNAseq$metaclusters,alldata.1$snRNAseq$timepoint))
cells <- cells %>% filter(Freq != 0)

g5 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g5 <- g5 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g5

ggsave("snRNAseqMetaclusterBarPlotOverTimepointNewTimePoints.png")


cells <- as.data.frame(table(alldata.1$CITEseq$metaclusters,alldata.1$CITEseq$timepoint))
cells <- cells %>% filter(Freq != 0)

g6 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g6 <- g6 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g6

ggsave("CITEseqMetaclusterBarPlotOverTimepointNewTimepoints.png")

g4 <- g4 + NoLegend()
g5 <- g5 + NoLegend()
g6 <- g6 + NoLegend()
g4+g6+g5

ggsave("ModalitiesMetaclusterBarPlotOverTimepointNewTimepoints.png", width = 10, height = 5)

alldata$celltype <- Idents(alldata)


cells <- as.data.frame(table(alldata$metaclusters,alldata$modality))

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4


#ggsave(filename = paste0(name, "Subset_RatioOverModalityVer2.tiff"))



cells <- as.data.frame(table(alldata$metaclusters,alldata$condition))

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4


#ggsave(filename = paste0(name, "Subset_RatioOverConditionsyVer2.tiff"))

```

LineGraphs

```{r}
library(ggpattern)
metadata <- alldata@meta.data

cols = c("mediumblue", "turquoise", "coral2", "springgreen2", "gold2", "mediumpurple2", "orange2", "orchid2", "darkseagreen2" ,"olivedrab2", "grey")

result_df <- metadata %>%
  group_by(orig.ident, modality, metaclusters) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident) %>%
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(modality, metaclusters) %>%
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

ggplot(result_df, aes(x = metaclusters, y = AverageCount, fill = metaclusters)) +
  geom_col_pattern(
    aes(fill = metaclusters, pattern_density = modality),
    colour          = 'black', 
    size            = 0.5,  
    position        = "stack",
    pattern_fill    = 'white',
    pattern_colour  = NA
  ) +
  theme_bw() +
  scale_fill_manual(values = cols, guide = FALSE) +
  labs(
    x = "Cell Type",
    y = "Average Number of Cells",
    fill = "Cell Type"
  )  + 
  coord_flip() +
  scale_pattern_manual(values = c(snRNAseq = "none", scRNAseq = "stripe", CITEseq = "circle")) +  # Specify patterns for modalities
  scale_pattern_density_manual(values = c(snRNAseq = 0, scRNAseq = 0.5, CITEseq = 1)) +
  theme(axis.text.y = element_text(size = 12), panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# Save the plot
ggsave("Global_ModalityCelltype_Averages.tiff", width = 7, height = 7)

ggplot(result_df, aes(x = metaclusters, y = AverageProportion, fill = modality)) +
  geom_bar(
    stat = "identity",
    position = "fill",
    colour = 'black',
    size = 0.5) +
  theme_bw()  +
  labs(
    x = "Cell Type",
    y = "Average Proportion",
    fill = "Modality"
  ) +
  coord_flip() +
  theme(
    axis.text.y = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  scale_y_continuous(labels = scales::percent_format())  # Convert y-axis to percentage format

# Save the plot
ggsave("Global_ModalityCMetaclusters_Proportions.tiff", width = 7, height = 7)
```

Repeat above plot, faceted by timepoint

```{r}
result_df <- metadata %>%
  group_by(orig.ident, timepoint, modality, metaclusters) %>%   # Include timepoint here
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident, timepoint) %>%                      # Also include timepoint here
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(modality, timepoint, metaclusters) %>%              # Include timepoint in the final grouping
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

ggplot(result_df, aes(x = metaclusters, y = AverageCount, fill = metaclusters)) +
  geom_col_pattern(
    aes(fill = metaclusters, pattern_density = modality),
    colour          = 'black', 
    size            = 0.5,  
    position        = "stack",
    pattern_fill    = 'white',
    pattern_colour  = NA
  ) +
  theme_bw() +
  scale_fill_manual(values = cols, guide = FALSE) +
  labs(
    x = "Cell Type",
    y = "Average Number of Cells",
    fill = "Cell Type"
  )  + 
  coord_flip() +
  facet_wrap(~timepoint, ncol = 3) +
  scale_pattern_manual(values = c(snRNAseq = "none", scRNAseq = "stripe", CITEseq = "circle")) +  # Specify patterns for modalities
  scale_pattern_density_manual(values = c(snRNAseq = 0, scRNAseq = 0.5, CITEseq = 1)) +
  theme(axis.text.y = element_text(size = 12), panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# Save the plot
ggsave("Global_ModalityMetaclusters_Averages_TimepointFacet.tiff", width = 16, height = 14)

```

```{r}
# Extract meta data information from unaltered data (alldata) and data downsampled to size of smallest batch (alldata_downsample)
min.cells <- min(table(alldata$orig.ident))
#Set identities to the meta data that we want to downsample on
Idents(alldata) <- "orig.ident"
#Subset the data, specifying that we want to downsample to the smallest batch size
alldata_downsample <- subset(alldata, downsample = min.cells)
#View the sizes of all the downsampled batches
table(alldata_downsample$orig.ident)
```

```{r}
# Extract meta data information from unaltered data (alldata) and data downsampled to size of smallest batch (alldata_downsample)
metadata <- alldata@meta.data
metadata_downsample <- alldata_downsample@meta.data

# Set color palette
pal <- c("royalblue2", "turquoise2", "deepskyblue2", "red2", "springgreen2", "gold2", "mediumpurple2", "orange2", "darkseagreen2", "orchid2", "magenta2", "tomato3")

# Calculate counts, average counts, count SD, proportions, average proportions, and proportion SD in one pipeline for NOT downsampled data
result_df <- metadata %>%
  group_by(orig.ident, timepoint, metaclusters) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident) %>%
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(timepoint,metaclusters) %>%
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

# Calculate counts, average counts, count SD, proportions, average proportions, and proportion SD in one pipeline for downsampled data
result_df_down <- metadata_downsample %>%
  group_by(orig.ident, timepoint, metaclusters) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident) %>%
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(timepoint, metaclusters) %>%
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

# Define the correct order of timepoints
timepoint_levels <- c("Unwounded", "Wounded_D1PW", "Wounded_D2PW", "Wounded_D4PW", "Wounded_D7PW", "Wounded_D15PW", "Wounded_D30PW")

# Ensure that timepoint columns are factors with the correct levels
result_df$timepoint <- factor(result_df$timepoint, levels = timepoint_levels)
result_df_down$timepoint <- factor(result_df_down$timepoint, levels = timepoint_levels)

# Create ggplot for average cell type counts with error bars (downsampled data)
ggplot(result_df, aes(x = timepoint, y = AverageCount, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = AverageCount - Count_SD, ymax = AverageCount + Count_SD, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  #geom_errorbar(aes(ymin = AverageCount - SD, ymax = AverageCount + SD), width = 0.2) + #Can include error bars if desired
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  )  +
  labs(title = "Average Cell Type Counts Over Time",
       x = "Timepoint",
       y = "Average Cell Count") +
  NoLegend()
ggsave("Global_AverageCelltypeCount.tiff", width = 9, height = 5.5)

# Create ggplot for average cell type counts with error ribbon (downsampled data)
ggplot(result_df_down, aes(x = timepoint, y = AverageCount, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = AverageCount - Count_SD, ymax = AverageCount + Count_SD, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  #geom_errorbar(aes(ymin = AverageCount - SD, ymax = AverageCount + SD), width = 0.2) + #Can include error bars if desired
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  )  +
  labs(title = "Average Downsampled Cell Type Counts Over Time",
       x = "Timepoint",
       y = "Average Cell Count") +
  NoLegend()
ggsave("Global_AverageCelltypeCount_Downsampled.tiff", width = 9, height = 5.5)

#Create ggplot for proportions over time
ggplot(result_df, aes(x = timepoint, y = Percentage, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = Percentage - SD_Percentage, ymax = Percentage + SD_Percentage, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  ) +
  labs(title = "Average Cell Type Proportions Over Time",
       x = "Timepoint",
       y = "Average Proportion (% of Total)") +
  NoLegend()
ggsave("Global_AverageCelltypeProp.tiff", width = 9, height = 5.5)

#Create ggplot for proportions over time using downsampled data
ggplot(result_df_down, aes(x = timepoint, y = Percentage, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = Percentage - SD_Percentage, ymax = Percentage + SD_Percentage, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  ) +
  labs(
    title = "Averaged Downsampled Cell Type Proportions Over Time",
    x = "Timepoint",
    y = "Average Proportion (% of Total)"
  ) +
  NoLegend()

ggsave("Global_AverageCelltypeProp_Downsampled.tiff", width = 9, height = 5.5)
#Save the dataframe as an xlsx file
writexl::write_xlsx(list(result_df, result_df_down), "Global_CellCountsProps.xlsx")

```

```{r}
alldata <- BuildClusterTree(alldata, reduction = "pca")
PlotClusterTree(alldata, direction = "rightwards")
tiff("GlobalSubset_PhyloTree.tiff")
dev.off()
```

```{r}

set.seed()
levels(alldata$orig.ident)
pal <- c("royalblue2", "turquoise2", "deepskyblue2", "red2", "springgreen2", "gold2", "mediumpurple2", "orange2", "darkseagreen2", "orchid2", "magenta2", "tomato3")

# Define the specific orig.ident values to keep
selected_orig_idents <- c("UW_Haensel1", "UW_Haensel2", "UW_Vu_v3","UW_Joost_1","D1PW_Justynski",
                          "D2PW_Justynski","D4PW_Haensel1", "D4PW_Haensel2", "D4PW_Haensel3", "D4PW_Vu_v3","D7PW_Vu_v2_1", "D7PW_Vu_v2_2", "D7PW_Vu_v3")

# Subset the Seurat object using the specified orig.ident values
alldata_subset <- subset(alldata, subset = orig.ident %in% selected_orig_idents)

alldata_subset <- IntegrateLayers(
  object = alldata_subset, 
  method = RPCAIntegration, 
  normalization.method = "SCT", 
  orig.reduction = "pca", 
  new.reduction = "rpca",
  verbose = F)

pcs = npcs(alldata_subset, var.total = 0.9)
 
alldata_subset <- FindNeighbors(alldata_subset, reduction = "rpca", dims = 1:pcs) %>% 
  FindClusters(resolution = 1, cluster.name = "rpca_clusters") %>%
  RunUMAP(reduction = "rpca", dims = 1:pcs, reduction.name = "umap.rpca")

Idents(alldata_subset) = alldata_subset$metaclusters
DimPlot(alldata_subset, cols = pal, reduction = "umap.rpca", pt.size = .3, label = F) + NoLegend() + NoAxes() 
ggsave("UMAPPublicAvaibleData.tiff")


#Subset Singel Nuceli Data
# Subset the Seurat object using the specified orig.ident values
alldata_subset <- subset(alldata, subset = modality %in% "snRNAseq")

alldata_subset <- IntegrateLayers(
  object = alldata_subset, 
  method = RPCAIntegration, 
  normalization.method = "SCT", 
  orig.reduction = "pca", 
  new.reduction = "rpca",
  verbose = F)

pcs = npcs(alldata_subset, var.total = 0.9)
 
alldata_subset <- FindNeighbors(alldata_subset, reduction = "rpca", dims = 1:pcs) %>% 
  FindClusters(resolution = 1, cluster.name = "rpca_clusters") %>%
  RunUMAP(reduction = "rpca", dims = 1:pcs, reduction.name = "umap.rpca")

Idents(alldata_subset) = alldata_subset$metaclusters
DimPlot(alldata_subset, cols = pal, reduction = "umap.rpca", pt.size = .3, label = F) + NoLegend() + NoAxes() 
ggsave("UMAPSingleNuceliData.tiff")

# Subset the CITEseq data and make individual UMAP
alldata_subset <- subset(alldata, subset = modality %in% "CITEseq")

alldata_subset <- IntegrateLayers(
  object = alldata_subset, 
  method = RPCAIntegration, 
  normalization.method = "SCT", 
  orig.reduction = "pca", 
  new.reduction = "rpca",
  verbose = F)

pcs = npcs(alldata_subset, var.total = 0.9)
 
alldata_subset <- FindNeighbors(alldata_subset, reduction = "rpca", dims = 1:pcs) %>% 
  FindClusters(resolution = 1, cluster.name = "rpca_clusters") %>%
  RunUMAP(reduction = "rpca", dims = 1:pcs, reduction.name = "umap.rpca")

Idents(alldata_subset) = alldata_subset$metaclusters
DimPlot(alldata_subset, cols = pal, reduction = "umap.rpca", pt.size = .3, label = F) + NoLegend() + NoAxes() 
ggsave("UMAPCITEseqData.tiff")


```

Differential Expression

```{r Differential Expression}
run_DE_analysis <- function(alldata, cell_types) {
  # Load required libraries
  library(DElegate)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(ggplot2)
  library(dplyr)
  library(Seurat)
  
  for (cell_type in cell_types) {
    # Subset data based on the specified cell type
    Idents(alldata) <- "metaclusters"
    alldata.1 <- subset(alldata, ident = c(cell_type))
    
    # Perform DE analysis for different comparisons
    de_results.04 <- DElegate::findDE(
      object = alldata.1, 
      compare = c("D4PW", "UW"),
      method = "deseq",
      group_column = "timepoint",
      replicate_column = "orig.ident"
    )
    
    de_results.47 <- DElegate::findDE(
      object = alldata.1, 
      compare = c("D7PW", "D4PW"),
      method = "deseq",
      group_column = "timepoint",
      replicate_column = "orig.ident"
    )
    
    de_results.715 <- DElegate::findDE(
      object = alldata.1, 
      compare = c("D15PW", "D7PW"),
      method = "deseq",
      group_column = "timepoint",
      replicate_column = "orig.ident"
    )
    
    de_results.1530 <- DElegate::findDE(
      object = alldata.1, 
      compare = c("D30PW", "D15PW"),
      method = "deseq",
      group_column = "timepoint",
      replicate_column = "orig.ident"
    )
    
    # Save DE results as CSV files
    write.csv(de_results.04, file = paste0("DE_results_", gsub(" ", "_", cell_type), "_UWOvsD4PW.csv"), row.names = FALSE)
    write.csv(de_results.47, file = paste0("DE_results_", gsub(" ", "_", cell_type), "_D4PWvsD7PW.csv"), row.names = FALSE)
    write.csv(de_results.715, file = paste0("DE_results_", gsub(" ", "_", cell_type), "_D7PWvsD15PW.csv"), row.names = FALSE)
    write.csv(de_results.1530, file = paste0("DE_results_", gsub(" ", "_", cell_type), "_D15PWvsD30PW.csv"), row.names = FALSE)
    
    # Set filtering thresholds
    x <- 1
    y <- 0.01
    
    # Subset DE results based on thresholds
    res.04 <- subset(de_results.04, padj < y & log_fc > x)
    res.47 <- subset(de_results.47, padj < y & log_fc > x)
    res.715 <- subset(de_results.715, padj < y & log_fc > x)
    res.1530 <- subset(de_results.1530, padj < y & log_fc > x)
    
    # Convert SYMBOL to ENTREZID
    ids.04 <- bitr(res.04$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
    ids.47 <- bitr(res.47$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
    ids.715 <- bitr(res.715$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
    ids.1530 <- bitr(res.1530$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
    
    # Create a gene list
    genelist <- list(
      UWOvsD4PW = ids.04$ENTREZID,
      D4PWvsD7PW = ids.47$ENTREZID,
      D7PWvsD15PW = ids.715$ENTREZID,
      D15PWvsD30PW = ids.1530$ENTREZID
    )
    
    # Run ClusterProfiler
    ck <- compareCluster(geneCluster = genelist, fun = enrichGO, OrgDb = org.Mm.eg.db,
                         pvalueCutoff = 0.05, ont = "BP")
    ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")
    
    # Plot dotplot
    s1 <- dotplot(ck, showCategory = 5, font.size = 12, label_format = 100) + 
      theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm"))
    
    s2 <- s1 + 
      theme(axis.text.x = element_text(size = 12, face = "bold", color = "black", angle = 45)) + 
      theme(axis.text.y = element_text(size = 12, face = "bold", color = "black"))
    
    print(s2)
    
    # Save plot
    ggsave(paste0("GO_", gsub(" ", "_", cell_type), ".tiff"), width = 10, height = 5)
    
    # Save cluster summary to a CSV file
    cluster_summary <- data.frame(ck)
    write.csv(cluster_summary, file = paste0("GO_", gsub(" ", "_", cell_type), ".csv"), row.names = FALSE)
  }
}

# Define cell types to iterate over
cell_types <- c("Keratinocytes", "Immune", "Sebocytes", "Fibroblasts",
                "Schwann cells", "Endothelial", "Pericytes",
                "Muscle", "Melanocytes", "Adipocytes")

# Run the function
run_DE_analysis(alldata, cell_types)


```

Running GO using FindMarkers for Each Timepoint

```{r}

#Differntial Expression
library(clusterProfiler)
library(org.Mm.eg.db)
#Perform Gene Ontology Anaylsis for Each Individual Timepoint Upregulated using Find Markers
Idents(alldata.1) <- "timepoint"
DefaultAssay(alldata.1) <- "RNA"

#Find markers for celltype
cellmarkers <- FindAllMarkers(alldata.1, min.pct = 0.5, recorrect_umi = F, only.pos = T)


res.uwo = subset(cellmarkers, cluster == "Unwounded", avg_log2FC > 1)
res.wo4 = subset(cellmarkers, cluster == "Wounded_D4PW", avg_log2FC > 1)
res.wo7 = subset(cellmarkers, cluster == "Wounded_D7PW", avg_log2FC > 1)
res.wo15 = subset(cellmarkers, cluster == "Wounded_D15PW", avg_log2FC > 1)
res.wo30 = subset(cellmarkers, cluster == "Wounded_D30PW", avg_log2FC > 1)


ids.uwo<-bitr(res.uwo$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo4<-bitr(res.wo4$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo7<-bitr(res.wo7$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo15<-bitr(res.wo15$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo30<-bitr(res.wo30$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)

genelist = list(UWO = ids.uwo$ENTREZID, 
           D4PW = ids.wo4$ENTREZID,
           D7PW = ids.wo7$ENTREZID,
           D15PW = ids.wo15$ENTREZID,
           D30PW = ids.wo30$ENTREZID)
           

# Run ClusterProifler

ck <- compareCluster(geneCluster = genelist, fun = enrichGO, OrgDb = org.Mm.eg.db, pvalueCutoff = 0.05, ont = "BP")

ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType="ENTREZID")
dotplot(ck) 
ck <- simplify(ck)

s1 = dotplot(ck, showCategory = 10, font.size = 18, label_format = 100)
s1 = s1 + theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
s2 = s1 + theme(axis.text.x = element_text(size=18, face="bold", color = "black"))
s2 = s2 + theme(axis.text.y = element_text(size=18, face="bold", color = "black"))
s2


# Convert to data.frame
ck_df <- as.data.frame(ck)

# Find unique pathways for each cluster
unique_df <- ck_df %>%
  group_by(ID) %>%
  filter(n_distinct(Cluster) == 1) %>%
  ungroup()


ck@compareClusterResult <- unique_df
ck <- simplify(ck)

# Create plot
s1 <- dotplot(ck, showCategory = 5, font.size = 12, label_format = 100) +
  scale_color_gradientn(colors = c("blue", "white", "red")) +  # Blue to red gradient
  theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm"))

s2 <- s1 +
  theme(
    axis.text.x = element_text(size = 12, face = "bold", color = "black", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12, face = "bold", color = "black")
  )

s2 + customtheme


ggsave("GeneOntologyGlobalPathwaysOvertime5Pathways.png", width = 12, height = 7)

ck_df <- as.data.frame(ck)

# Save to CSV
write.csv(ck_df, file = "GlobalGOcompareCluster_results.csv", row.names = FALSE)

```


```{r Gene Ontology TimeCourse Adipocytes}
library(dplyr) 
library(ggplot2)
gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Adipocytes.csv")


top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster     = factor(Cluster,     levels = unique(Cluster)),
    Description = factor(Description, levels = unique(Description))
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "olivedrab3",
    high = "olivedrab4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_AdipocytesOvertimeDotplot.png", p, width = 10, height = 5)

```

```{r Gene Ontology TimeCourse Endothelial}
library(dplyr) 
library(ggplot2)
gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Endothelial_cells.csv")


top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "gold2",
    high = "gold3",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_EndothelialOvertimeDotplot.png", p, width = 10, height = 10)

```

```{r Gene Ontology TimeCourse Fibroblasts}
library(dplyr) 
library(ggplot2)
gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology (Delegate By Metacluster)/Fibroblasts_GO_Upregulated.csv")


top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("DPW4vsUWO", "DPW7vsD4PW", "DPW15vsD7PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "coral2",
    high = "coral4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_FibroblastsOvertimeDotplot.png", p, width = 10, height = 10)

```

```{r Gene Ontology TimeCourse HF}
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_HF.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "skyblue1",
    high = "skyblue3",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_HFOvertimeDotplot.png", p, width = 10, height = 10)

```

```{r Gene Ontology TimeCourse IFE }
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_IFE.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "royalblue1",
    high = "royalblue4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_IFEOvertimeDotplot.png", p, width = 10, height = 10)

```

```{r Gene Ontology TimeCourse Immune Cells }
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Immune cells.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "springgreen1",
    high = "springgreen4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_ImmuneCellsOvertimeDotplot.png", p, width = 10, height = 10)

```

```{r Gene Ontology TimeCourse Melanocytes }
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Melanocytes.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "orchid1",
    high = "orchid4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_MelanocytesOvertimeDotplot.png", p, width = 10, height = 10)
```

```{r Gene Ontology TimeCourse Muscle Cells }
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Muscle_cells.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "orange1",
    high = "orange3",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_MuscleCellsOvertimeDotplot.png", p, width = 10, height = 10)
```

```{r Gene Ontology TimeCourse Perciytes}
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Pericytes.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "mediumpurple1",
    high = "mediumpurple4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_PericytesOvertimeDotplot.png", p, width = 10, height = 10)
```

```{r Gene Ontology TimeCourse RBCs}
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_RBCs.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "tomato1",
    high = "tomato4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_RBCsOvertimeDotplot.png", p, width = 10, height = 10)
```

```{r Gene Ontology TimeCourse Schwann Cellss}
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Schwann cells.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "seagreen1",
    high = "seagreen4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_SchwannCellsOvertimeDotplot.png", p, width = 10, height = 10)
```


```{r Gene Ontology TimeCourse Sebocytes}
library(dplyr) 
library(ggplot2)
library(clusterProfiler)

gene.ont = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Sebocytes.csv")

top_n <- 10

top_go <- gene.ont %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = top_n, with_ties = FALSE) %>%
  ungroup()



# (optional) ensure plotting order matches your data
top_go <- top_go %>%
  mutate(
    Cluster = factor(Cluster, levels = c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")),
    Description = factor(Description, levels = unique(Description))  # preserves order of appearance
  )

p <- ggplot(top_go, aes(x = Cluster, 
                        y = Description)) +
  geom_point(
    aes(size = Count, fill = pvalue),
    shape  = 21,        # circle with fill & outline
    color  = "black",   # dot outline
    stroke = 0.5        # outline thickness
  ) +
  scale_fill_gradient(
    low  = "deepskyblue1",
    high = "deepskyblue4",
    name = "p‑value"
  ) +
  scale_size_continuous(name = "Gene count") +
  theme_bw(base_size = 14) +
  theme(
    text            = element_text(family = "Arial", face = "bold", color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y     = element_text(hjust = 1, color = "black"),
    axis.title      = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  ) +
  ggtitle("Top GO pathways per condition")

# display
print(p)

# save
ggsave("GO_SeboctytesOvertimeDotplot.png", p, width = 10, height = 10)
```

```{r}
library(clusterProfiler)
library(org.Mm.eg.db)
#Perform Gene Ontology Anaylsis for Each Individual Timepoint Upregulated using Find Markers
Idents(alldata) <- "timepoint"
DefaultAssay(alldata) <- "RNA"

#Find markers for celltype
cellmarkers <- FindAllMarkers(alldata, min.pct = 0.2, recorrect_umi = F, only.pos = T)

# Subset for each condition with avg_log2FC > 1 and pct.1 > 0.3
res.uwo <- subset(cellmarkers, cluster == "Unwounded" & avg_log2FC > 0 & pct.1 > 0.1)
res.wo4 <- subset(cellmarkers, cluster == "Wounded_D4PW" & avg_log2FC > 1 & pct.1 > 0.1)
res.wo7 <- subset(cellmarkers, cluster == "Wounded_D7PW" & avg_log2FC > 0.5 & pct.1 > 0.1)
res.wo15 <- subset(cellmarkers, cluster == "Wounded_D15PW" & avg_log2FC > 0 & pct.1 > 0.1)
res.wo30 <- subset(cellmarkers, cluster == "Wounded_D30PW" & avg_log2FC > 0 & pct.1 > 0.1)

# Get top 20 genes by avg_log2FC for each condition
top.uwo <- head(res.uwo[order(-res.uwo$avg_log2FC), ], 30)
top.wo4 <- head(res.wo4[order(-res.wo4$avg_log2FC), ], 30)
top.wo7 <- head(res.wo7[order(-res.wo7$avg_log2FC), ], 30)
top.wo15 <- head(res.wo15[order(-res.wo15$avg_log2FC), ], 30)
top.wo30 <- head(res.wo30[order(-res.wo30$avg_log2FC), ], 30)

# Optionally, bind all together into one dataframe with a column for the condition
top.genes <- rbind(
  transform(top.uwo, condition = "Unwounded"),
  transform(top.wo4, condition = "Wounded_D4PW"),
  transform(top.wo7, condition = "Wounded_D7PW"),
  transform(top.wo15, condition = "Wounded_D15PW"),
  transform(top.wo30, condition = "Wounded_D30PW")
)

# View or extract just gene names if needed
genes <- unique(top.genes$gene)

alldata <- ScaleData(alldata, features = genes)
genes <- setdiff(genes, c("Fth1", "Gm42418"))

p1 = DoHeatmap(alldata, features = genes, size = 5, group.by = "timepoint") + 
  scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(n = 7, name = "RdBu"))) +
  theme(
    axis.text.x = element_text(face = "bold", size = 12, family = "Arial"),
    axis.text.y = element_text(face = "bold", size = 12, family = "Arial"),
    legend.text = element_text(face = "bold", size = 12, family = "Arial"),
    legend.title = element_text(face = "bold", size = 12, family = "Arial"),
    plot.title = element_text(face = "bold", size = 12, family = "Arial"),
    strip.text = element_text(face = "bold", size = 12, family = "Arial")
  )

ggsave("HeatmapAcrossTimepointMajorMarkers.png", p1, width = 20, height = 25)


# Generate the DotPlot
library(Seurat)
library(ggplot2)
library(RColorBrewer)

# Create initial DotPlot to get the data
p <- DotPlot(alldata, features = unique(genes), dot.scale = 7, group.by = "timepoint_modality")
plot_data <- p$data  # Extract DotPlot data

# Invert red and blue in the RdYlBu palette
inverted_colors <- rev(brewer.pal(n = 11, name = "RdYlBu"))

# Rebuild the plot with black outlines and inverted colors
p <- ggplot(plot_data, aes(x = id, y = features.plot)) +
  geom_point(
    aes(size = pct.exp, fill = avg.exp.scaled),
    shape = 21,         # filled circle
    color = "black",    # black outline
    stroke = 0.5
  ) +
  scale_fill_gradientn(colors = inverted_colors) +
  scale_size(range = c(0, 7)) +
  theme_bw() +
  RotatedAxis() +
  theme(
    text = element_text(size = 12, face = "bold", family = "Arial"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_text(face = "bold", family = "Arial"),
    legend.text = element_text(family = "Arial")
  ) +
  guides(
    fill = guide_colorbar(title = "Average Expression"),
    size = guide_legend(title = "Percent Expressed")
  ) +
  labs(x = "Timepoint", y = "Gene") + customtheme

ggsave("DotPlotAcrossTimePointModalityMajorMarkers.png", p, height = 20, width = 8)




```

```{r}

```



```{r}
#Load PSU Dataset
alldata.1 = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1-2/PSUUnit.rds")
#Change identity levels so that they are grouped without cell cluster
Idents(alldata.1) <- gsub("\\s+[IVX]+$", "", Idents(alldata))
Idents(alldata.1) <- gsub("\\s+[0-9]+$", "", Idents(alldata))
levels(alldata.1)
```

Run GSEA analysis

```{r}
library(clusterProfiler)
library(DElegate)
library(org.Mm.eg.db)


celltype <- "PSU"
#In case the celltype column in the metadata has a different name 
identity_metadata <- "celltype"


alldata$celltype <- Idents(alldata)

alldata.1 <- subset(alldata, idents = "PSU")


de_results.04 <- DElegate::findDE(object = alldata.1, compare = c("Wounded_DPW4","Unwounded"), method = "deseq", group_column = "timepoint",replicate_column = "orig.ident")
de_results.47 <- DElegate::findDE(object = alldata.1, compare = c("Wounded_DPW7","Wounded_D4PW"), method = "deseq", group_column = "timepoint",replicate_column = "orig.ident")
de_results.715 <- DElegate::findDE(object = alldata.1, compare = c("Wounded_DPW15","Wounded_D7PW"), method = "deseq", group_column = "timepoint",replicate_column = "orig.ident")
de_results.1530 <- DElegate::findDE(object = alldata.1, compare = c("Wounded_DPW30","Wounded_D15PW"), method = "deseq", group_column = "timepoint",replicate_column = "orig.ident")



x = 1
y = .01

res.01 = subset(de_results.01, log_fc > x & pvalue < y)
res.02 = subset(de_results.02, log_fc > x & pvalue < y)
res.04 = subset(de_results.04, log_fc > x & pvalue < y)
res.07 = subset(de_results.07, log_fc > x & pvalue < y)
res.015 = subset(de_results.015, log_fc > x & pvalue < y)
res.030 = subset(de_results.030, log_fc > x & pvalue < y)


res.47 = intersect(res.04$feature,res.07$feature)
res.715 = intersect(res.04$feature,res.07$feature)

ids.04<-bitr(res.04$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.07<-bitr(res.07$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.015<-bitr(res.015$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.030<-bitr(res.030$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)


genelist = list(UWOvsDPW4 = ids.04$ENTREZID, 
           UWOvsDPW7 = ids.07$ENTREZID,
           UWOvsDPW15 = ids.015$ENTREZID,
           UWOvsDPW30 = ids.030$ENTREZID)

# Run ClusterProifler

ck <- compareCluster(geneCluster = genelist, fun = enrichGO, OrgDb = org.Mm.eg.db, pvalueCutoff = 0.05, ont = "BP")

ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType="ENTREZID")
dotplot(ck) 
s1 = dotplot(ck, showCategory = 10, font.size = 18, label_format = 100)
s1 = s1 + theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
s2 = s1 + theme(axis.text.x = element_text(size=18, face="bold", color = "black"))
s2 = s2 + theme(axis.text.y = element_text(size=18, face="bold", color = "black"))
s2

ggsave("PSUDotPlotUpregulatedGenesBetweenTimepoints.tiff", width = 16)

x = -1
y = .01

res.01 = subset(de_results.01, log_fc > x & pvalue < y)
res.02 = subset(de_results.02, log_fc > x & pvalue < y)
res.04 = subset(de_results.04, log_fc > x & pvalue < y)
res.07 = subset(de_results.07, log_fc > x & pvalue < y)
res.015 = subset(de_results.015, log_fc > x & pvalue < y)
res.030 = subset(de_results.030, log_fc > x & pvalue < y)


res.47 = intersect(res.04$feature,res.07$feature)
res.715 = intersect(res.04$feature,res.07$feature)

ids.04<-bitr(res.04$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.07<-bitr(res.07$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.015<-bitr(res.015$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.030<-bitr(res.030$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)


genelist = list(UWOvsDPW4 = ids.04$ENTREZID, 
           UWOvsDPW7 = ids.07$ENTREZID,
           UWOvsDPW15 = ids.015$ENTREZID,
           UWOvsDPW30 = ids.030$ENTREZID)

# Run ClusterProifler

ck <- compareCluster(geneCluster = genelist, fun = enrichGO, OrgDb = org.Mm.eg.db, pvalueCutoff = 0.05, ont = "BP")

ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType="ENTREZID")
dotplot(ck) 
s1 = dotplot(ck, showCategory = 10, font.size = 18, label_format = 100)
s1 = s1 + theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
s2 = s1 + theme(axis.text.x = element_text(size=18, face="bold", color = "black"))
s2 = s2 + theme(axis.text.y = element_text(size=18, face="bold", color = "black"))
s2

ggsave("PSUDotPlotDownreuglatedGenesBetweenTimepoints.tiff", width = 16)
```

```{r R loop for determining Gene Ontology Enrichment For Each Metacluster}
library(clusterProfiler)
library(DElegate)
library(org.Mm.eg.db)
library(ggplot2)
library(dplyr)

# Define cell types
cell_types <- c(
  "IFE Keratinocytes", "HF Keratinocytes", "Sebocytes", "Fibroblasts",
  "Immune", "Endothelial", "Pericytes", "Muscle",
  "Melanocytes", "Schwann cells", "Adipocytes", "RBCs"
)

# Timepoint comparisons
comparisons <- list(
  "DPW4vsUWO"    = c("Wounded_D4PW", "Unwounded"),
  "DPW7vsD4PW"   = c("Wounded_D7PW", "Wounded_D4PW"),
  "DPW15vsD7PW"  = c("Wounded_D15PW", "Wounded_D7PW"),
  "DPW30vsD15PW" = c("Wounded_D30PW", "Wounded_D15PW")
)

# Prepare DE analysis for each cell type
for (celltype in cell_types) {
  message("Processing: ", celltype)
  
  alldata$celltype <- Idents(alldata)
  alldata_subset <- subset(alldata, ident = celltype)
  
  celltype_clean <- gsub(" ", "_", celltype)
  de_list <- list()

  # --- Differential Expression ---
  for (comp_name in names(comparisons)) {
    comp <- comparisons[[comp_name]]

    de_result <- tryCatch({
      DElegate::findDE(
        object = alldata_subset,
        compare = comp,
        method = "deseq",
        group_column = "timepoint"
      )
    }, error = function(e) {
      warning("Skipping comparison ", comp_name, " due to error: ", e$message)
      return(NULL)
    })

    if (!is.null(de_result)) {
      write.csv(de_result, paste0(celltype_clean, "_", comp_name, "_DE_results.csv"), row.names = FALSE)
    }

    de_list[[comp_name]] <- de_result
  }

  # ---- UPREGULATED ----
  up_genelist <- list()
  for (comp_name in names(de_list)) {
    de_df <- de_list[[comp_name]]
    if (!is.null(de_df)) {
      up_df <- subset(de_df, log_fc > 1 & pvalue < 0.01)
      if (nrow(up_df) > 0) {
        entrez <- bitr(up_df$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
        up_genelist[[comp_name]] <- entrez$ENTREZID
      }
    }
  }

  if (length(up_genelist) > 0) {
    ck_up <- compareCluster(
      geneCluster = up_genelist,
      fun = enrichGO,
      OrgDb = org.Mm.eg.db,
      pvalueCutoff = 0.05,
      ont = "BP"
    )
    ck_up <- setReadable(ck_up, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")

    # Save GO enrichment results
    write.csv(as.data.frame(ck_up), paste0(celltype_clean, "_GO_Upregulated.csv"), row.names = FALSE)

    s2_up <- dotplot(ck_up, showCategory = 10, font.size = 18, label_format = 100) +
      theme(
        plot.margin = grid::unit(c(0, 0, 0, 0), "mm"),
        axis.text.x = element_text(size = 18, face = "bold", color = "black"),
        axis.text.y = element_text(size = 18, face = "bold", color = "black")
      )

    ggsave(
      filename = paste0(celltype_clean, "_DotPlot_Upregulated_Timepoints.png"),
      plot = s2_up,
      width = 16,
      height = 20
    )
  }

  # ---- DOWNREGULATED ----
  down_genelist <- list()
  for (comp_name in names(de_list)) {
    de_df <- de_list[[comp_name]]
    if (!is.null(de_df)) {
      down_df <- subset(de_df, log_fc < -1 & pvalue < 0.01)
      if (nrow(down_df) > 0) {
        entrez <- bitr(down_df$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
        down_genelist[[comp_name]] <- entrez$ENTREZID
      }
    }
  }

  if (length(down_genelist) > 0) {
    ck_down <- compareCluster(
      geneCluster = down_genelist,
      fun = enrichGO,
      OrgDb = org.Mm.eg.db,
      pvalueCutoff = 0.05,
      ont = "BP"
    )
    ck_down <- setReadable(ck_down, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")

    # Save GO enrichment results
    write.csv(as.data.frame(ck_down), paste0(celltype_clean, "_GO_Downregulated.csv"), row.names = FALSE)

    s2_down <- dotplot(ck_down, showCategory = 10, font.size = 18, label_format = 100) +
      theme(
        plot.margin = grid::unit(c(0, 0, 0, 0), "mm"),
        axis.text.x = element_text(size = 18, face = "bold", color = "black"),
        axis.text.y = element_text(size = 18, face = "bold", color = "black")
      )

    ggsave(
      filename = paste0(celltype_clean, "_DotPlot_Downregulated_Timepoints.png"),
      plot = s2_down,
      width = 16,
      height = 20
    )
  }
}


```


```{r}
library(DElegate)
library(org.Mm.eg.db)
library(clusterProfiler)
library(ggVennDiagram)
library(ggplot2)
library(dplyr)

# Define comparisons against Unwounded
comparisons <- list(
  "D1PW"  = c("Wounded_D1PW", "Unwounded"),
  "D2PW"  = c("Wounded_D2PW", "Unwounded"),
  "D4PW"  = c("Wounded_D4PW", "Unwounded"),
  "D7PW"  = c("Wounded_D7PW", "Unwounded"),
  "D15PW" = c("Wounded_D15PW", "Unwounded"),
  "D30PW" = c("Wounded_D30PW", "Unwounded")
)

# Initialize list to store upregulated gene sets
upregulated_genes_list <- list()


options(future.globals.maxSize = 8000 * 1024^2)

# Loop through comparisons and perform DE
for (tp in names(comparisons)) {
  comp <- comparisons[[tp]]
  message("Processing comparison: ", tp)

  # Try DE analysis
  de_result <- tryCatch({
    DElegate::findDE(
      object = alldata,
      compare = comp,
      method = "deseq",
      group_column = "timepoint"
    )
  }, error = function(e) {
    warning("Skipping ", tp, ": ", e$message)
    return(NULL)
  })

  if (!is.null(de_result)) {
    # Filter for upregulated genes
    up_df <- subset(de_result, log_fc > 0 & pvalue < 0.01)
    if (nrow(up_df) > 0) {
      upregulated_genes_list[[tp]] <- up_df$feature
    } else {
      upregulated_genes_list[[tp]] <- character(0)
    }
  }
}

# Optional: Save the upregulated gene lists
for (tp in names(upregulated_genes_list)) {
  write.csv(upregulated_genes_list[[tp]],
            file = paste0("UpregulatedGenes_", tp, "_vs_UW.csv"),
            row.names = FALSE)
}

# Plot Venn diagram
# Filter to timepoints with non-empty gene sets
filtered_upsets <- upregulated_genes_list[sapply(upregulated_genes_list, length) > 0]


# Chronological order
ordered_timepoints <- c("D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")

# Reorder filtered_upsets to match
filtered_upsets <- filtered_upsets[ordered_timepoints]

library(ggVennDiagram)
library(ggplot2)

venn_plot_bold_labels <- ggVennDiagram(
  filtered_upsets,
  label = "count",
  category.names = names(filtered_upsets),
  category.col = "black",          # label color
  category_fontface = "bold",      # make set labels bold
  label_geom = "text",
  label_alpha = 1,
  label_size = 5,
  label_color = "black",
  label_fontface = "bold"
) +
  scale_fill_gradient(low = "#D0F0C0", high = "#006400") +
  theme(
    text = element_text(size = 14),
    legend.position = "none"
  )

ggsave(
  filename = "6_SetVennDiagram_green.svg",  # name of the file
  plot = venn_plot_bold_labels,
  width = 10,     # width in inches
  height = 8,     # height in inches
  dpi = 300       # resolution
)


# Create a new list where D1PW and D2PW are merged
filtered_upsets_grouped <- filtered_upsets

# Merge the two sets
filtered_upsets_grouped$Inflammation <- unique(c(
  filtered_upsets$D1PW,
  filtered_upsets$D2PW
))

# Remove the original D1PW and D2PW entries
filtered_upsets_grouped <- filtered_upsets_grouped[!names(filtered_upsets_grouped) %in% c("D1PW", "D2PW")]

# Optional: reorder timepoints if you like
filtered_upsets_grouped <- filtered_upsets_grouped[c("Inflammation", "D4PW", "D7PW", "D15PW", "D30PW")]

# Plot
venn_plot_grouped <- ggVennDiagram(
  filtered_upsets_grouped,
  label = "count",
  category.names = names(filtered_upsets_grouped),
  category.col = "black",
  category_fontface = "bold",
  label_geom = "text",
  label_alpha = 1,
  label_size = 5,
  label_color = "black",
  label_fontface = "bold"
) +
  scale_fill_gradient(low = "#D0F0C0", high = "#006400") +
  theme(
    text = element_text(size = 14),
    legend.position = "none"
  )

# Save the plot
ggsave("6set_VennDiagram_green_grouped_inflammation.svg", plot = venn_plot_grouped, width = 10, height = 8, dpi = 300)



```







```{r}
library(clusterProfiler)
library(org.Mm.eg.db)


Idents(alldata) <- alldata$orig.ident
alldata = subset(alldata, ident = c("UW_snRNAseq_2","D4PW_snRNAseq_1","D4PW_snRNAseq_2"), invert = T)

  # Perform DE analysis for different comparisons
  de_results.04 <- DElegate::findDE(
    object = alldata, 
    compare = c("Wounded_D4PW", "Unwounded"),
    method = "deseq",
    group_column = "timepoint",
    replicate_column = "orig.ident"
  )
  de_results.47 <- DElegate::findDE(
    object = alldata, 
    compare = c("Wounded_D7PW", "Unwounded"),
    method = "deseq",
    group_column = "timepoint"
  )
  de_results.715 <- DElegate::findDE(
    object = alldata, 
    compare = c("Wounded_D15PW", "Unwounded"),
    method = "deseq",
    group_column = "timepoint"
  )
  de_results.1530 <- DElegate::findDE(
    object = alldata, 
    compare = c("Wounded_D30PW", "Unwounded"),
    method = "deseq",
    group_column = "timepoint"
  )
  
   # Set filtering thresholds
  x <- 1
  y <- 0.01
  
  # Subset DE results based on thresholds
  res.04 <- subset(de_results.04, padj < y & log_fc > x)
  res.47 <- subset(de_results.47, padj < y & log_fc > x)
  res.715 <- subset(de_results.715, padj < y & log_fc > x)
  res.1530 <- subset(de_results.1530, padj < y & log_fc > x)
  
  # Convert SYMBOL to ENTREZID
  ids.04 <- bitr(res.04$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.47 <- bitr(res.47$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.715 <- bitr(res.715$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.1530 <- bitr(res.1530$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  
  # Create a gene list
  genelist <- list(
    UWOvsD4PW = ids.04$ENTREZID,
    UWOvsD7PW = ids.47$ENTREZID,
    UWOvsD15PW = ids.715$ENTREZID,
    UWOvsD30PW = ids.1530$ENTREZID
  )
  
  # Run ClusterProfiler
  ck <- compareCluster(geneCluster = genelist, fun = enrichGO, OrgDb = org.Mm.eg.db, 
                       pvalueCutoff = 0.05, ont = "BP")
  ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")
  
  # Plot dotplot
  s1 <- dotplot(ck, showCategory = 5, font.size = 12, label_format = 100) + 
    theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm"))
  s2 <- s1 + 
    theme(axis.text.x = element_text(size = 12, face = "bold", color = "black", angle = 45)) + 
    theme(axis.text.y = element_text(size = 12, face = "bold", color = "black"))
  print(s2)
  
  # Save plot
  ggsave(paste0("GO_", gsub(" ", "_", "Global"), ".png"), width = 10, height = 5)
  
  # Save cluster summary to a CSV file
  cluster_summary <- data.frame(ck)
  write.csv(cluster_summary, file = paste0("GO_", gsub(" ", "_", "Global"), ".csv"), row.names = FALSE)
  
  
# Filter based on thresholds
res.04 <- subset(de_results.04, padj < y & log_fc > x)
res.47 <- subset(de_results.47, padj < y & log_fc > x)
res.715 <- subset(de_results.715, padj < y & log_fc > x)
res.1530 <- subset(de_results.1530, padj < y & log_fc > x)

# Sort by padj (ascending), then log_fc (descending), then rate1 (descending)
res.04 <- res.04[order(res.04$padj, -res.04$log_fc, -res.04$rate1), ]
res.47 <- res.47[order(res.47$padj, -res.47$log_fc, -res.47$rate1), ]
res.715 <- res.715[order(res.715$padj, -res.715$log_fc, -res.715$rate1), ]
res.1530 <- res.1530[order(res.1530$padj, -res.1530$log_fc, -res.1530$rate1), ]

# Extract top 200 genes from each
top.04 <- head(res.04$feature, 50)
top.47 <- head(res.47$feature, 50)
top.715 <- head(res.715$feature, 50)
top.1530 <- head(res.1530$feature, 50)

# Combine and get unique genes
all_genes <- c(top.04, top.47, top.715, top.1530)
unique_genes <- unique(all_genes)

# Assuming your metadata columns are named "timepoint" and "modality"
alldata$timepoint_modality <- paste(alldata$timepoint, alldata$modality, sep = "_")
new_levels <- c(
  # Unwounded
  "Unwounded_scRNAseq",
  "Unwounded_snRNAseq",
  "Unwounded_CITEseq",
  
  # Wounded_D1PW
  "Wounded_D1PW_scRNAseq",
  
  # Wounded_D2PW
  "Wounded_D2PW_scRNAseq",
  
  # Wounded_D4PW
  "Wounded_D4PW_scRNAseq",
  "Wounded_D4PW_snRNAseq",
  
  # Wounded_D7PW
  "Wounded_D7PW_scRNAseq",
  "Wounded_D7PW_snRNAseq",
  
  # Wounded_D15PW
  "Wounded_D15PW_snRNAseq",
  "Wounded_D15PW_CITEseq",
  
  # Wounded_D30PW
  "Wounded_D30PW_snRNAseq",
  "Wounded_D30PW_CITEseq"
)

alldata$timepoint_modality <- factor(alldata$timepoint_modality, levels = new_levels)


Idents(alldata) <- alldata$timepoint_modality

p <- DotPlot(alldata, unique_genes, dot.scale = 7, cols =  "RdYlBu") +RotatedAxis() +
theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") # Outlines circles in black
# Display the plot

ggsave(filename = "MajorExpressionForEachTimepointModality_RatioOvertimeVer3.png", width = 40, height = 4)
```






```{r}
library(dplyr)
library(clusterProfiler)
library(org.Mm.eg.db)
library(ggplot2)

library(dplyr)
library(clusterProfiler)
library(org.Mm.eg.db)
library(ggplot2)

# Define cell types to iterate through
cell_types <- c(
  "IFE Keratinocytes", "HF Keratinocytes", "Sebocytes", "Fibroblasts", 
  "Immune", "Endothelial", "Pericytes", "Muscle", 
  "Melanocytes", "Schwann cells", "Adipocytes", "RBCs"
)

# Make sure identity column is consistent
alldata$celltype <- Idents(alldata)

# Loop through each cell type
for (celltype in cell_types) {
  message("Processing cell type: ", celltype)

  # Subset Seurat object to just this cell type
  alldata_subset <- subset(alldata, celltype == celltype)

  # Set timepoint as the identity for DE analysis
  Idents(alldata_subset) <- alldata_subset$timepoint

  # Run differential expression comparing timepoints within the cell type
  cellmarkers_bulk <- findDE(
    alldata_subset, 
    group_column = "timepoint",
    replicate_column = "orig.ident",
    method = "deseq", 
    lfc_shrinkage = "apeglm"
  )

  # Sort by log fold change
  cellmarkers_bulk <- cellmarkers_bulk %>%
    group_by(group1) %>%
    arrange(desc(log_fc)) %>%
    ungroup()

  # GO enrichment analysis on DE genes
  gse_bulk <- compareCluster(
    feature | log_fc ~ group1, 
    data = cellmarkers_bulk, 
    fun = gseGO, 
    OrgDb = org.Mm.eg.db,
    keyType = "SYMBOL", 
    pvalueCutoff = 0.05, 
    pAdjustMethod = "BH", 
    ont = "BP", 
    nPermSimple = 10000, 
    eps = 0
  )

  # Plotting GO enrichment
  s1 <- dotplot(gse_bulk, showCategory = 3, font.size = 8, label_format = 100) +
    theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm"))

  s2 <- s1 +
    theme(
      axis.text.y = element_text(size = 11, face = "bold", color = "black"),
      axis.text.x = element_text(size = 11, face = "bold", color = "black", angle = 45)
    )

  # Save dotplot
  plot_filename <- paste0("DotPlot_", gsub(" ", "_", celltype), "_Timepoint.tiff")
  ggsave(plot_filename, s2, height = 8, width = 10)

  # Save GO enrichment summary
  cluster_summary <- data.frame(gse_bulk)
  csv_filename <- paste0("GO_", gsub(" ", "_", celltype), "_Timepoint.csv")
  write.csv(cluster_summary, file = csv_filename, row.names = FALSE)

  # Save raw DE results (optional, but helpful!)
  de_filename <- paste0("DEG_", gsub(" ", "_", celltype), "_Timepoint.csv")
  write.csv(cellmarkers_bulk, file = de_filename, row.names = FALSE)
}

```






Figure 1

```{r}

Idents(alldata) <- "metaclusters"
features = c("Krt14","Itga6", #Keratincoytes Labels
             "Krt15","Sox9",
             "Pparg", #Sebocytes
             "Col1a2","Pdgfra", #Fibroblast Markers
             "Ptprc", # Immune Cell Markers
             "Pecam1", #Endothelial Cell Markers
                          "Rgs5", #Pericytes
             "Mylpf", "Des", # Muscle Markers
                          "Mpz", #Schwann Cell Markers
             "Mlana", #Melanocyte Marker
             "Adipoq", #Adipocyte Markers
             "Hba-a1","Hbb-bt" #RBC marker
)



DotPlot(alldata, features, dot.scale = 7) + RotatedAxis()



p <- DotPlot(alldata, features, dot.scale = 7, cols = c("white", "blue3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") # Outlines circles in black

# Display the plot
p

ggsave("DotPlotAllModalityIntergrationAcrossCellTypes.tiff",dpi = 300,height = 6, width =7.3 )

```

```{r}
process_go_and_plot <- function(celltype, color, csv_path) {
  # Load GO result
  onto.genes <- read.csv(csv_path)
  
  # Rebuild and simplify
  ck_rebuilt <- new("compareClusterResult", compareClusterResult = onto.genes, fun = "enrichGO")
  ck_rebuilt <- simplify(ck_rebuilt)
  ck_df <- as.data.frame(ck_rebuilt)
  
  # Select unique pathways
  unique_pathways <- ck_df %>%
    arrange(p.adjust) %>%
    distinct(Description, .keep_all = TRUE)
  
  # Select top N per cluster
  top_n <- 2
  top_pathways <- unique_pathways %>%
    group_by(Cluster) %>%
    arrange(p.adjust) %>%
    slice_head(n = top_n) %>%
    ungroup()
  
  # Split gene list
  genes_by_pathway <- top_pathways %>%
    select(Cluster, Description, geneID) %>%
    separate_rows(geneID, sep = "/")
  
  genes_listed <- genes_by_pathway %>%
    group_by(Cluster, Description) %>%
    summarise(Genes = list(unique(geneID)), .groups = "drop")
  
  # Order clusters
  cluster_order <- c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")
  genes_listed <- genes_listed %>%
    mutate(Cluster = factor(Cluster, levels = cluster_order)) %>%
    arrange(Cluster)
  
  # Top 5 genes
  pick_top5 <- function(genes_vector) genes_vector[1:min(length(genes_vector), 5)]
  top5_genes_per_pathway <- genes_listed %>%
    rowwise() %>%
    mutate(Top5Genes = list(pick_top5(Genes))) %>%
    ungroup()
  
  unique_top5_genes <- top5_genes_per_pathway %>%
    pull(Top5Genes) %>%
    unlist() %>%
    unique() %>%
    as.character()
  
  # Subset Seurat and plot
  Idents(alldata) <- "metaclusters"
  alldata.1 <- subset(alldata, ident = celltype)
  Idents(alldata.1) <- "timepoint"
  
  p <- DotPlot(alldata.1, unique_top5_genes, dot.scale = 7, cols = c("white", color)) +
    RotatedAxis() +
    theme(text = element_text(size = 15, face = "bold", family = "Arial")) +
    scale_size(range = c(0, 7)) +
    geom_point(aes(size = pct.exp), shape = 21, color = "black")
  
  # Save plots
  ggsave(paste0("DotPlot", celltype, "GO.png"), dpi = 300, height = 3.5, width = 8)
  ggsave(paste0("DotPlot", celltype, "GOUnsupervisedList.png"), dpi = 300, height = 4.5, width = 14)
  
  # Save table
  top5_genes_per_pathway_flat <- top5_genes_per_pathway %>%
    mutate(
      Cluster = as.character(Cluster),
      Description = as.character(Description),
      Genes = sapply(Genes, function(x) paste(x, collapse = "; ")),
      Top5Genes = sapply(Top5Genes, function(x) paste(x, collapse = "; "))
    ) %>%
    as.data.frame()
  
  write.csv(top5_genes_per_pathway_flat, file = paste0(celltype, "_genes_per_pathway.csv"), row.names = FALSE)
}

```






```{r}

onto.genes = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Adipocytes.csv")
ck_rebuilt <- new(
  "compareClusterResult",
  compareClusterResult = onto.genes,
  fun = "enrichGO"
)

# Step 3: Now you can use simplify() and other clusterProfiler functions
ck_rebuilt<- simplify(ck_rebuilt)

# 1. Extract data.frame from the compareClusterResult object
ck_df <- as.data.frame(ck_rebuilt)

# 2. Ensure each pathway (Description) is unique across ALL clusters
# Keep only the first appearance (lowest p.adjust)
unique_pathways <- ck_df %>%
  arrange(p.adjust) %>%
  distinct(Description, .keep_all = TRUE)

# 3. Then for each cluster, pick top N pathways (among unique ones)
top_n <- 2# change if you want more/fewer per cluster

top_pathways <- unique_pathways %>%
  group_by(Cluster) %>%
  arrange(p.adjust) %>%
  slice_head(n = top_n) %>%
  ungroup()

# 4. Extract the gene lists
genes_by_pathway <- top_pathways %>%
  select(Cluster, Description, geneID) %>%
  separate_rows(geneID, sep = "/")  # split multiple gene IDs into rows

# 5. Summarize nicely
genes_listed <- genes_by_pathway %>%
  group_by(Cluster, Description) %>%
  summarise(Genes = list(unique(geneID)), .groups = "drop")

# 1. Define your custom cluster order
cluster_order <- c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")

# 2. Set 'Cluster' as a factor with the desired order
genes_listed <- genes_listed %>%
  mutate(Cluster = factor(Cluster, levels = cluster_order)) %>%
  arrange(Cluster)

# Function to pick top 5 genes from each pathway
pick_top5 <- function(genes_vector) {
  genes_vector[1:min(length(genes_vector), 5)]  # No sorting, just first 5
}

# Apply to each pathway
top5_genes_per_pathway <- genes_listed %>%
  rowwise() %>%
  mutate(Top5Genes = list(pick_top5(Genes))) %>%
  ungroup()

# 3. Flatten to get all genes, make them unique
unique_top5_genes <- top5_genes_per_pathway %>%
  pull(Top5Genes) %>%
  unlist() %>%
  unique() %>%
  as.character()


Idents(alldata) <- "metaclusters"
alldata.1 <- subset(alldata, ident = "Adipocytes")  


#DotPlots for Gene Ontology Cell Types
Idents(alldata.1) <- "timepoint"


features = c("Cxcl5","Cxcr2","Acp5","Il23a", #D4PW Leukocote Recruitment, Tissue Remodeling
             "Pparg","Gpc6","Ctnna3") #D7PW Proliferation
             
genes_to_remove <- c("Krt77", "Krt10","Krt5", "Krt1")

unique_top5_genes <- unique_top5_genes[!unique_top5_genes %in% genes_to_remove]

p <- DotPlot(alldata.1, unique_top5_genes, dot.scale = 7, cols = c("white", "springgreen4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotImmuneCellsGOEnhanced.png",dpi = 300,height = 4.5, width = 7 )

ggsave("DotPlotImmuneCellsGOUnsuprisedList.png",dpi = 300,height = 4.5, width = 14 )


# Now you can save it
# Force everything to be character/text (no lists allowed)
top5_genes_per_pathway_flat <- top5_genes_per_pathway %>%
  mutate(
    Cluster = as.character(Cluster),
    Description = as.character(Description),
    Genes = sapply(Genes, function(x) paste(x, collapse = "; ")),
    Top5Genes = sapply(Top5Genes, function(x) paste(x, collapse = "; "))
  ) %>%
  as.data.frame()

# Now save it
write.csv(top5_genes_per_pathway_flat, file = "ImmuneCells_genes_per_pathway.csv", row.names = FALSE)


```

Gene Ontology Dot Plots for Immune Cells

```{r}
#DotPlots for Gene Ontology Cell Types

  
alldata.1 <- subset(alldata, ident = "Immune cells")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl5","Cxcr2","Acp5","Il23a", #D4PW Leukocote Recruitment, Tissue Remodeling
             "Pparg","Gpc6","Ctnna3") #D7PW Proliferation
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "springgreen4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotImmuneCellsGO2.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots For Fibroblasts

```{r}
onto.genes = read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Fibroblasts.csv")
ck_rebuilt <- new(
  "compareClusterResult",
  compareClusterResult = onto.genes,
  fun = "enrichGO"
)

# Step 3: Now you can use simplify() and other clusterProfiler functions
ck_rebuilt<- simplify(ck_rebuilt)

# 1. Extract data.frame from the compareClusterResult object
ck_df <- as.data.frame(ck_rebuilt)

# 2. Ensure each pathway (Description) is unique across ALL clusters
# Keep only the first appearance (lowest p.adjust)
unique_pathways <- ck_df %>%
  arrange(p.adjust) %>%
  distinct(Description, .keep_all = TRUE)

# 3. Then for each cluster, pick top N pathways (among unique ones)
top_n <- 2# change if you want more/fewer per cluster

top_pathways <- unique_pathways %>%
  group_by(Cluster) %>%
  arrange(p.adjust) %>%
  slice_head(n = top_n) %>%
  ungroup()

# 4. Extract the gene lists
genes_by_pathway <- top_pathways %>%
  select(Cluster, Description, geneID) %>%
  separate_rows(geneID, sep = "/")  # split multiple gene IDs into rows

# 5. Summarize nicely
genes_listed <- genes_by_pathway %>%
  group_by(Cluster, Description) %>%
  summarise(Genes = list(unique(geneID)), .groups = "drop")

# 1. Define your custom cluster order
cluster_order <- c("UWOvsD4PW", "D4PWvsD7PW", "D7PWvsD15PW", "D15PWvsD30PW")

# 2. Set 'Cluster' as a factor with the desired order
genes_listed <- genes_listed %>%
  mutate(Cluster = factor(Cluster, levels = cluster_order)) %>%
  arrange(Cluster)

# Function to pick top 5 genes from each pathway
pick_top5 <- function(genes_vector) {
  genes_vector[1:min(length(genes_vector), 5)]  # No sorting, just first 5
}

# Apply to each pathway
top5_genes_per_pathway <- genes_listed %>%
  rowwise() %>%
  mutate(Top5Genes = list(pick_top5(Genes))) %>%
  ungroup()

# 3. Flatten to get all genes, make them unique
unique_top5_genes <- top5_genes_per_pathway %>%
  pull(Top5Genes) %>%
  unlist() %>%
  unique() %>%
  as.character()



#DotPlots for Gene Ontology Cell Types


Idents(alldata) <- "metaclusters"
  
alldata.1 <- subset(alldata, ident = "Fibroblasts")  
Idents(alldata.1) <- "timepoint"


features = c("S100a8","S100a9","Ccl22","Cxcl12", #D4PW Leukocote Recruitment, Tissue Remodeling
                          "Acta2","Nox4", #Contraction
             "Egfr","Esr1","Pdgfd","Fgf2","Pak3", #D7PW Fibroblast Profliferation/Migration
             "Wnt10b","Msx2" #Anagen
                 ) 
             


p <- DotPlot(alldata.1, unique_top5_genes, dot.scale = 7, cols = c("white", "red3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotFibroblastsGO.png",dpi = 300,height = 3.5, width = 8 )

ggsave("DotPlotFibroblastsGOUnsupervisedList.png",dpi = 300,height = 4.5, width = 14 )


# Now you can save it
# Force everything to be character/text (no lists allowed)
top5_genes_per_pathway_flat <- top5_genes_per_pathway %>%
  mutate(
    Cluster = as.character(Cluster),
    Description = as.character(Description),
    Genes = sapply(Genes, function(x) paste(x, collapse = "; ")),
    Top5Genes = sapply(Top5Genes, function(x) paste(x, collapse = "; "))
  ) %>%
  as.data.frame()

# Now save it
write.csv(top5_genes_per_pathway_flat, file = "Fibroblasts_genes_per_pathway.csv", row.names = FALSE)

```

Gene Ontology Dot Plots for IFE

```{r}

process_go_and_plot("IFE Keratinocytes","royalblue4","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_IFE.csv")
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "IFE")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl3","Il1b",
             "Ubb","Sfn",#Cell Inflamation
             "Egfr","Notch2","Bmpr1a","Fgfr2",
             
             "Mki67", "Barx2","Dlx2",
             "Myc")
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "royalblue4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotIFEGO.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots for HF

```{r}
#DotPlots for Gene Ontology Cell Types

process_go_and_plot("HF Keratinocytes","skyblue4","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_HF.csv")


alldata.1 <- subset(alldata, ident = "HF")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl3","Fcgr2b", "Cxcl14","Ubc", #Inflammation
             "Ext1","Lrig1", "Lgals1","Atf2", #Axonogensis/Muscle
             "Krt27","Dlx3","Smad3", #Hair Follicle
             "Sox9"
             
        )
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "skyblue4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotHFGO.tiff",dpi = 300,height = 3.5, width = 8 )

```

Gene Ontology Dot Plots for Sebocytes

```{r}
#DotPlots for Gene Ontology Cell Types

process_go_and_plot("Sebocytes","turquoise4","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Sebocytes.csv")

alldata.1 <- subset(alldata, ident = "Sebocytes")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl2","Ccl4","Tagln2","S100a11", #Inflammation
             "Cdk14","Tanc2a", "Efna5", #Hair Follicle
             "Lef1","Sox5", "Mef2c", "Esr1","Bmpr2"
        )
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "turquoise4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotSebocytesGO.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots for Schwann Cells

```{r}
#DotPlots for Gene Ontology Cell Types

process_go_and_plot("Schwann cells","darkseagreen3","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Schwann_cells.csv")

alldata.1 <- subset(alldata, ident = "Schwann cells")  
Idents(alldata.1) <- "timepoint"


features = c("Cttn","Cfl1","Tubb2b", #modification of postsynaptic structure
             "Ptprg","Nf1","Ptprz1","Magi2", #Hair Follicle
             "Chl1","Hmox1"
        )
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "darkseagreen3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotSchwannCells.tiff",dpi = 300,height = 3.5, width = 8 )
```



```{r Gene Ontology Dot Plots for Skeletal Muscle}
#DotPlots for Gene Ontology Cell Types

process_go_and_plot("Muscle","orange3","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Muscle_cells.csv")

alldata.1 <- subset(alldata, ident = "Muscle cells")  
Idents(alldata.1) <- "timepoint"


features = c("S100a9","Il1b","Ccl3", #modification of postsynaptic structure
             "Ttn","Mlip3","Mef2c","Sox6", #Hair Follicle
             "Smad4","Lef1","Notch1", "Dab2"
        )
             

p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "orange3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotSkeletalMuscle.tiff",dpi = 300,height = 3.5, width = 8 )

```



```{r Gene Ontology Dot Plots for Endothelial}
#DotPlots for Gene Ontology Cell Types
process_go_and_plot("Endothelial","gold3","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Endothelial_cells.csv")


alldata.1 <- subset(alldata, ident = "Endothelial")  
Idents(alldata.1) <- "timepoint"


features = c("Serpine1","Spp1","Ccl3","Thbs1","Egr1", #modification of postsynaptic structure
             "Ptprm","Cdh13","Vegfc", #Hair Follicle
             "Bmpr1a","Dlx3","Cdh3", "Lncpint")
             

p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "gold3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotEndothelial.tiff",dpi = 300,height = 3.5, width = 8 )

```

Gene Ontology for Pericytes

```{r Gene Ontology Dot Plots for Pericytes}
#DotPlots for Gene Ontology Cell Types
process_go_and_plot("Pericytes","mediumpurple4","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Pericytes.csv")


alldata.1 <- subset(alldata, ident = "Pericytes")  
Idents(alldata.1) <- "timepoint"


features = c("Serpine1","Spp1","Ccl3","Thbs1","Egr1", #modification of postsynaptic structure
             "Ptprm","Cdh13","Vegfc", #Hair Follicle
             "Bmpr1a","Dlx3","Cdh3", "Lncpint")
             

p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "mediumpurple")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotPericytes.tiff",dpi = 300,height = 3.5, width = 8 )
```

```{r Gene Ontology for Melanocytes}
#DotPlots for Gene Ontology Cell Types
process_go_and_plot("Melanocytes","orchid3","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Melanocytes.csv")


alldata.1 <- subset(alldata, ident = "Melanocytes")  
Idents(alldata.1) <- "timepoint"


features = c("Serpine1","Spp1","Ccl3","Thbs1","Egr1", #modification of postsynaptic structure
             "Ptprm","Cdh13","Vegfc", #Hair Follicle
             "Bmpr1a","Dlx3","Cdh3", "Lncpint")
             

p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "orchid3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotMleanocytes.tiff",dpi = 300,height = 3.5, width = 8 )
```

```{r Gene Ontology for Adipocytes}
#DotPlots for Gene Ontology Cell Types

process_go_and_plot("Adipocytes","olivedrab3","/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Gene Ontology/csv files/GO_Adipocytes.csv")


alldata.1 <- subset(alldata, ident = "Adipocytes")  
Idents(alldata.1) <- "timepoint"


p <- DotPlot(alldata.1, unique_top5_genes, dot.scale = 7, cols = c("white", "olivedrab4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotAdipocytesGOEnhanced.png",dpi = 300,height = 4.5, width = 7 )

ggsave("DotPlotAdipocytesGOUnsuprisedList.png",dpi = 300,height = 4.5, width = 14 )


# Now you can save it
# Force everything to be character/text (no lists allowed)
top5_genes_per_pathway_flat <- top5_genes_per_pathway %>%
  mutate(
    Cluster = as.character(Cluster),
    Description = as.character(Description),
    Genes = sapply(Genes, function(x) paste(x, collapse = "; ")),
    Top5Genes = sapply(Top5Genes, function(x) paste(x, collapse = "; "))
  ) %>%
  as.data.frame()

# Now save it
write.csv(top5_genes_per_pathway_flat, file = "Adipocytes_genes_per_pathway.csv", row.names = FALSE)

```
Run Cell Chat on the Global Cell Types

```{r}
##Prep the cellchat database
CellChatDB <- CellChatDB.mouse

showDatabaseCategory(CellChatDB)
#This removes non-protein interactions that are contained in the CellChat DB such as synaptic interactions and metabolic signals
CellChatDB.use <- subsetDB(CellChatDB)

```

```{r}
#Hold off Until RoadMap
cellchat <- createCellChat(object = alldata.1, group.by = "celltype", assay = "RNA")
levels(cellchat@idents)
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multisession", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

```


LGALS Comparing Unwounded vs Wounded Global Expression
```{r}
#LGALS GLOBAL Expression Plots
View(alldata@meta.data)

# Create a new metadata column "Wound_Status"
alldata$Wound_Status <- ifelse(alldata$timepoint == "Unwounded", "Unwounded", "Wounded")

# Check if the new column is added correctly
table(alldata$Wound_Status)


FeaturePlot_scCustom(alldata, "Lgals1", split.by = "Wound_Status", reduction = "umap.rpca", na_cutoff = 1)

ggsave("FeauturePlotLGAS1ExpressionUWOvsWOGlobal.png", width = 9, height = 5)

p = VlnPlot_scCustom(alldata, "Lgals1", split.by = "Wound_Status", pt.size = 0, assay = "RNA", colors_use = c("cornflowerblue","red"))
p <- p & theme(
  text = element_text(family = "Arial", face = "bold", size = 11),
  axis.text = element_text(family = "Arial", face = "bold", size = 11),
  axis.title = element_text(family = "Arial", face = "bold", size = 11),
  legend.text = element_text(family = "Arial", face = "bold", size = 11),
  legend.title = element_text(family = "Arial", face = "bold", size = 11)
)

# Display the plot
p

ggsave("VlnPlotLGAS1ExpressionUWOvsWOGlobal.png", width = 9, height = 5)


#PSU LGALS1 EXPRESSION PLOTS
alldata = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1-2/PSUUnit.rds")


# Create a new metadata column "Wound_Status"
alldata$Wound_Status <- ifelse(alldata$timepoint == "Unwounded", "Unwounded", "Wounded")

# Check if the new column is   correctly
table(alldata$Wound_Status)


FeaturePlot_scCustom(alldata, "Lgals1", split.by = "Wound_Status", reduction = "umap.harmony", na_cutoff = 0.1)

ggsave("FeauturePlotLGAS1ExpressionUWOvsWOPSU.png", width = 9, height = 5)

p = VlnPlot_scCustom(alldata, "Lgals1", split.by = "Wound_Status", pt.size = 0, assay = "RNA", colors_use = c("cornflowerblue","red"))
p <- p & theme(
  text = element_text(family = "Arial", face = "bold", size = 11),
  axis.text = element_text(family = "Arial", face = "bold", size = 11),
  axis.title = element_text(family = "Arial", face = "bold", size = 11),
  legend.text = element_text(family = "Arial", face = "bold", size = 11),
  legend.title = element_text(family = "Arial", face = "bold", size = 11)
)

# Display the plot
p

ggsave("VlnPlotLGAS1ExpressionUWOvsWOPSU.png", width = 9, height = 5)
```


```{r Find Markers Heatmap}

alldata = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Data/OWHA RDS files/IntegratedMultimodal_FineCellTypes_061225.rds")
# Load libraries
library(Seurat)
library(clusterProfiler)
library(org.Mm.eg.db)  # ⬅️ Mouse annotation database
library(ComplexHeatmap)
library(circlize)
library(dplyr)
library(khroma)

# Set identity and assay
Idents(alldata) <- "timepoint"
DefaultAssay(alldata) <- "RNA"
alldata<- JoinLayers(alldata)

# Recode and factor timepoints
alldata@meta.data$timepoint <- recode(alldata@meta.data$timepoint,
  "Unwounded" = "UW",
  "Wounded_D1PW" = "D1PW",
  "Wounded_D2PW" = "D2PW",
  "Wounded_D4PW" = "D4PW",
  "Wounded_D7PW" = "D7PW",
  "Wounded_D15PW" = "D15PW",
  "Wounded_D30PW" = "D30PW"
)
alldata@meta.data$timepoint <- factor(alldata@meta.data$timepoint,
  levels = c("UW", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW"))

# Find markers
cellmarkers_mouse <- FindAllMarkers(alldata, min.pct = 0.30, recorrect_umi = FALSE, only.pos = TRUE)

# Subset DE genes per timepoint
timepoints <- levels(alldata@meta.data$timepoint)
de_genes_list <- lapply(timepoints, function(tp) {
  subset(cellmarkers_mouse, cluster == tp & avg_log2FC > 1)
})
names(de_genes_list) <- timepoints

# Convert to ENTREZ (mouse)
gene_lists_entrez <- lapply(de_genes_list, function(df) {
  bitr(df$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)$ENTREZID
})

# GO enrichment
ck <- compareCluster(geneCluster = gene_lists_entrez, fun = enrichGO, OrgDb = org.Mm.eg.db,
                     pvalueCutoff = 0.05, ont = "BP")
ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")
ck <- simplify(ck)

# Unique pathways
ck_df <- as.data.frame(ck)
unique_df <- ck_df %>% group_by(ID) %>% filter(n_distinct(Cluster) == 1) %>% ungroup()
ck@compareClusterResult <- unique_df


# Dotplot
s1 <- dotplot(ck, showCategory = 8, font.size = 12, label_format = 100) +
  scale_color_gradientn(colors = c("blue", "white", "red")) +
  theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm")) +
  theme(axis.text.x = element_text(size = 12, face = "bold", color = "black", angle = 45, hjust = 1),
        axis.text.y = element_text(size = 12, face = "bold", color = "black"))

s1 + customtheme
ggsave("Mouse_GeneOntology_GlobalPathways_AllTimepoints.png", width = 10, height = 7)

# Save GO results
write.csv(as.data.frame(ck), file = "Mouse_GlobalGOcompareCluster_results.csv", row.names = FALSE)
ck <- read.csv("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 2/Mouse_GlobalGOcompareCluster_results.csv")

library(ggplot2)
library(dplyr)
library(khroma)

# ---- Define Nightfall colors for timepoints ----
timepoint_colors <- color("nightfall")(7)
names(timepoint_colors) <- c("UW", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")

# ---- Ensure Cluster is a factor in correct order ----
timepoint_order <- c("UW", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")
ck_df$Cluster <- factor(ck_df$Cluster, levels = timepoint_order)

# ---- Optional: Keep only unique pathways (appear in 1 cluster) ----
unique_df <- ck_df %>%
  group_by(ID) %>%
  filter(n_distinct(Cluster) == 1) %>%
  ungroup()

# ---- Select top 5 GO terms per cluster by adjusted p-value ----
top_terms <- unique_df %>%
  group_by(Cluster) %>%
  slice_min(order_by = p.adjust, n = 5) %>%
  ungroup()

# ---- Create bar plot ----
ggplot(top_terms, aes(x = zScore, y = reorder(Description, zScore), fill = Cluster)) +
  geom_col() +
  scale_fill_manual(values = timepoint_colors) +
  facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
  theme_minimal(base_size = 14) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 12)
  ) +
  labs(
    x = "-log10(adj. p-value)",
    y = "GO Term",
    fill = "Timepoint",
    title = "Top Enriched GO Terms per Timepoint"
  ) + theme_ggprism_mod()

# ---- Save figure ----
ggsave("Mouse_GO_TopTerms_Barplot_Nightfall.svg", width = 12, height = 12)



# Heatmap section
# Top 30 genes per timepoint
top_genes_list <- lapply(names(de_genes_list), function(tp) {
  top <- head(de_genes_list[[tp]][order(-de_genes_list[[tp]]$avg_log2FC), ], 50)
  transform(top, condition = tp)
})
top.genes <- do.call(rbind, top_genes_list)
mouse_genes <- unique(top.genes$gene)
mouse_genes <- setdiff(mouse_genes, "Gm42418")

# Scale and extract matrix
alldata <- ScaleData(alldata, features = mouse_genes)
mouse_genes <- intersect(mouse_genes, rownames(alldata))
matDE1 <- GetAssayData(alldata, slot = "scale.data")[mouse_genes, ]

# Order meta
meta <- alldata@meta.data[colnames(matDE1), ]
meta$order <- paste(meta$timepoint, meta$metaclusters, sep = "_")
ordered_cells <- rownames(meta[order(meta$timepoint, meta$metaclusters), ])
matDE1 <- matDE1[, ordered_cells, drop = FALSE]

# Annotations
metacluster_colors <- c(
  "IFE Keratinocytes" = "#364B9A",
  "HF Keratinocytes"  = "#56B4E9",
  "Sebocytes"         = "#8DD3C7",
  "Fibroblasts"       = "indianred2",
  "Immune cells"      = "#198F1B",
  "Endothelial cells" = "#F6C800",
  "Pericytes"         = "#D12E82",
  "Muscle cells"            = "orange2",
  "Schwann cells"     = "purple",
  "Melanocytes"       = "pink1",
  "Adipocytes"        = "deeppink",
  "Red Blood Cells"   = "#A7222B"
)
timepoint_colors <- color("nightfall")(7)
names(timepoint_colors) <- c("UW", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")

col_annot <- HeatmapAnnotation(
  df = meta[ordered_cells, c("metaclusters", "timepoint")],
  col = list(
    metaclusters = metacluster_colors,
    timepoint = timepoint_colors
  )
)

col_fun <- colorRamp2(c(-1, 0, 2), c("blue", "white", "red"))

# Save heatmap
png(file = "Mouse_HeatmapByTimepoint-Metacluster_TopGenes.png",
    width = 12, height = 15, units = 'in', bg = "transparent", res = 300)
hm <- Heatmap(
  matDE1,
  name = "logFC\n(Z-score)",
  top_annotation = col_annot,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  row_names_gp = gpar(fontsize = 8, fontface = "bold"),
  row_names_max_width = unit(10, "cm"),
  show_column_names = FALSE,
  show_heatmap_legend = TRUE,
  col = col_fun,
  column_names_rot = 45,
  column_gap = unit(5, 'mm')
)
draw(hm, merge_legend = TRUE)
dev.off()



```
```{r Keratinocyte Specfic Gene Ontology Bar Plots (Differential Expression)}
name = "Keratinocytes"
alldata.1 = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/scRNA SEQ of wounds/Skin Wound Healing Atlas/3. Subclustering/Keratinocyte Subclustering/KeratinocytesAnnotated_2025-07-07.rds")
colors_pals = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1/color_palette.rds")


ife_levels <- c("Ker.Cycling I",
  "Basal II", "Basal III", "Basal IV",
  "Spinous I", "Spinous II", "Spinous III", "Spinous IV",
  "Cornified"
)

alldata.1 = subset(alldata.1, ident = ife_levels)


alldata.1$celltype <- Idents(alldata.1)
# Update timepoint names
alldata.1@meta.data$timepoint <- recode(alldata.1@meta.data$timepoint,
  "Unwounded" = "UW",
  "Wounded_D1PW" = "D1PW",
  "Wounded_D2PW" = "D2PW",
  "Wounded_D4PW" = "D4PW",
  "Wounded_D7PW" = "D7PW",
  "Wounded_D15PW" = "D15PW",
  "Wounded_D30PW" = "D30PW"
)

# ---- Libraries ----
library(Seurat)
library(DElegate)
library(clusterProfiler)
library(org.Mm.eg.db)
library(dplyr)
library(ggplot2)
library(khroma)

# ---- DE + GO on full alldata.1 ----
run_DE_GO_analysis <- function(alldata.1) {
  
  # ---- Run DE for pairwise comparisons ----
  de_results.04 <- DElegate::findDE(
    object = alldata.1,
    compare = c("D4PW", "UW"),
    method = "deseq",
    group_column = "timepoint",
    replicate_column = "orig.ident"
  )
  
  de_results.47 <- DElegate::findDE(
    object = alldata.1,
    compare = c("D7PW", "D4PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  
  de_results.715 <- DElegate::findDE(
    object = alldata.1,
    compare = c("D15PW", "D7PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  
  de_results.1530 <- DElegate::findDE(
    object = alldata.1,
    compare = c("D30PW", "D15PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  
  # ---- Filter DE results ----
  x <- .5     # logFC threshold
  y <- 0.01   # padj threshold
  
  res.04   <- subset(de_results.04, padj < y & log_fc > x)
  res.47   <- subset(de_results.47, padj < y & log_fc > x)
  res.715  <- subset(de_results.715, padj < y & log_fc > x)
  res.1530 <- subset(de_results.1530, padj < y & log_fc > x)
  
  # ---- Map SYMBOL to ENTREZ ----
  ids.04   <- bitr(res.04$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.47   <- bitr(res.47$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.715  <- bitr(res.715$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.1530 <- bitr(res.1530$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  
  genelist <- list(
    D4PW_vs_UW      = ids.04$ENTREZID,
    D7PW_vs_D4PW    = ids.47$ENTREZID,
    D15PW_vs_D7PW   = ids.715$ENTREZID,
    D30PW_vs_D15PW  = ids.1530$ENTREZID
  )
  
  # ---- GO enrichment ----
  ck <- compareCluster(
    geneCluster = genelist,
    fun = enrichGO,
    OrgDb = org.Mm.eg.db,
    pvalueCutoff = 0.05,
    ont = "BP"
  )
  ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")
  
  ck_df <- as.data.frame(ck)
  
  # ---- Keep unique pathways ----
  unique_df <- ck_df %>%
    group_by(ID) %>%
    filter(n_distinct(Cluster) == 1) %>%
    ungroup()
  
  # ---- Select top 5 per comparison ----
  top_terms <- unique_df %>%
    group_by(Cluster) %>%
    slice_min(order_by = p.adjust, n = 8) %>%
    ungroup()
  
  # ---- Order comparisons ----
  comp_order <- c("D4PW_vs_UW", "D7PW_vs_D4PW", "D15PW_vs_D7PW", "D30PW_vs_D15PW")
  top_terms$Cluster <- factor(top_terms$Cluster, levels = comp_order)
  
  # ---- Colors (nightfall palette) ----
  comp_colors <- color("nightfall")(length(comp_order))
  names(comp_colors) <- comp_order
  
  # ---- Plot by Z-score ----
  p <- ggplot(top_terms,
              aes(x = zScore,
                  y = reorder(Description, zScore),
                  fill = Cluster)) +
    geom_col() +
    scale_fill_manual(values = comp_colors) +
    facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
    theme_ggprism_mod(base_size = 14) +
    theme(
      strip.text = element_text(face = "bold", size = 14),
      axis.text.y = element_text(size = 12, face = "bold"),
      axis.text.x = element_text(size = 12),
      legend.position = "none"
    ) +
    labs(
      x = "Z-score",
      y = "GO Term",
      title = ""
    )
  
  # ---- Save outputs ----
  ggsave("GO_TopTerms_Keratinocytes_Barplot_Nightfall.svg",
         p, width = 10, height = 10, dpi = 300)
  write.csv(ck_df, "GO_Keratinocytes_results.csv", row.names = FALSE)
  
  print(p)
  
}

# ---- Run ----
run_DE_GO_analysis(alldata.1)




p <- ggplot(top_terms,
              aes(x = zScore,
                  y = reorder(Description, zScore),
                  fill = Cluster)) +
    geom_col() +
    scale_fill_manual(values = comp_colors) +
    facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
    theme_ggprism_mod(base_size = 14) +
    theme(
      strip.text = element_text(face = "bold", size = 14),
      axis.text.y = element_text(size = 12, face = "bold"),
      axis.text.x = element_text(size = 12),
      legend.position = "none"
    ) +
    labs(
      x = "Z-score",
      y = "GO Term",
      title = "Top Enriched GO Terms across Timepoints (Keratinocytes)"
    )

```


```{r Global Gene Ontology Across Timepoints}

alldata = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/scRNA SEQ of wounds/Skin Wound Healing Atlas/IntegratedMultimodal_FineCellTypes_061225.rds")
colors_pals = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1/color_palette.rds")

alldata$celltype <- Idents(alldata)
# Update timepoint names
alldata@meta.data$timepoint <- recode(alldata@meta.data$timepoint,
  "Unwounded" = "UW",
  "Wounded_D1PW" = "D1PW",
  "Wounded_D2PW" = "D2PW",
  "Wounded_D4PW" = "D4PW",
  "Wounded_D7PW" = "D7PW",
  "Wounded_D15PW" = "D15PW",
  "Wounded_D30PW" = "D30PW"
)

# ---- Libraries ----
library(Seurat)
library(DElegate)
library(clusterProfiler)
library(org.Mm.eg.db)
library(dplyr)
library(ggplot2)
library(khroma)
library(CellChat)

# ---- DE + GO on full alldata ----
run_DE_GO_analysis <- function(alldata) {
  
  # ---- Run DE for pairwise comparisons ----
  de_results.04 <- DElegate::findDE(
    object = alldata,
    compare = c("D4PW", "UW"),
    method = "deseq",
    group_column = "timepoint",
    replicate_column = "orig.ident"
  )
  
  de_results.47 <- DElegate::findDE(
    object = alldata,
    compare = c("D7PW", "D4PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  
  de_results.715 <- DElegate::findDE(
    object = alldata,
    compare = c("D15PW", "D7PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  
  de_results.1530 <- DElegate::findDE(
    object = alldata,
    compare = c("D30PW", "D15PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  
  # ---- Filter DE results ----
  x <- 0     # logFC threshold
  y <- 0.01   # padj threshold
  
  res.04   <- subset(de_results.04, padj < y & log_fc > x)
  res.47   <- subset(de_results.47, padj < y & log_fc > x)
  res.715  <- subset(de_results.715, padj < y & log_fc > x)
  res.1530 <- subset(de_results.1530, padj < y & log_fc > x)
  
  # ---- Map SYMBOL to ENTREZ ----
  ids.04   <- bitr(res.04$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.47   <- bitr(res.47$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.715  <- bitr(res.715$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.1530 <- bitr(res.1530$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  
  genelist <- list(
    D4PW_vs_UW      = ids.04$ENTREZID,
    D7PW_vs_D4PW    = ids.47$ENTREZID,
    D15PW_vs_D7PW   = ids.715$ENTREZID,
    D30PW_vs_D15PW  = ids.1530$ENTREZID
  )
  
  # ---- GO enrichment ----
  ck <- compareCluster(
    geneCluster = genelist,
    fun = enrichGO,
    OrgDb = org.Mm.eg.db,
    pvalueCutoff = 0.05,
    ont = "BP"
  )
  ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")
  
  ck_df <- as.data.frame(ck)
  
  # ---- Keep unique pathways ----
  unique_df <- ck_df %>%
    group_by(ID) %>%
    filter(n_distinct(Cluster) == 2) %>%
    ungroup()
  
  # ---- Select top 5 per comparison ----
  top_terms <- unique_df %>%
    group_by(Cluster) %>%
    slice_min(order_by = p.adjust, n = 8) %>%
    ungroup()
  
  # ---- Order comparisons ----
  comp_order <- c("D4PW_vs_UW", "D7PW_vs_D4PW", "D15PW_vs_D7PW", "D30PW_vs_D15PW")
  top_terms$Cluster <- factor(top_terms$Cluster, levels = comp_order)
  
  # ---- Colors (nightfall palette) ----
  comp_colors <- color("nightfall")(length(comp_order))
  names(comp_colors) <- comp_order
  
  # ---- Plot by Z-score ----
  p <- ggplot(top_terms,
              aes(x = zScore,
                  y = reorder(Description, zScore),
                  fill = Cluster)) +
    geom_col() +
    scale_fill_manual(values = comp_colors) +
    facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
    theme_ggprism_mod(base_size = 14) +
    theme(
      strip.text = element_text(face = "bold", size = 14),
      axis.text.y = element_text(size = 12, face = "bold"),
      axis.text.x = element_text(size = 12),
      legend.position = "none"
    ) +
    labs(
      x = "Z-score",
      y = "GO Term",
      title = ""
    )
  
  # ---- Save outputs ----
  ggsave("GO_TopTerms_Gloabal_Barplot_NightfallNew.svg",
         p, width = 10, height = 10, dpi = 300)
  write.csv(ck_df, "GO_Global_resultsNew.csv", row.names = FALSE)
  
  print(p)
  
}

# ---- Run ----
options(future.globals.maxSize = 8000 * 1024^2)
run_DE_GO_analysis(alldata)




p <- ggplot(top_terms,
              aes(x = zScore,
                  y = reorder(Description, zScore),
                  fill = Cluster)) +
    geom_col() +
    scale_fill_manual(values = comp_colors) +
    facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
    theme_ggprism_mod(base_size = 14) +
    theme(
      strip.text = element_text(face = "bold", size = 14),
      axis.text.y = element_text(size = 12, face = "bold"),
      axis.text.x = element_text(size = 12),
      legend.position = "none"
    ) +
    labs(
      x = "Z-score",
      y = "GO Term",
      title = "Top Enriched GO Terms across Timepoints (Keratinocytes)"
    )
```
```{r}
# ---- Libraries ----
# ========================================
# Differential Expression Heatmap (Grouped by Peak Timepoint)
# ========================================

# ---- Libraries ----
library(Seurat)
library(dplyr)
library(ComplexHeatmap)
library(circlize)
library(khroma)
library(ggplot2)
library(grid)

# ---- Create a named list of DE results ----
de_results_list <- list(
  D4_vs_UW   = res.04,
  D7_vs_D4   = res.47,
  D15_vs_D7  = res.715,
  D30_vs_D15 = res.1530
)

# ---- Define Nightfall colors for timepoints ----
timepoint_colors <- color("nightfall")(7)
names(timepoint_colors) <- c("UW", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")

# ---- Define metacluster colors ----
metacluster_colors <- c(
  "IFE Keratinocytes" = "#364B9A",
  "HF Keratinocytes"  = "#56B4E9",
  "Sebocytes"         = "#8DD3C7",
  "Fibroblasts"       = "#F95B3C",
  "Immune cells"      = "#198F1B",
  "Endothelial cells" = "#F6C800",
  "Pericytes"         = "#D12E82",
  "Muscle cells"      = "#EC6F2D",
  "Schwann cells"     = "#7A4FCF",
  "Melanocytes"       = "#FA5B77",
  "Adipocytes"        = "#19A1D1",
  "Red Blood Cells"   = "#A7222B"
)

# ---- Prepare top genes list, filter by logFC and rate thresholds ----
logFC_min <- -10
logFC_max <- 10

top_genes_list <- lapply(names(de_results_list), function(tp) {
  df <- de_results_list[[tp]] %>%
    filter(
      log_fc > logFC_min & log_fc <= logFC_max,
      (rate1 > 0.15 | rate2 > 0.15)   # rate filter
    ) %>%
    arrange(-log_fc) %>%
    head(200)
  df$timepoint <- tp
  return(df)
})

top_genes_df <- do.call(rbind, top_genes_list)

# Remove unwanted genes
genes_to_plot <- setdiff(unique(top_genes_df$feature), "Gm42418")

# ---- Scale data in Seurat ----
alldata <- ScaleData(alldata, features = genes_to_plot)
genes_to_plot <- intersect(genes_to_plot, rownames(alldata))
mat <- GetAssayData(alldata, slot = "scale.data")[genes_to_plot, ]

# ---- Order cells by timepoint and metacluster ----
meta <- alldata@meta.data[colnames(mat), ]
meta$order <- paste(meta$timepoint, meta$metaclusters, sep = "_")
ordered_cells <- rownames(meta[order(meta$timepoint, meta$metaclusters), ])
mat <- mat[, ordered_cells, drop = FALSE]

# ---- Compute average expression per timepoint ----
time_levels <- c("UW", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")

avg_by_time <- sapply(time_levels, function(tp) {
  cells_tp <- rownames(meta[meta$timepoint == tp, ])
  if (length(cells_tp) > 1) {
    rowMeans(mat[, cells_tp, drop = FALSE])
  } else if (length(cells_tp) == 1) {
    mat[, cells_tp, drop = FALSE]
  } else {
    # handle missing timepoints gracefully
    rep(NA, nrow(mat))
  }
})

# ---- Align row names and remove incomplete rows ----
rownames(avg_by_time) <- rownames(mat)
avg_by_time <- avg_by_time[complete.cases(avg_by_time), , drop = FALSE]

# ---- Determine peak timepoint for each gene ----
gene_peak_tp <- apply(avg_by_time, 1, function(x) names(which.max(x)))
gene_peak_tp <- factor(gene_peak_tp, levels = time_levels)

# ---- Ensure matching order ----
gene_peak_tp <- gene_peak_tp[rownames(avg_by_time)]
mat <- mat[rownames(avg_by_time), , drop = FALSE]

# ---- Reorder genes by peak timepoint and mean expression ----
gene_order <- order(gene_peak_tp, rowMeans(avg_by_time))
mat <- mat[gene_order, , drop = FALSE]
# ---- Column annotations ----
col_annot <- HeatmapAnnotation(
  df = meta[ordered_cells, c("metaclusters", "timepoint")],
  col = list(
    metaclusters = metacluster_colors,
    timepoint = timepoint_colors
  )
)

# ---- Heatmap color scale ----
col_fun <- colorRamp2(c(-1, 0, 2), c("blue", "white", "red"))

# ---- Plot heatmap ----
png(
  filename = "Mouse_Heatmap_TopGenes_GroupedByTimepoint.png",
  width = 12, height = 15, units = "in", res = 300, bg = "transparent"
)

hm <- Heatmap(
  mat,
  name = "Expression\n(Z-score)",
  top_annotation = col_annot,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  row_names_gp = gpar(fontsize = 8, fontface = "bold"),
  row_names_max_width = unit(10, "cm"),
  show_column_names = FALSE,
  show_heatmap_legend = TRUE,
  col = col_fun,
  column_names_rot = 45,
  column_gap = unit(5, "mm")
)

draw(hm, merge_legend = TRUE)
dev.off()

```
```{r}
library(dplyr)
library(ComplexHeatmap)
library(circlize)
library(grid)

# ---- Combine DE results into one matrix ----
de_results_list <- list(
  D4_vs_UW   = de_results.04,
  D7_vs_D4   = de_results.47,
  D15_vs_D7  = de_results.715,
  D30_vs_D15 = de_results.1530
)

# ---- Create gene x timepoint matrix and cap logFC ----
logFC_cap <- 5

all_genes <- unique(unlist(lapply(de_results_list, function(x) x$feature)))

logFC_mat <- sapply(names(de_results_list), function(tp) {
  df <- de_results_list[[tp]]
  vals <- setNames(df$log_fc, df$feature)
  vals[all_genes]
})
logFC_mat[is.na(logFC_mat)] <- 0
logFC_mat <- pmin(pmax(logFC_mat, -logFC_cap), logFC_cap)  # cap logFC at ±5
rownames(logFC_mat) <- all_genes

# ---- Define colors for timepoints ----
timepoint_colors <- color("nightfall")(length(names(de_results_list)))
names(timepoint_colors) <- names(de_results_list)

col_fun <- colorRamp2(c(-logFC_cap, 0, logFC_cap), c("blue", "white", "red"))

# ---- Save heatmap as PNG ----
png(filename = "Mouse_LogFC_Heatmap_Capped.png",
    width = 12, height = 15, units = "in", res = 300, bg = "transparent")

Heatmap(
  logFC_mat,
  name = "logFC",
  cluster_rows = TRUE,          # cluster genes to show temporal patterns
  cluster_columns = FALSE,      # keep timepoints in chronological order
  col = col_fun,
  show_row_names = TRUE,
  show_column_names = TRUE,
  column_names_side = "top",
  row_names_gp = gpar(fontsize = 6),
  clustering_method_rows = "complete"
)

dev.off()
```

```{r}
# 1. Load required libraries
library(Seurat)
library(clusterProfiler)
library(org.Mm.eg.db)
library(dplyr)
library(ggplot2)
library(khroma)

# 2. Prepare Data & Create Grouping
# Group scRNAseq and CITEseq as 'Whole Cell', and snRNAseq as 'Nuclei'
alldata@meta.data <- alldata@meta.data %>%
  mutate(modality_group = case_when(
    modality %in% c("scRNAseq", "CITEseq") ~ "Whole Cell",
    modality == "snRNAseq" ~ "Nuclei",
    TRUE ~ modality
  ))

# Set active identity and factor levels
group_order <- c("Whole Cell", "Nuclei")
alldata@meta.data$modality_group <- factor(alldata@meta.data$modality_group, levels = group_order)
Idents(alldata) <- "modality_group"
DefaultAssay(alldata) <- "RNA"
alldata <- JoinLayers(alldata)

# 3. Find markers for the new groups
cellmarkers_group <- FindAllMarkers(alldata, 
                                    min.pct = 0.30, 
                                    recorrect_umi = FALSE, 
                                    only.pos = TRUE)

# 4. Gene Ontology Analysis
# Convert to ENTREZ IDs
gene_lists_entrez <- lapply(group_order, function(grp) {
  genes <- subset(cellmarkers_group, cluster == grp & avg_log2FC > 1)$gene
  if(length(genes) == 0) return(NULL)
  bitr(genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)$ENTREZID
})
names(gene_lists_entrez) <- group_order
gene_lists_entrez <- gene_lists_entrez[!sapply(gene_lists_entrez, is.null)]

# Run compareCluster
ck <- compareCluster(geneCluster = gene_lists_entrez, 
                    fun = enrichGO, 
                    OrgDb = org.Mm.eg.db, 
                    pvalueCutoff = 0.05, 
                    ont = "BP")

# Clean up results
ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")
ck_df <- as.data.frame(ck)

# 5. Save the CSV results
write.csv(ck_df, file = "GO_Results_WholeCell_vs_Nuclei.csv", row.names = FALSE)

# 6. Filter for Unique Pathways & Top Terms
# Keeping pathways that only appear in one of the two groups
unique_terms <- ck_df %>%
  group_by(ID) %>%
  filter(n_distinct(Cluster) == 1) %>%
  ungroup() %>%
  mutate(logP = -log10(p.adjust))

# 7. Generate Separate Barplots
# Define specific colors from the Nightfall palette
group_colors <- c("Whole Cell" = "#1A1A1A", "Nuclei" = "#5D3A9B")

for (grp in names(gene_lists_entrez)) {
  
  # Select top 10 unique terms for this group
  plot_data <- unique_terms %>%
    filter(Cluster == grp) %>%
    slice_min(order_by = p.adjust, n = 10)
  
  # Create Plot
  p <- ggplot(plot_data, aes(x = logP, y = reorder(Description, logP))) +
    geom_col(fill = group_colors[grp], color = "black", width = 0.7) +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.y = element_text(size = 12, face = "bold", color = "black"),
      axis.text.x = element_text(size = 12, color = "black"),
      panel.grid.minor = element_blank(),
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16)
    ) +
    labs(
      x = "-log10(adj. p-value)",
      y = NULL,
      title = paste("Unique GO Terms:", grp)
    )
  
  # Save plot
  file_name <- paste0("GO_Barplot_", gsub(" ", "_", grp), ".svg")
  ggsave(file_name, plot = p, width = 9, height = 7)
  
  # Print to show in plot window
  print(p)
}
```

```{r}
# 1. Setup Groups
alldata@meta.data <- alldata@meta.data %>%
  mutate(modality_group = case_when(
    modality %in% c("scRNAseq", "CITEseq") ~ "Whole Cell",
    modality == "snRNAseq" ~ "Nuclei",
    TRUE ~ modality
  ))

modality_groups <- c("Whole Cell", "Nuclei")
full_timepoints <- c("UW", "D4PW", "D7PW", "D15PW", "D30PW")
nuclei_timepoints <- c("UW", "D4PW", "D7PW", "D15PW", "D30PW")

# 2. Loop with robust subsetting
for (m_grp in modality_groups) {
  
  message(paste("\n--- Starting Analysis for:", m_grp, "---"))
  
  # A. Subset by Modality Group
  obj_sub <- subset(alldata, modality_group == m_grp)
  
  # B. Filter out any cells that don't belong to the desired timepoints
  target_tps <- if(m_grp == "Nuclei") nuclei_timepoints else full_timepoints
  
  # Ensure we only keep cells that are actually in our target list
  cells_to_keep <- rownames(obj_sub@meta.data[obj_sub@meta.data$timepoint %in% target_tps, ])
  obj_sub <- subset(obj_sub, cells = cells_to_keep)
  
  # C. Fix Factors: This removes 'ghost' levels that have 0 cells
  obj_sub$timepoint <- factor(as.character(obj_sub$timepoint), levels = target_tps)
  Idents(obj_sub) <- "timepoint"
  
  # D. Prepare Layers (Required for FindAllMarkers in Seurat v5)
  obj_sub <- JoinLayers(obj_sub)
  
  # 3. Find Markers
  message("Finding markers...")
  markers <- tryCatch({
    FindAllMarkers(obj_sub, 
                   only.pos = TRUE, 
                   min.pct = 0.25, 
                   logfc.threshold = 0.5) 
  }, error = function(e) {
    message(paste("Error in FindAllMarkers for", m_grp, ":", e))
    return(NULL)
  })
  
  if (is.null(markers) || nrow(markers) == 0) {
    message(paste("No markers found for", m_grp))
    next
  }

  # 4. Filter and Convert to Entrez
  # Only use clusters that actually returned markers
  existing_clusters <- intersect(levels(obj_sub$timepoint), unique(markers$cluster))
  
  gene_list <- lapply(existing_clusters, function(tp) {
    genes <- markers %>% filter(cluster == tp) %>% pull(gene)
    if(length(genes) < 3) return(NULL) 
    
    entrez <- bitr(genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db, drop = TRUE)$ENTREZID
    return(entrez)
  })
  names(gene_list) <- existing_clusters
  gene_list <- gene_list[!sapply(gene_list, is.null)]

  if (length(gene_list) == 0) {
    message("No gene lists found for GO.")
    next
  }

  # 5. GO Enrichment
  message("Running GO Enrichment...")
  ck <- compareCluster(geneCluster = gene_list, 
                       fun = enrichGO, 
                       OrgDb = org.Mm.eg.db, 
                       ont = "BP", 
                       pvalueCutoff = 0.05)
  
  ck <- simplify(setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID"))
  
  # 6. Save CSV
  write.csv(as.data.frame(ck), 
            file = paste0("GO_Timepoint_Results_", gsub(" ", "_", m_grp), ".csv"), 
            row.names = FALSE)
  
  # 7. Visualization
  p <- dotplot(ck, showCategory = 5) + 
    scale_color_gradientn(colors = c("blue", "white", "red")) +
    labs(title = paste("Pathways Over Time:", m_grp), x = "Timepoint") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
          axis.text.y = element_text(size = 9, face = "bold", color = "black"))
  
  ggsave(paste0("GO_Timepoint_Dotplot_", gsub(" ", "_", m_grp), ".svg"), 
         plot = p, width = 10, height = 12)
  
  print(p)
}
```

