---
title: "Figure 2 UMAP Making"
author: "Jonathan Chin Cheong"
date: "2024-10-21"
output: html_document
---

```{r}
set.seed(123)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1-2")
```

```{r pressure, echo=FALSE}
#Load Libraries
library(Seurat)
library(SeuratData)
library(ggplot2)
library(patchwork)
library(dplyr)
library(Rfast2)
library(viridis)
library(scCustomize)
library(harmony)
library(nichenetr)
library(khroma)
```

Load Data Sets

```{r}
alldata = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/scRNA SEQ of wounds/Skin Wound Healing Atlas/IntegratedMultimodal_FineCellTypes_061225.rds")
```

```{r}
library(colorspace)

metacluster_groups <- list(
  "IFE Keratinocytes" = c("Basal II", "Basal III", "Basal IV", "uHF I", "uHF II", 
                          "Spinous I", "Spinous II", "Spinous III", "Spinous IV", "Cornified", 
                          "Ker.Cycling I", "Ker.Cycling II", "Ker.Cycling III"),
  
  "HF Keratinocytes" = c("IRS I", "IRS II", "Medulla", "Cortex I", "Cortex II" ,"Outer Bulge I", 
                         "Outer Bulge II", "Outer Bulge III", "HF I", "HF II"),
  
  "Sebocytes" = c("Transitional Basal Cells", "Sebocyte II", "Sebocyte I", "Sebocyte III"),
  
  "Fibroblasts" = c("Papillary I", "Papillary II", "Papillary III", "Papillary IV",
                    "Reticular I", "Reticular II", "Reticular III", "Reticular IV", "Reticular V",
                    "Fascia", "Myofibroblasts I", "Myofibroblasts II", "Myofibroblasts III",
                    "Lef1+ Fibroblasts", "Tnc+ Fibroblasts", "Myoc+ Fibroblasts", "Lox+ Fibroblasts","Nfkb+ Fibroblasts"),
  
  "Immune cells" = c("Monocytes I", "Monocytes II", "Monocytes III", 
                     "Macrophages I", "Macrophages II", "Macrophages III", "Macrophages IV", 
                     "Cycling Macrophages", "OC-like cells", "moDCs", "cDC1", "cDC2", 
                     "Migratory DCs", "Langerhans cells", "pDCs", "Neutrophils I", 
                     "Neutrophils II", "Neutrophils III", "Basophils", "Mast cells", "NK cells", 
                     "γδT cells", "T cells", "Tregs", "ILCs","B cells"),
  
  "Endothelial cells" = c( "Tacr1+ ECs", "Capillary ECs", "Mature ECs", "LEC I", 
                          "LEC II", "Artery ECs", "Proliferating ECs", "Progenitor ECs", 
                          "EC I", "EC II", "Vein ECs", "PVMs", "Proliferating LECs", "LEC III", "EC III","VWF-hi ECs"
                          ),
  
  "Pericytes" = c("Pericytes", "VSMCs", "Pericyte I", "Pericyte II", "Pericyte III","Pericyte IV"),
  
  "Muscle cells" = c("Muscle Progenitors", "Myonuclei", "Skeletal Muscle"),
  
  "Schwann cells" = c("Myelinating SC I", "Myelinating SC II", 
                      "Repair SC I", "Repair SC II","Repair SC III"),
  
  "Melanocytes" = c("Oca2- Melanocytes", "Oca2-lo Melanocytes","Oca2+ Melanocytes"),
  
  "Adipocytes" = c("Adipocyte I", "Adipocyte II"),
  
  "Red Blood Cells" = c("Red Blood Cells")
)


# Your metacluster colors and subcluster counts
metacluster_colors <- c(
  "IFE Keratinocytes" = "#364B9A",
  "HF Keratinocytes"  = "#56B4E9",
  "Sebocytes"         = "#8DD3C7",
  "Fibroblasts"       = "indianred2",
  "Immune cells"      = "#198F1B",
  "Endothelial cells" = "#F6C800",
  "Pericytes"         = "#D12E82",
  "Muscle cells"            = "orange2",
  "Schwann cells"     = "purple",
  "Melanocytes"       = "pink1",
  "Adipocytes"        = "deeppink",
  "Red Blood Cells"   = "#A7222B"
)



fine_colors <- list()

for (meta in names(metacluster_groups)) {
  num_celltypes <- length(metacluster_groups[[meta]])
  
  # Generate a perceptually distinct palette around the base color
  pal <- colorway::build_palette(
    central_color = metacluster_colors[[meta]],
    n_colors = num_celltypes,
    hue_range = 0.2,  # optional: narrow hue range for better consistency
    val_range = 0.4,  # adjust value range
    sat_range = -0.3  # adjust saturation
  )
  
  # Assign color vector to the names of subclusters in this metacluster
  names(pal$palette) <- metacluster_groups[[meta]]
  fine_colors[[meta]] <- pal$palette
}

# Combine all subcluster color mappings into a single named vector
color_palette <- unlist(fine_colors)
names(color_palette) <- sub("^[^.]+\\.", "", names(color_palette))

color_palette
color_palette["Skeletal Muscle"] <- "#FF6600"
color_palette["Papillary I"]        <- "#A66C8F"
color_palette["Papillary II"]       <- "#B178A8"
color_palette["Papillary III"]      <- "#A04B72"
color_palette["Papillary IV"]       <- "#8C2B5D"
color_palette["Reticular I"]        <- "#BD5B99"
color_palette["Reticular II"]       <- "#D1647A"
color_palette["Reticular III"]      <- "#E07682"
color_palette["Reticular IV"]       <- "#9B3040"  # a vivid dark red
color_palette["Reticular V"]        <- "#6B1E28"  # deeper, more burgundy/maroon
color_palette["Fascia"]              <- "#E37471"
color_palette["Myofibroblasts I"]   <- "#A74337"
color_palette["Myofibroblasts II"]  <- "#B85D4A"
color_palette["Myofibroblasts III"] <- "#FFAC90"
color_palette["Lef1+ Fibroblasts"]  <- "#BC6C45"
color_palette["Tnc+ Fibroblasts"]   <- "#6A4F2A"
color_palette["Myoc+ Fibroblasts"]  <- "#E3A571"
color_palette["Lox+ Fibroblasts"]   <- "#D49F63"
color_palette["Nfkb+ Fibroblasts"]  <- "#A97132"

# Pericytes updates (cooler purples and distinct greens)
color_palette["Pericytes"]       <- "#A7DD0B"
color_palette["VSMCs"]           <- "#A2E813"
color_palette["Pericyte I"]      <- "#8669B8" # deep dark purple, slightly desaturated
color_palette["Pericyte II"]     <- "#5E3490"  # rich medium purple, balanced saturation
color_palette["Pericyte III"]    <- "#3A1A65" # soft medium-light purple, lighter brightness
color_palette["Pericyte IV"]     <- "#B8A3DA"  # pale pastel purple, low saturation for contrast
#3A1A65"

color_palette["Adipocyte I"]   <- "#FF1493"  # Deep Pink  
color_palette["Adipocyte II"]  <- "#FF69B4"  # Hot Pink  
color_palette["Myelinating SC I"]   <- "#9C27B0"  # Bright Purple  
color_palette["Myelinating SC II"]  <- "#BE4FD9"  # Electric Purple  
color_palette["Repair SC I"]        <- "#D18EFF"  # Neon Lavender  
color_palette["Repair SC II"]       <- "#E6B3FF"  # Soft Bright Lavender  
color_palette["Repair SC III"]      <- "#A63FCC"  # Vivid Purple  

Idents(alldata) <- "celltype_fine"
# 5. Now, use this palette in your plots
DimPlot(alldata, cols = color_palette, reduction = "umap.rpca") + custom_theme
ggsave("UMAPwithLedgend.png", height = 11, width = 15)
DimPlot(alldata, cols = color_palette, reduction = "umap.rpca", raster =F) & NoLegend() +custom_theme
ggsave("UMAPGLoabalBlank.png", height = 11, width = 11)

DimPlot(alldata, cols = color_palette, reduction = "umap.rpca", label = T, repel = T, label.size = 3, raster = F) & NoLegend() &custom_theme
ggsave("UMAPwithOutLedgendWithLabels.png", height = 11, width = 11)

saveRDS(color_palette, file = "color_palette.rds")

alldata <- subset(alldata, subset = orig.ident %in% c("UW_Han", "UW_Joost_1", "UW_Joost_2", "UW_Joost_3"), invert = T)


color_palette <- readRDS("color_palette.rds")
```

MakeUMAPS for Figure 1

```{r}

Idents(alldata) <- alldata$celltype_fine
DimPlot(alldata, cols = color_palette, reduction = "umap.rpca", raster =F) & NoLegend() + NoAxes() + customtheme
ggsave("UMAPofFineCellTypeAnnotationsNoLabels.svg", width = 6, height = 6)
ggsave("UMAPofFineCellTypeAnnotationsNoLabels.png", width = 6, height = 6)


DimPlot(alldata, cols = color_palette, reduction = "umap.rpca", raster =F, split.by = "condition") & NoLegend() + NoAxes() + customtheme
ggsave("UMAPofFineCellTypeSplitbyCondition.svg", width = 12, height = 6)
ggsave("UMAPofFineCellTypeSplitbyCondition.png", width = 12, height = 6)

alldata@meta.data$condition <- ifelse(
  alldata@meta.data$timepoint == "UW", 
  "Unwounded", 
  "Wounded"
)



Idents(alldata) <- alldata$metaclusters

metacluster_colors <- c(
  "IFE Keratinocytes" = "#364B9A",
  "HF Keratinocytes"  = "#56B4E9",
  "Sebocytes"         = "#8DD3C7",
  "Fibroblasts"       = "indianred2",
  "Immune cells"      = "#198F1B",
  "Endothelial cells" = "#F6C800",
  "Pericytes"         = "#D12E82",
  "Muscle cells"            = "orange2",
  "Schwann cells"     = "purple",
  "Melanocytes"       = "pink1",
  "Adipocytes"        = "deeppink",
  "Red Blood Cells"   = "#A7222B"
)


levels(alldata$timepoint) <- c("UWO", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")

DimPlot(alldata, cols = color_palette, reduction = "umap.rpca", raster =F, split.by = "condition") & NoLegend() + NoAxes() + customtheme
ggsave("UMAPofFineCellTypeSplitbyConditionNoLegend.png", width = 12, height = 6)


```


```{r}
# Set target number of cells per condition
target_n <- 43686

# Downsample cells per condition
set.seed(1234)  # For reproducibility

downsampled_cells <- alldata@meta.data %>%
  dplyr::mutate(cell_id = rownames(.)) %>%
  dplyr::group_by(condition) %>%
  dplyr::slice_sample(n = target_n) %>%
  dplyr::pull(cell_id)

# Subset Seurat object
alldata_ds <- subset(alldata, cells = downsampled_cells)

# Plot
p <- DimPlot(
  alldata_ds,
  cols = color_palette,
  reduction = "umap.rpca",
  raster = FALSE,
  split.by = "condition"
) & NoLegend() + NoAxes() + customtheme

# Save
ggsave("UMAPofFineCellTypeSplitbyConditionDownSampledNoLegend.png", plot = p, width = 12, height = 6)
```





```{r}
VlnPlot(alldata, "percent_mito", group.by = modality)
```

```{r}
library(dplyr)
#Add new modality group 
alldata$modality_group <- ifelse(alldata$modality %in% c("scRNAseq", "CITEseq"), "whole cell", "nuclei")

# Add a new metadata column called 'condition' based on 'timepoint'
alldata@meta.data <- alldata@meta.data %>%
  mutate(condition = case_when(
    timepoint == "UW" ~ "Unwounded",
    timepoint %in% c("D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW") ~ "Wounded",
    TRUE ~ as.character(timepoint)  # Keeps any other potential values unchanged
  ))
```

```{r}
cols = c("springgreen2", "orange2", "steelblue2")

p <- VlnPlot(alldata, 
             features = c("nFeature_RNA", "percent_mito", "percent_ribo", "TF Percentage"), 
             cols = cols, 
             group.by = "modality", 
             pt.size = 0)

# Customize the font
p <- p & theme(
  text = element_text(family = "Arial", face = "bold", size = 11),
  axis.text = element_text(family = "Arial", face = "bold", size = 11),
  axis.title = element_text(family = "Arial", face = "bold", size = 11),
  legend.text = element_text(family = "Arial", face = "bold", size = 11),
  legend.title = element_text(family = "Arial", face = "bold", size = 11)
)

# Display the plot
p

ggsave("QualityControlofModalities.png", height = 6, width = 7)


DimPlot(alldata, group.by = "orig.ident", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByBatch.png", height = 6, width = 7)

DimPlot(alldata, group.by = "timepoint", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByTimepoint.png", height = 6, width = 7)

DimPlot(alldata, group.by = "modality", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByModality.png", height = 6, width = 7)

DimPlot(alldata, group.by = "chemistry", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupByChemistry.png", height = 6, width = 7)

DimPlot(alldata, group.by = "sex", reduction = "umap.rpca", raster = F) + NoAxes()
ggsave("RPCAUMAP_GroupBySex.png", height = 6, width = 7)


DimPlot(alldata, group.by = "orig.ident", reduction = "umap.rpca", raster = F,shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByBatchNoLegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "timepoint", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByTimepointNolegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "modality", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByModalityNoLegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "chemistry", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() + NoLegend()
ggsave("RPCAUMAP_GroupByChemistryNoLegend.png", height = 6, width = 6)

DimPlot(alldata, group.by = "sex", reduction = "umap.rpca", raster = F, shuffle = T) + NoAxes() +  NoLegend()
ggsave("RPCAUMAP_GroupBySexNoLegend.png", height = 6, width = 6)



```

Making Graphs for CSCI WIP Progress

```{r}
name = "AllCells"
cols <- metacluster_colors


alldata$celltype <- Idents(alldata)


cells <- as.data.frame(table(alldata$metaclusters,alldata$timepoint))

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4


ggsave(filename = paste0(name, "Subset_RatioOvertimeVer2.tiff"))

alldata.1 = SplitObject(alldata, split.by = "modality")
cells <- as.data.frame(table(alldata.1$scRNAseq$metaclusters,alldata.1$scRNAseq$timepoint))

cells <- cells %>% filter(Freq != 0)

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4

ggsave("scRANSeqMetaclusterBarPlotOverTimepoint.png")


cells <- as.data.frame(table(alldata.1$snRNAseq$metaclusters,alldata.1$snRNAseq$timepoint))
cells <- cells %>% filter(Freq != 0)

g5 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g5 <- g5 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g5

ggsave("snRNAseqMetaclusterBarPlotOverTimepoint.png")


cells <- as.data.frame(table(alldata.1$CITEseq$metaclusters,alldata.1$CITEseq$timepoint))
cells <- cells %>% filter(Freq != 0)

g6 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g6 <- g6 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g6

ggsave("CITEseqMetaclusterBarPlotOverTimepoint.png")

g4 <- g4 + NoLegend()
g5 <- g5 + NoLegend()
g6 <- g6 + NoLegend()
g4+g6+g5

ggsave("ModalitiesMetaclusterBarPlotOverTimepoint.png", width = 10, height = 5)

alldata$celltype <- Idents(alldata)


cells <- as.data.frame(table(alldata$metaclusters,alldata$modality))

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4


ggsave(filename = paste0(name, "Subset_RatioOverModalityVer2.tiff"))



cells <- as.data.frame(table(alldata$metaclusters,alldata$condition))

g4 <- ggplot(cells, aes(fill = Var1, x = Var2, y = Freq)) +
  geom_bar(position = 'fill', stat = "identity", width = 0.5, color = "black") +
  scale_fill_manual(values = cols) +
  ggtitle('Cell Type Representation by Count Ratio') +
  xlab('') +
  ylab('% Total Cells') +
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
g4 <- g4 + theme(
    text = element_text(family = "Arial", face = "bold", size = 12),
    axis.text = element_text(family = "Arial", face = "bold", size = 14, color = "black"),
    axis.title = element_text(family = "Arial", face = "bold", size = 12),
    legend.text = element_text(family = "Arial", face = "bold", size = 12),
    legend.title = element_text(family = "Arial", face = "bold", size = 12))
g4


ggsave(filename = paste0(name, "Subset_RatioOverConditionsyVer2.tiff"))

```

LineGraphs

```{r}
library(ggpattern)
metadata <- alldata@meta.data

 cols <- metacluster_colors
result_df <- metadata %>%
  group_by(orig.ident, modality, metaclusters) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident) %>%
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(modality, metaclusters) %>%
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

ggplot(result_df, aes(x = metaclusters, y = AverageCount, fill = metaclusters)) +
  geom_col_pattern(
    aes(fill = metaclusters, pattern_density = modality),
    colour          = 'black', 
    size            = 0.5,  
    position        = "stack",
    pattern_fill    = 'white',
    pattern_colour  = NA
  ) +
  theme_bw() +
  scale_fill_manual(values = cols, guide = FALSE) +
  labs(
    x = "Cell Type",
    y = "Average Number of Cells",
    fill = "Cell Type"
  )  + 
  coord_flip() +
  scale_pattern_manual(values = c(snRNAseq = "none", scRNAseq = "stripe", CITEseq = "circle")) +  # Specify patterns for modalities
  scale_pattern_density_manual(values = c(snRNAseq = 0, scRNAseq = 0.5, CITEseq = 1)) +
  theme(axis.text.y = element_text(size = 12), panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# Save the plot
ggsave("Global_ModalityCelltype_Averages.png", width = 7, height = 7)

cols <- c("springgreen2", "orange2", "steelblue2")

ggplot(result_df, aes(x = metaclusters, y = AverageProportion, fill = modality)) + 
  geom_bar(
    stat = "identity",
    position = "fill",
    colour = 'black',
    size = 0.5
  ) + 
  theme_bw() + 
  labs(
    x = "Cell Type",
    y = "Average Proportion",
    fill = "Modality"
  ) + 
  coord_flip() + 
  theme(
    axis.text.y = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) + 
  scale_y_continuous(labels = scales::percent_format()) + 
  scale_fill_manual(values = cols)
# Save the plot
ggsave("Global_ModalityCMetaclusters_Proportions.png", width = 7, height = 7)
```

Repeat above plot, faceted by timepoint

```{r}
result_df <- metadata %>%
  group_by(orig.ident, timepoint, modality, metaclusters) %>%   # Include timepoint here
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident, timepoint) %>%                      # Also include timepoint here
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(modality, timepoint, metaclusters) %>%              # Include timepoint in the final grouping
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

ggplot(result_df, aes(x = metaclusters, y = AverageCount, fill = metaclusters)) +
  geom_col_pattern(
    aes(fill = metaclusters, pattern_density = modality),
    colour          = 'black', 
    size            = 0.5,  
    position        = "stack",
    pattern_fill    = 'white',
    pattern_colour  = NA
  ) +
  theme_bw() +
  scale_fill_manual(values = cols, guide = FALSE) +
  labs(
    x = "Cell Type",
    y = "Average Number of Cells",
    fill = "Cell Type"
  )  + 
  coord_flip() +
  facet_wrap(~timepoint, ncol = 3) +
  scale_pattern_manual(values = c(snRNAseq = "none", scRNAseq = "stripe", CITEseq = "circle")) +  # Specify patterns for modalities
  scale_pattern_density_manual(values = c(snRNAseq = 0, scRNAseq = 0.5, CITEseq = 1)) +
  theme(axis.text.y = element_text(size = 12), panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# Save the plot
ggsave("Global_ModalityMetaclusters_Averages_TimepointFacet.png", width = 16, height = 14)

```

```{r}
# Extract meta data information from unaltered data (alldata) and data downsampled to size of smallest batch (alldata_downsample)
min.cells <- min(table(alldata$orig.ident))
#Set identities to the meta data that we want to downsample on
Idents(alldata) <- "orig.ident"
#Subset the data, specifying that we want to downsample to the smallest batch size
alldata_downsample <- subset(alldata, downsample = min.cells)
#View the sizes of all the downsampled batches
table(alldata_downsample$orig.ident)
```

```{r}
# Extract meta data information from unaltered data (alldata) and data downsampled to size of smallest batch (alldata_downsample)
metadata <- alldata@meta.data
metadata_downsample <- alldata_downsample@meta.data

# Set color palette
pal <- metacluster_colors

# Calculate counts, average counts, count SD, proportions, average proportions, and proportion SD in one pipeline for NOT downsampled data
result_df <- metadata %>%
  group_by(orig.ident, timepoint, metaclusters) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident) %>%
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(timepoint,metaclusters) %>%
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

# Calculate counts, average counts, count SD, proportions, average proportions, and proportion SD in one pipeline for downsampled data
result_df_down <- metadata_downsample %>%
  group_by(orig.ident, timepoint, metaclusters) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(orig.ident) %>%
  mutate(TotalCount = sum(Count),
         Proportion = Count / TotalCount) %>%
  ungroup() %>%
  group_by(timepoint, metaclusters) %>%
  summarise(AverageCount = mean(Count, na.rm = TRUE),
            Count_SD = sd(Count, na.rm = TRUE),
            AverageProportion = mean(Proportion, na.rm = TRUE),
            Proportion_SD = sd(Proportion, na.rm = TRUE), 
            Percentage = AverageProportion * 100,
            SD_Percentage = Proportion_SD * 100, .groups = "drop")

# Define the correct order of timepoints
timepoint_levels <- c("UWO", "D1PW", "D2PW", "D4PW", "D7PW", "D15PW", "D30PW")

# Ensure that timepoint columns are factors with the correct levels
result_df$timepoint <- factor(result_df$timepoint, levels = timepoint_levels)
result_df_down$timepoint <- factor(result_df_down$timepoint, levels = timepoint_levels)

# Create ggplot for average cell type counts with error bars (downsampled data)
ggplot(result_df, aes(x = timepoint, y = AverageCount, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = AverageCount - Count_SD, ymax = AverageCount + Count_SD, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  #geom_errorbar(aes(ymin = AverageCount - SD, ymax = AverageCount + SD), width = 0.2) + #Can include error bars if desired
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  )  +
  labs(title = "Average Cell Type Counts Over Time",
       x = "Timepoint",
       y = "Average Cell Count") +
  NoLegend()
ggsave("Global_AverageCelltypeCount.png", width = 9, height = 5.5)

# Create ggplot for average cell type counts with error ribbon (downsampled data)
ggplot(result_df_down, aes(x = timepoint, y = AverageCount, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = AverageCount - Count_SD, ymax = AverageCount + Count_SD, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  #geom_errorbar(aes(ymin = AverageCount - SD, ymax = AverageCount + SD), width = 0.2) + #Can include error bars if desired
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  )  +
  labs(title = "Average Downsampled Cell Type Counts Over Time",
       x = "Timepoint",
       y = "Average Cell Count") +
  NoLegend()
ggsave("Global_AverageCelltypeCount_Downsampled.png", width = 9, height = 5.5)

#Create ggplot for proportions over time
ggplot(result_df, aes(x = timepoint, y = Percentage, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = Percentage - SD_Percentage, ymax = Percentage + SD_Percentage, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  ) +
  labs(title = "Average Cell Type Proportions Over Time",
       x = "Timepoint",
       y = "Average Proportion (% of Total)") +
  NoLegend()
ggsave("Global_AverageCelltypeProp.png", width = 9, height = 5.5)

#Create ggplot for proportions over time using downsampled data
ggplot(result_df_down, aes(x = timepoint, y = Percentage, color = metaclusters, group = metaclusters)) +
  geom_ribbon(aes(ymin = Percentage - SD_Percentage, ymax = Percentage + SD_Percentage, fill = metaclusters), alpha = 0.2) +
  geom_line() +
  geom_point() +
  facet_wrap(~ metaclusters, scales = "free_y") +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11, family = "Arial", face = "bold"),
    axis.text.y = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.x = element_text(size = 11, family = "Arial", face = "bold"),
    axis.title.y = element_text(size = 11, family = "Arial", face = "bold"),
    strip.text = element_text(size = 11, family = "Arial", face = "bold"),
    plot.title = element_text(size = 11, family = "Arial", face = "bold", hjust = 0.5)
  ) +
  labs(
    title = "Averaged Downsampled Cell Type Proportions Over Time",
    x = "Timepoint",
    y = "Average Proportion (% of Total)"
  ) +
  NoLegend()

ggsave("Global_AverageCelltypeProp_Downsampled.png", width = 9, height = 5.5)
#Save the dataframe as an xlsx file
writexl::write_xlsx(list(result_df, result_df_down), "Global_CellCountsProps.xlsx")



# Extract baseline UWO percentages per metacluster
baseline <- result_df %>%
  filter(timepoint == "UWO") %>%
  select(metaclusters, UWO_Percentage = Percentage)

# Join to result_df and calculate fold change vs. UWO
result_df <- result_df %>%
  left_join(baseline, by = "metaclusters") %>%
  mutate(RelativeToUWO = Percentage / UWO_Percentage)

# Plot relative proportion (UWO = 1)
ggplot(result_df, aes(x = timepoint, y = RelativeToUWO, color = metaclusters, group = metaclusters)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(values = pal) +
  labs(
    title = "Cell Type Proportions Relative to UWO (Baseline = 1)",
    y = "Relative Proportion",
    x = "Timepoint"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    legend.title = element_blank(),
    panel.grid.major = element_line(color = "grey50"),  # Show major grid lines
    panel.grid.minor = element_line(color = "grey50")   # Show minor grid lines
  )
ggsave("GlobalLineGraphOverTimeForProportionComparedToUWO.png", height =5, width = 10)
```

```{r}
alldata <- BuildClusterTree(alldata, reduction = "pca")
PlotClusterTree(alldata, direction = "rightwards")
tiff("GlobalSubset_PhyloTree.tiff")
dev.off()
```

```{r Runnning Differntial Expression Across Modalities}
# 
# set.seed()
# levels(alldata$orig.ident)
# pal <- c("royalblue2", "turquoise2", "deepskyblue2", "red2", "springgreen2", "gold2", "mediumpurple2", "orange2", "darkseagreen2", "orchid2", "magenta2", "tomato3")
# 
# # Define the specific orig.ident values to keep
# selected_orig_idents <- c("UW_Haensel1", "UW_Haensel2", "UW_Vu_v3","UW_Joost_1","D1PW_Justynski",
#                           "D2PW_Justynski","D4PW_Haensel1", "D4PW_Haensel2", "D4PW_Haensel3", "D4PW_Vu_v3","D7PW_Vu_v2_1", "D7PW_Vu_v2_2", "D7PW_Vu_v3")
# 
# # Subset the Seurat object using the specified orig.ident values
# alldata_subset <- subset(alldata, subset = orig.ident %in% selected_orig_idents)
# 
# alldata_subset <- IntegrateLayers(
#   object = alldata_subset, 
#   method = RPCAIntegration, 
#   normalization.method = "SCT", 
#   orig.reduction = "pca", 
#   new.reduction = "rpca",
#   verbose = F)
# 
# pcs = npcs(alldata_subset, var.total = 0.9)
#  
# alldata_subset <- FindNeighbors(alldata_subset, reduction = "rpca", dims = 1:pcs) %>% 
#   FindClusters(resolution = 1, cluster.name = "rpca_clusters") %>%
#   RunUMAP(reduction = "rpca", dims = 1:pcs, reduction.name = "umap.rpca")
# 
# Idents(alldata_subset) = alldata_subset$metaclusters
# DimPlot(alldata_subset, cols = pal, reduction = "umap.rpca", pt.size = .3, label = F) + NoLegend() + NoAxes() 
# ggsave("UMAPPublicAvaibleData.tiff")
# 
# 
# #Subset Singel Nuceli Data
# # Subset the Seurat object using the specified orig.ident values
# alldata_subset <- subset(alldata, subset = modality %in% "snRNAseq")
# 
# alldata_subset <- IntegrateLayers(
#   object = alldata_subset, 
#   method = RPCAIntegration, 
#   normalization.method = "SCT", 
#   orig.reduction = "pca", 
#   new.reduction = "rpca",
#   verbose = F)
# 
# pcs = npcs(alldata_subset, var.total = 0.9)
#  
# alldata_subset <- FindNeighbors(alldata_subset, reduction = "rpca", dims = 1:pcs) %>% 
#   FindClusters(resolution = 1, cluster.name = "rpca_clusters") %>%
#   RunUMAP(reduction = "rpca", dims = 1:pcs, reduction.name = "umap.rpca")
# 
# Idents(alldata_subset) = alldata_subset$metaclusters
# DimPlot(alldata_subset, cols = pal, reduction = "umap.rpca", pt.size = .3, label = F) + NoLegend() + NoAxes() 
# ggsave("UMAPSingleNuceliData.tiff")
# 
# # Subset the CITEseq data and make individual UMAP
# alldata_subset <- subset(alldata, subset = modality %in% "CITEseq")
# 
# alldata_subset <- IntegrateLayers(
#   object = alldata_subset, 
#   method = RPCAIntegration, 
#   normalization.method = "SCT", 
#   orig.reduction = "pca", 
#   new.reduction = "rpca",
#   verbose = F)
# 
# pcs = npcs(alldata_subset, var.total = 0.9)
#  
# alldata_subset <- FindNeighbors(alldata_subset, reduction = "rpca", dims = 1:pcs) %>% 
#   FindClusters(resolution = 1, cluster.name = "rpca_clusters") %>%
#   RunUMAP(reduction = "rpca", dims = 1:pcs, reduction.name = "umap.rpca")
# 
# Idents(alldata_subset) = alldata_subset$metaclusters
# DimPlot(alldata_subset, cols = pal, reduction = "umap.rpca", pt.size = .3, label = F) + NoLegend() + NoAxes() 
# ggsave("UMAPCITEseqData.tiff")


```

Differential Expression

```{r}

library(clusterProfiler)
library(org.Mm.eg.db)

# Define cell types to iterate over
cell_types <- c("IFE", "HF", "Immune cells", "Sebocytes", "Fibroblasts", 
                "Schwann cells", "Endothelial cells", "Pericytes", 
                "Muscle cells", "Melanocytes", "RBCs", "Adipocytes")

# Iterate over each cell type
for (cell_type in cell_types) {
  # Subset data based on the specified cell type
  Idents(alldata) <- "metaclusters"
  alldata.1 <- subset(alldata, ident = c(cell_type))
  
  # Perform DE analysis for different comparisons
  de_results.04 <- DElegate::findDE(
    object = alldata.1, 
    compare = c("Wounded_D4PW", "Unwounded"),
    method = "deseq",
    group_column = "timepoint",
    replicate_column = "orig.ident"
  )
  de_results.47 <- DElegate::findDE(
    object = alldata.1, 
    compare = c("Wounded_D7PW", "Wounded_D4PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  de_results.715 <- DElegate::findDE(
    object = alldata.1, 
    compare = c("Wounded_D15PW", "Wounded_D7PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  de_results.1530 <- DElegate::findDE(
    object = alldata.1, 
    compare = c("Wounded_D30PW", "Wounded_D15PW"),
    method = "deseq",
    group_column = "timepoint"
  )
  
  # Set filtering thresholds
  x <- 1
  y <- 0.01
  
  # Subset DE results based on thresholds
  res.04 <- subset(de_results.04, padj < y & log_fc > x)
  res.47 <- subset(de_results.47, padj < y & log_fc > x)
  res.715 <- subset(de_results.715, padj < y & log_fc > x)
  res.1530 <- subset(de_results.1530, padj < y & log_fc > x)
  
  # Convert SYMBOL to ENTREZID
  ids.04 <- bitr(res.04$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.47 <- bitr(res.47$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.715 <- bitr(res.715$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  ids.1530 <- bitr(res.1530$feature, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Mm.eg.db)
  
  # Create a gene list
  genelist <- list(
    UWOvsD4PW = ids.04$ENTREZID,
    D4PWvsD7PW = ids.47$ENTREZID,
    D7PWvsD15PW = ids.715$ENTREZID,
    D15PWvsD30PW = ids.1530$ENTREZID
  )
  
  # Run ClusterProfiler
  ck <- compareCluster(geneCluster = genelist, fun = enrichGO, OrgDb = org.Mm.eg.db, 
                       pvalueCutoff = 0.05, ont = "BP")
  ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType = "ENTREZID")
  
  # Plot dotplot
  s1 <- dotplot(ck, showCategory = 5, font.size = 12, label_format = 100) + 
    theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm"))
  s2 <- s1 + 
    theme(axis.text.x = element_text(size = 12, face = "bold", color = "black", angle = 45)) + 
    theme(axis.text.y = element_text(size = 12, face = "bold", color = "black"))
  print(s2)
  
  # Save plot
  ggsave(paste0("GO_", gsub(" ", "_", cell_type), ".tiff"), width = 10, height = 5)
  
  # Save cluster summary to a CSV file
  cluster_summary <- data.frame(ck)
  write.csv(cluster_summary, file = paste0("GO_", gsub(" ", "_", cell_type), ".csv"), row.names = FALSE)
}


```











Running GO using FindMarkers for Each Timepoint

```{r}

#Differntial Expression
library(clusterProfiler)
library(org.Mm.eg.db)
#Perform Gene Ontology Anaylsis for Each Individual Timepoint Upregulated using Find Markers
Idents(alldata.1) <- "timepoint"
DefaultAssay(alldata.1) <- "RNA"

#Find markers for celltype
cellmarkers <- FindAllMarkers(alldata.1, min.pct = 0.5, recorrect_umi = F, only.pos = T)


res.uwo = subset(cellmarkers, cluster == "Unwounded", avg_log2FC > 1)
res.wo4 = subset(cellmarkers, cluster == "Wounded_D4PW", avg_log2FC > 1)
res.wo7 = subset(cellmarkers, cluster == "Wounded_D7PW", avg_log2FC > 1)
res.wo15 = subset(cellmarkers, cluster == "Wounded_D15PW", avg_log2FC > 1)
res.wo30 = subset(cellmarkers, cluster == "Wounded_D30PW", avg_log2FC > 1)


ids.uwo<-bitr(res.uwo$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo4<-bitr(res.wo4$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo7<-bitr(res.wo7$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo15<-bitr(res.wo15$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)
ids.wo30<-bitr(res.wo30$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb=org.Mm.eg.db)

genelist = list(UWO = ids.uwo$ENTREZID, 
           D4PW = ids.wo4$ENTREZID,
           D7PW = ids.wo7$ENTREZID,
           D15PW = ids.wo15$ENTREZID,
           D30PW = ids.wo30$ENTREZID)
           

# Run ClusterProifler

ck <- compareCluster(geneCluster = genelist, fun = enrichGO, OrgDb = org.Mm.eg.db, pvalueCutoff = 0.05, ont = "BP")

ck <- setReadable(ck, OrgDb = org.Mm.eg.db, keyType="ENTREZID")
dotplot(ck) 
s1 = dotplot(ck, showCategory = 5, font.size = 18, label_format = 100)
s1 = s1 + theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
s2 = s1 + theme(axis.text.x = element_text(size=18, face="bold", color = "black"))
s2 = s2 + theme(axis.text.y = element_text(size=18, face="bold", color = "black"))
s2
```

```{r}
#Load PSU Dataset
alldata.1 = readRDS("/Users/jchincheong/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1-2/PSUUnit.rds")
#Change identity levels so that they are grouped without cell cluster
Idents(alldata.1) <- gsub("\\s+[IVX]+$", "", Idents(alldata))
Idents(alldata.1) <- gsub("\\s+[0-9]+$", "", Idents(alldata))
levels(alldata.1)
```

Run GSEA analysis

```{r}
library(clusterProfiler)
library(DElegate)
library(org.Mm.eg.db)


celltype <- "PSU"
#In case the celltype column in the metadata has a different name 
identity_metadata <- "celltype"


alldata.1$celltype <- Idents(alldata.1)

cellmarkers_bulk <- findDE(alldata.1, group_column = identity_metadata, replicate_column = "orig.ident", method = "deseq", lfc_shrinkage = "apeglm")


#Order the results by decreasing log2FC
cellmarkers_bulk <- cellmarkers_bulk %>%
  group_by(group1) %>%
  arrange(desc(log_fc)) %>%
  ungroup()

gse_bulk <- compareCluster(feature|log_fc~group1, data = cellmarkers_bulk, fun = gseGO, OrgDb = org.Mm.eg.db,
                            keyType ="SYMBOL", pvalueCutoff = 0.05, pAdjustMethod = "BH", ont = "BP", nPermSimple = 10000, eps = 0)



dotplot(gse_bulk) 



s1 <- dotplot(gse_bulk, showCategory =3, font.size = 8, label_format = 100) +
  theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm"))

s2 <- s1 +
  theme(axis.text.y  = element_text(size = 11, face = "bold", color = "black")) +
  theme(axis.text.x = element_text(size = 11, face = "bold", color = "black", angle = 45))
s2
ggsave("DotPlotPSUGroupedByCluster.tiff",height = 8, width = 10)

cluster_summary <- data.frame(gse)
write.csv(cluster_summary, file = paste0("GO_PSUDELEGATE.csv"), row.names = FALSE)
```

Figure 1

```{r}

Idents(alldata) <- "metaclusters"
features = c("Krt14","Itga6", #Keratincoytes Labels
             "Krt15","Sox9",
             "Pparg", #Sebocytes
             "Col1a2","Pdgfra", #Fibroblast Markers
             "Ptprc", # Immune Cell Markers
             "Pecam1", #Endothelial Cell Markers
                          "Rgs5", #Pericytes
             "Mylpf", "Des", # Muscle Markers
                          "Mpz", #Schwann Cell Markers
             "Mlana", #Melanocyte Marker
             "Adipoq", #Adipocyte Markers
             "Hba-a1","Hbb-bt" #RBC marker
)



DotPlot(alldata, features, dot.scale = 7) + RotatedAxis()



p <- DotPlot(alldata, features, dot.scale = 7, cols = "RdYlBu") + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") # Outlines circles in black

# Display the plot
p

ggsave("DotPlotAllModalityIntergrationAcrossCellTypesNew.png",dpi = 300,height = 6, width =7.3 )

```

```{r}
#DotPlots for Gene Ontology Cell Types
Idents(alldata.1) <- "timepoint"


features = c("Cxcl5","Cxcr2","Acp5","Il23a", #D4PW Leukocote Recruitment, Tissue Remodeling
             "Pparg","Gpc6","Ctnna3") #D7PW Proliferation
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "springgreen4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotImmuneCellsGO.tiff",dpi = 300,height = 4.5, width = 7 )


```

Gene Ontology Dot Plots for Immune Cells

```{r}
#DotPlots for Gene Ontology Cell Types

  
alldata.1 <- subset(alldata, ident = "Immune cells")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl5","Cxcr2","Acp5","Il23a", #D4PW Leukocote Recruitment, Tissue Remodeling
             "Pparg","Gpc6","Ctnna3") #D7PW Proliferation
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "springgreen4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotImmuneCellsGO2.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots For Fibroblasts

```{r}
#DotPlots for Gene Ontology Cell Types

  
alldata.1 <- subset(alldata, ident = "Fibroblasts")  
Idents(alldata.1) <- "timepoint"


features = c("S100a8","S100a9","Ccl22","Cxcl12", #D4PW Leukocote Recruitment, Tissue Remodeling
                          "Acta2","Nox4", #Contraction
             "Egfr","Esr1","Pdgfd","Fgf2","Pak3", #D7PW Fibroblast Profliferation/Migration
             "Wnt10b","Msx2" #Anagen
                 ) 
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "red3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotFibroblastsGO.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots for IFE

```{r}
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "IFE")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl3","Il1b",
             "Ubb","Sfn",#Cell Inflamation
             "Egfr","Notch2","Bmpr1a","Fgfr2",
             
             "Mki67", "Barx2","Dlx2",
             "Myc")
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "royalblue4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotIFEGO.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots for HF

```{r}
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "HF")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl3","Fcgr2b", "Cxcl14","Ubc", #Inflammation
             "Ext1","Lrig1", "Lgals1","Atf2", #Axonogensis/Muscle
             "Krt27","Dlx3","Smad3", #Hair Follicle
             "Sox9"
             
        )
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "skyblue4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotHFGO.tiff",dpi = 300,height = 3.5, width = 8 )

```

Gene Ontology Dot Plots for Sebocytes

```{r}
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "Sebocytes")  
Idents(alldata.1) <- "timepoint"


features = c("Cxcl2","Ccl4","Tagln2","S100a11", #Inflammation
             "Cdk14","Tanc2a", "Efna5", #Hair Follicle
             "Lef1","Sox5", "Mef2c", "Esr1","Bmpr2"
        )
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "turquoise4")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotSebocytesGO.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots for Schwann Cells

```{r}
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "Schwann cells")  
Idents(alldata.1) <- "timepoint"


features = c("Cttn","Cfl1","Tubb2b", #modification of postsynaptic structure
             "Ptprg","Nf1","Ptprz1","Magi2", #Hair Follicle
             "Chl1","Hmox1"
        )
             


p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "darkseagreen3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotSchwannCells.tiff",dpi = 300,height = 3.5, width = 8 )
```

Gene Ontology Dot Plots for Skeletal Muscle

```{r}
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "Muscle cells")  
Idents(alldata.1) <- "timepoint"


features = c("S100a9","Il1b","Ccl3", #modification of postsynaptic structure
             "Ttn","Mlip3","Mef2c","Sox6", #Hair Follicle
             "Smad4","Lef1","Notch1", "Dab2"
        )
             

p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "orange3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotSkeletalMuscle.tiff",dpi = 300,height = 3.5, width = 8 )

```

Gene Ontology Dot Plots for Skeletal Muscle

```{r}
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "Endothelial")  
Idents(alldata.1) <- "timepoint"


features = c("Serpine1","Spp1","Ccl3","Thbs1","Egr1", #modification of postsynaptic structure
             "Ptprm","Cdh13","Vegfc", #Hair Follicle
             "Bmpr1a","Dlx3","Cdh3", "Lncpint")
             

p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "gold3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotEndothelial.tiff",dpi = 300,height = 3.5, width = 8 )

```

Gene Ontology for Hair Follicle

```{r}
#DotPlots for Gene Ontology Cell Types

alldata.1 <- subset(alldata, ident = "Endothelial")  
Idents(alldata.1) <- "timepoint"


features = c("Serpine1","Spp1","Ccl3","Thbs1","Egr1", #modification of postsynaptic structure
             "Ptprm","Cdh13","Vegfc", #Hair Follicle
             "Bmpr1a","Dlx3","Cdh3", "Lncpint")
             

p <- DotPlot(alldata.1, features, dot.scale = 7, cols = c("white", "gold3")) + 
  RotatedAxis() +
  theme(
    text = element_text(size = 15, face = "bold", family = "Arial")
  ) +
  scale_size(range = c(0, 7)) + # Ensures size consistency
  geom_point(aes(size = pct.exp), shape = 21, color = "black") 
# Outlines circles in black
p
# Display the plot
ggsave("DotPlotEndothelial.tiff",dpi = 300,height = 3.5, width = 8 )
```

```{r}
#Phylogentic Genes
alldata <- BuildClusterTree(alldata, reduction = "pca")
PlotClusterTree(alldata, direction = "rightwards")
tiff("AllCellsSubset_PhyloTree.tiff")
PlotClusterTree(alldata, direction = "rightwards")
dev.off()
```


Run Cell Chat on the Global Cell Types

```{r}
##Prep the cellchat database
CellChatDB <- CellChatDB.mouse

showDatabaseCategory(CellChatDB)
#This removes non-protein interactions that are contained in the CellChat DB such as synaptic interactions and metabolic signals
CellChatDB.use <- subsetDB(CellChatDB)

```

```{r}
#Hold off Until RoadMap
cellchat <- createCellChat(object = alldata.1, group.by = "celltype", assay = "RNA")
levels(cellchat@idents)
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multisession", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

```

```{r BarPlot Showing Cell Number Across Samples}
# Load necessary libraries
library(Seurat)
library(dplyr)
library(ggplot2)

# Step 1: Extract the orig.ident counts
orig_counts <- table(alldata$orig.ident) %>% as.data.frame()
colnames(orig_counts) <- c("orig.ident", "cell_count")

# Step 2: Add group based on substrings in the orig.ident names
orig_counts$group <- case_when(
  grepl("Haensel", orig_counts$orig.ident)    ~ "Haensel (2020)",
  grepl("Vu", orig_counts$orig.ident)         ~ "Vu (2022)",
  grepl("snRNAseq", orig_counts$orig.ident)   ~ "snRNAseq",
  grepl("CITEseq", orig_counts$orig.ident)    ~ "CITEseq",
  grepl("Justynski", orig_counts$orig.ident)  ~ "Justynski (2023)",
  TRUE                                        ~ "Other"
)

# Step 3: Summarize total cells per group
group_counts <- orig_counts %>%
  group_by(group) %>%
  summarise(total_cells = sum(cell_count)) %>%
  arrange(desc(total_cells))

# Step 4: Plot
# Step 4: Horizontal bar plot with black outlines
ggplot(group_counts, aes(x = reorder(group, total_cells), y = total_cells, fill = group)) +
  geom_bar(stat = "identity", width = 0.9, color = "black", size = 0.8) +
  geom_text(aes(label = total_cells),
            hjust = -0.4,
            size = 4.5,
            fontface = "bold") +
  labs(title = "Total Cell Counts by Dataset Group",
       x = "Sample Group",
       y = "Number of Cells") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  coord_flip() +
  theme_ggprism_mod(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text.y = element_text(face = "bold")
  ) +
  custom_theme

ggsave("Grouped_Barplot_by_SampleIdentiy.png",
       width = 8, height = 5, dpi = 300)
```
