---
title: "Spatial Figure Making"
author: "Simon Van Deursen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Sets root directory for chunks to the directory where the file is located
knitr::opts_knit$set(root.dir = "~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 3 Spatial")

library(Seurat)
library(ggplot2)
library(patchwork)
library(cowplot)
library(scales)
library(viridis)
library(dplyr)
library(tidyr)
library(purrr)
library(writexl)
library(SeuratWrappers)
library(Banksy)
library(hdf5r)
library(spacexr)
library(sfarrow)
library(Rfast2)
library(rlang)
#install.packages("ggsignif")  # only once
library(ggsignif)
library(colorway)
library(khroma)
library(ggbreak)
library(extrafont)

#font_import() # Needed to import fonts from computer 
#Load needed functions
source("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/scRNA SEQ of wounds/Skin Wound Healing Atlas/Useful Functions/GGsave_multidevice.R")
source("~/Documents/Shared-Utils/Useful Functions/CustomSpatialHighlightPlot2.R")

# needs to be set for large dataset analysis
options(future.globals.maxSize = 1e9)

#Save a custom theme for ggplots, including setting Arial as default font and
#making background transparent
custom_theme <- theme(
  text = element_text(family = "Arial"),
  plot.background = element_rect(fill = "transparent", color = NA),
  panel.background = element_rect(fill = "transparent", color = NA),
  legend.background = element_rect(fill = "transparent"),
  panel.grid = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)
```
Load in the datasets
```{r Load data}
object_uw <- readRDS("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Users/Simon/Spatial Data/RDS Files/YWV01_UW_FineAnnotation_092525.rds")

object_d4 <- readRDS("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Users/Simon/Spatial Data/RDS Files/YWV04_D4PW_FineAnnotation_092525.rds")

object_d7 <- readRDS("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Users/Simon/Spatial Data/RDS Files/YWV03_D7PW_FineAnnotation_092525.rds")

object_d30 <- readRDS("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Users/Simon/Spatial Data/RDS Files/YWV02_D30PW_FineAnnotation_092525.rds")
```

For ease of plotting, generate a list of objects
```{r Generate dataset list for plotting}
# List of all conditions with metadata
plot_info <- list(
  "YWV01_UW"   = list(object = object_uw,  regions = c("Region1", "Region2","ROI"), pt.size.factor = 1.2),
  "YWV04_D4PW" = list(object = object_d4, regions = c("Region1", "Region2","ROI1"),   pt.size.factor = 1.25),
  "YWV03_D7PW" = list(object = object_d7, regions = c("Region1", "Region2","ROI1"),   pt.size.factor = 1.25),
  "YWV02_D30PW"= list(object = object_d30, regions = c("Region1", "Region2","ROI"), pt.size.factor = 1)
)
```
Get counts of each region
```{r}
table(object_uw$Region)
table(object_d4$Region)
table(object_d7$Region)
table(object_d30$Region)
```

Save scatter plots of the relationship between nCount and nFeature
```{r Plot QC Scatter plot}
for(sample_name in names(plot_info)) {
  info <- plot_info[[sample_name]]
  seurat_obj <- info$object
  
  FeatureScatter(seurat_obj, feature1 = "nCount_Spatial", feature2 = "nFeature_Spatial",
               group.by = "orig.ident", cols = "grey25", raster = F) & custom_theme
  
  ggsave_multidevice(filename_base = paste0(sample_name,"nCount_nFeature_Scatter"), height = 7, width = 8)
}
```

```{r Loop to save plots of anatomical sites}
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 3 Spatial/Anatomical Sites")

anatomical_sites <- union(unique(object_d7$AnatomicalSite),unique(object_d30$AnatomicalSite))
# Set levels of anatomical sites
anatomical_sites <- factor(anatomical_sites, levels = sort(anatomical_sites))

anatomical_pal <- setNames(scCustomize::Hue_Pal(length(anatomical_sites)), levels(anatomical_sites)) 

# Loop over each object and its regions
for (sample_name in names(plot_info)) {
  info <- plot_info[[sample_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  for (region in regions) {
    p <- SpatialDimPlot(seurat_obj, group.by = "AnatomicalSite", images = region,
                        pt.size.factor = 3) &
      scale_fill_manual(values = anatomical_pal) &
      custom_theme +
      theme(panel.grid = element_blank())

    ggsave_multidevice(filename_base = paste0(sample_name, "_AnatomicalSite_", region))
    
  }
}
```

```{r Loop to save plots of 10X graph clusters}
# Loop over each object and its regions
for (sample_name in names(plot_info)) {
  info <- plot_info[[sample_name]]
  seurat_obj <- info$object
  regions <- info$regions
  pt.size <- info$pt.size.factor
  cluster_levels <- unique(seurat_obj$alignment_clusters)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), sort(cluster_levels))
  
  for (region in regions) {
    p <- SpatialDimPlot(seurat_obj, group.by = "alignment_clusters", images = region,
                        pt.size.factor = pt.size) &
      scale_fill_manual(values = cluster_colors) &
      custom_theme +
      theme(panel.grid = element_blank())

    ggsave_multidevice(filename_base = paste0(sample_name, "_10XClustering_", region))
    
    p <- SpatialDimPlot(seurat_obj, group.by = "alignment_clusters", images = region,
                        pt.size.factor = pt.size, label = T, repel = T, label.size = 3) & 
      scale_fill_manual(values = cluster_colors) & 
      custom_theme +
      theme(panel.grid = element_blank())
    
    ggsave_multidevice(filename_base = paste0(sample_name, "_10XClusteringLabeled_", region))
  }
}
```
We can also generate plots highlighting each cluster in its respective colors
```{r UW 10X clustering split plot}
Idents(object_uw) <- "alignment_clusters"
# Get cells grouped by metacluster
cells <- CellsByIdentities(object_uw)

full_pal <- setNames(viridis(length(levels(object_uw$alignment_clusters)),
                                     option = "H"), levels(object_uw$alignment_clusters)) 
# Build a full palette including “Unselected" - necessary to change background color
full_pal <- c(
  full_pal,
  Unselected = "grey65"
)

# Make SpatialDimPlot
p <- SpatialDimPlot(
  object_uw,
  images          = "slice1",
  cells.highlight = cells,
  cols.highlight  = full_pal,  # placeholder
  facet.highlight = TRUE,
  combine         = TRUE
) + NoLegend() & custom_theme

# Override the fill scale
p <- p & scale_fill_manual(
  values = full_pal,
  breaks = names(full_pal)  # ensures both the metacluster and “Unselected" levels are known
)

p & theme(title = element_text(face = "bold"),
          panel.grid = element_blank())

#ggsave_multidevice(filename_base = "YWV01_UW_10XClustersSplitDimPlot", width = 10, height = 15)
```
```{r Plot QC Scatter plot}
for(sample_name in names(plot_info)) {
  info <- plot_info[[sample_name]]
  seurat_obj <- info$object
  
  FeatureScatter(seurat_obj, feature1 = "nCount_Spatial", feature2 = "nFeature_Spatial",
               group.by = "orig.ident", cols = "grey25", raster = F) & custom_theme
  
  ggsave_multidevice(filename_base = paste0(sample_name,"nCount_nFeature_Scatter"), height = 7, width = 8)
}
```

```{r Loop to save plots of seurat clusters}
# Loop over each object and its regions
for (sample_name in names(plot_info)) {
  info <- plot_info[[sample_name]]
  seurat_obj <- info$object
  regions <- info$regions
  pt.size <- info$pt.size.factor
  cluster_levels <- levels(seurat_obj$seurat_cluster.projected)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), cluster_levels)
  
  for (region in regions) {
    p <- SpatialDimPlot(seurat_obj, group.by = "seurat_cluster.projected", images = region) &
      scale_fill_manual(values = cluster_colors) &
      custom_theme +
      theme(panel.grid = element_blank())

    ggsave_multidevice(filename_base = paste0(sample_name, "_SeuratClustering_", region))
    
    p <- SpatialDimPlot(seurat_obj, group.by = "seurat_cluster.projected", images = region,
                        label = T, repel = T, label.size = 3, pt.size.factor = pt.size) & 
      scale_fill_manual(values = cluster_colors) & 
      custom_theme +
      theme(panel.grid = element_blank())
    
    ggsave_multidevice(filename_base = paste0(sample_name, "_SeuratClusteringLabeled_", region))
    
  }
}
```
```{r Loop to save plots of BANKSY clusters}
# Loop over each object and its regions
for (sample_name in names(plot_info)) {
  info <- plot_info[[sample_name]]
  seurat_obj <- info$object
  regions <- info$regions
  pt.size <- info$pt.size.factor
  cluster_levels <- levels(seurat_obj$banksy_cluster_0.5)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), cluster_levels)
  
  for (region in regions) {
    p <- SpatialDimPlot(seurat_obj, group.by = "banksy_cluster_0.5", images = region) &
      scale_fill_manual(values = cluster_colors) &
      custom_theme +
      theme(panel.grid = element_blank())

    ggsave_multidevice(filename_base = paste0(sample_name, "_BANKSYClustering_", region))
    
    p <- SpatialDimPlot(seurat_obj, group.by = "banksy_cluster_0.5", images = region,
                        label = T, repel = T, label.size = 3, pt.size.factor = pt.size) & 
      scale_fill_manual(values = cluster_colors) & 
      custom_theme +
      theme(panel.grid = element_blank())
    
    ggsave_multidevice(filename_base = paste0(sample_name, "_BANKSYClusteringLabeled_", region))
    
  }
}
```
Similarly, we want to save plots of cell type annotations, both by cell type and
metacluster. We will start by saving defined palettes from each of these
```{r Save cell type color palette}
# Set color palette for main metacluster colors
metacluster_colors <- c(
  "IFE Keratinocytes" = "#364B9A",
  "HF Keratinocytes"  = "#56B4E9",
  "Sebocytes"         = "#8DD3C7",
  "Fibroblasts"       = "indianred2",
  "Immune cells"      = "#198F1B",
  "Endothelial cells" = "#F6C800",
  "Pericytes"         = "#D12E82",
  "Muscle cells"            = "orange2",
  "Schwann cells"     = "purple",
  "Melanocytes"       = "pink1",
  "Adipocytes"        = "deeppink",
  "Red Blood Cells"   = "#A7222B",
  "Unknown" = "grey70"
)
# Load fine cell type color palette
cell_colors <- readRDS("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Figure 1/color_palette.rds")
# Add Unknown to cell palettte
cell_colors <- c(cell_colors, "Unknown" = "white")
```

```{r Loop to save cell type and metacluster plots}
# Loop over each object and its regions, plotting cell types with associated palette
for (sample_name in names(plot_info)) {
  info <- plot_info[[sample_name]]
  seurat_obj <- info$object
  regions <- info$regions
  cluster_levels <- levels(seurat_obj$first_type)
  pt.size <- plot_info[[sample_name]]$pt.size.factor
  
  for (region in regions) {
    # Plot cell types unlabeled
    p <- SpatialDimPlot(seurat_obj, group.by = "first_type", images = region, pt.size.factor = pt.size) & 
      scale_fill_manual(values = cell_colors) & 
      custom_theme + NoLegend() +
      theme(panel.grid = element_blank())
    
    ggsave_multidevice(filename_base = paste0(sample_name, "_RCTDCellType_", region))
    
    # Plot cell type labeled
    p <- SpatialDimPlot(seurat_obj, group.by = "first_type", images = region, 
                        pt.size.factor = pt.size, label = T, repel = T, label.size = 2) & 
      scale_fill_manual(values = cell_colors) & 
      custom_theme + NoLegend() +
      theme(panel.grid = element_blank())
    
    ggsave_multidevice(filename_base = paste0(sample_name, "_RCTDCellTypeLabeled_", region))
    
    # Plot metaclusters for each region
    p <- SpatialDimPlot(seurat_obj, group.by = "metaclusters", images = region,
                        pt.size.factor = pt.size) & 
      scale_fill_manual(values = metacluster_colors) & 
      custom_theme + NoLegend() +
      theme(panel.grid = element_blank())
    
    ggsave_multidevice(filename_base = paste0(sample_name, "_RCTDMetaclusters_", region))
  }
}
```
Save UMAP plots with the 10X clustering or metacluster annotations overlayed
```{r Save UMAPs}
# Loop to save annotated UMAPs
for(sample_name in names(plot_info)) {
  
  seurat_obj <- plot_info[[sample_name]]$object
  
  # 10X clustering 
  cluster_levels <- unique(seurat_obj$alignment_clusters)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), sort(cluster_levels))
 
  p <- DimPlot(seurat_obj, reduction = "full.umap.sketch", group.by = "alignment_clusters",
               raster = F, cols = cluster_colors) + 
    labs(title = "") + 
    NoAxes() & 
    custom_theme
    
  p + NoLegend() 

  ggsave_multidevice(filename_base = paste0(sample_name,"10XClusterUMAP_NoLegend"), height = 7, width = 7)
  ggsave_multidevice(plot = p, filename_base = paste0(sample_name,"10XClusterUMAP_Legend"), height = 7, width = 8)
  
  # Seurat clustering 
  cluster_levels <- levels(seurat_obj$seurat_cluster.projected)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), cluster_levels)
  
  p <- DimPlot(seurat_obj, reduction = "full.umap.sketch", group.by = "seurat_cluster.projected",
               raster = F, cols = cluster_colors) + 
    labs(title = "") + 
    NoAxes() & 
    custom_theme
    
  p + NoLegend() 

  ggsave_multidevice(filename_base = paste0(sample_name,"SeuratClusterUMAP_NoLegend"), height = 7, width = 7)
  ggsave_multidevice(plot = p, filename_base = paste0(sample_name,"SeuratClusterUMAP_Legend"), height = 7, width = 8)
  
  # BANKSY clustering
  cluster_levels <- levels(seurat_obj$banksy_cluster_0.8)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), cluster_levels)
  
  p <- DimPlot(seurat_obj, reduction = "full.umap.sketch", group.by = "banksy_cluster_0.8",
               raster = F, cols = cluster_colors) + 
    labs(title = "") + 
    NoAxes() & 
    custom_theme
    
  p + NoLegend() 
  ggsave_multidevice(filename_base = paste0(sample_name,"BANKSYClusterUMAP_NoLegend"), height = 7, width = 7)
  ggsave_multidevice(plot = p, filename_base = paste0(sample_name,"BANKSYClusterUMAP_Legend"), height = 7, width = 8)
  
  # Repeat for metaclusters
  p <- DimPlot(seurat_obj, reduction = "full.umap.sketch", group.by = "metaclusters",
               raster = F, cols = metacluster_colors) + 
    labs(title = "") & 
    custom_theme
  
  p + NoLegend() + NoAxes() 

  ggsave_multidevice(filename_base = paste0(sample_name,"MetaclusterUMAP_NoLegend"), height = 7, width = 7)

  ggsave_multidevice(plot = p, filename_base = paste0(sample_name,"MetaclusterUMAP_Legend"), height = 7, width = 8)
}
```
We also need to save the fine cell type legend, which should be the same across all objects
```{r Save fine cell type legend}
p <- DimPlot(object_d7, cols = cell_colors, group.by = "first_type",
             reduction = "full.umap.sketch") + 
  theme(text = element_text(family = "Arial", size = 14)) & custom_theme

# Extract the legend using cowplot
library(cowplot)
legend <- get_legend(p)

ggsave("FineCellTypeLegend.png", plot = legend, width = 12, height = 5, 
       dpi = 300, bg = "transparent")
```
We are also interested in the quantification of the representation of celltypes
within each cluster
```{r Quantify metacluster representation by cluster}
# Quantify metacluster composition within defined clusters
quantify_metacluster_representation <- function(obj, sample_name, annotation_col) {
  meta <- obj@meta.data
  
  # Dynamically select the annotation column and metacluster column
  meta <- meta %>%
    dplyr::select(annotation = all_of(annotation_col),
                  metacluster = metaclusters) %>%
    filter(!is.na(annotation), !is.na(metacluster))
  
  # Count and normalize
  count_table <- meta %>%
    dplyr::count(annotation, metacluster, name = "count") %>%
    group_by(annotation) %>%
    mutate(proportion = count / sum(count)) %>%
    ungroup() %>%
    mutate(sample = sample_name)
  
  return(count_table)
}

# Define clustering columns of interest
clustering_cols <- c("alignment_clusters", "seurat_cluster.projected", "banksy_cluster_0.8")

# Generate all results for each clustering column
all_results_list <- map(clustering_cols, function(cluster_col) {
  
  # For each sample
  results_df <- map2_dfr(plot_info, names(plot_info), function(info, sample_name) {
    result <- quantify_metacluster_representation(info$object, sample_name, annotation_col = cluster_col)
    
    # Optional: generate a barplot for each sample & clustering
    p <- ggplot(result, aes(x = annotation, y = proportion, fill = metacluster)) +
      geom_bar(stat = "identity") +
      theme_minimal() +
      labs(title = paste(sample_name, "-", cluster_col),
           x = "Cluster", y = "Proportion", fill = "Metacluster") +
      scale_fill_manual(values = metacluster_colors) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
            axis.title.x = element_text(size = 14),
            axis.text.y = element_text(size = 12, angle = 45),
            axis.title.y = element_text(size = 14)) +
      custom_theme + NoLegend()
    
    ggsave_multidevice(
      filename_base = paste0(sample_name, "_", cluster_col, "_MetaclusterCluster_Barplot"),
      plot = p, width = 8, height = 5
    )
    
    return(result)
  })
  
  return(results_df)
})

names(all_results_list) <- clustering_cols

# Export to Excel
write_xlsx(all_results_list, path = "Metacluster_Cluster_Representation.xlsx")
```
Similarly, we want to compare cell type counts and proportions across samples
```{r Compare metacluster counts and props across objects}
get_metacluster_counts <- function(seurat_obj, annotation_col, label, scaling_factor = 10000) {
  # Count raw metacluster values by Region
  count_data <- seurat_obj@meta.data %>%
    dplyr::count(Region, !!sym(annotation_col)) %>%
    group_by(Region) %>%
    mutate(
      Normalized = n / sum(n) * scaling_factor,
      Proportion = n / sum(n) * 100
    ) %>%
    ungroup() %>%
    mutate(Object = label)
  
  # Average across regions
  summary_data <- count_data %>%
    group_by(!!sym(annotation_col), Object) %>%
    summarise(
      Normalized = mean(Normalized),
      Proportion = mean(Proportion),
      .groups = "drop"
    )
  
  # Return both in a named list
  return(list(
    region_counts = count_data,
    summary = summary_data
  ))
}

# Extract from each object
plot_data <- bind_rows(
  get_metacluster_counts(object_uw, annotation_col = "metaclusters", "Unwounded")$summary,
  get_metacluster_counts(object_d4, annotation_col = "metaclusters", "D4PW")$summary,
  get_metacluster_counts(object_d7, annotation_col = "metaclusters", "D7PW")$summary,
  get_metacluster_counts(object_d30, annotation_col = "metaclusters", "D30PW")$summary
)

# Get separator positions for vertical dashed lines
object_levels <- c("Unwounded","D4PW","D7PW","D30PW")
plot_data$Object <- factor(plot_data$Object, levels = object_levels)
separator_positions <- seq(1.5, length(object_levels) - 0.5, by = 1)

# Plot 1: Normalized bar plot (dodge)
p1 <- ggplot(plot_data, aes(x = Object, y = Normalized, fill = metaclusters)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = metacluster_colors) +
  labs(y = "Cell Count per 10,000 Cells",
       title = "Normalized Metacluster Counts") +
  geom_vline(xintercept = separator_positions, linetype = "dashed", color = "gray50") +
  theme_bw() +
  theme(
    text = element_text(face = "bold"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14),
    plot.title = element_text(size = 16, hjust = 0.5)
  )

# Plot 2: Stacked bar plot with proportions
p2 <- ggplot(plot_data, aes(x = Object, y = Proportion, fill = metaclusters)) +
  geom_bar(stat = "identity", width = 0.6) +
  scale_fill_manual(values = metacluster_colors) +
  labs(y = "Proportion (%)",
       title = "Average Metacluster Proportions per Object") +
  geom_vline(xintercept = separator_positions, linetype = "dashed", color = "gray50") +
  theme_void() +
  theme(
    text = element_text(face = "bold"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14, angle = 90),
    plot.title = element_text(size = 16, hjust = 0.5)
  )

# Print plots
print(p1)
print(p2)

ggsave_multidevice(filename_base = "MetaclusterComparison_CountPlot", plot = p1,
                   width = 10, height = 8)
ggsave_multidevice(filename_base = "MetaclusterComparison_PropPlot", plot = p2,
                   width = 6, height = 5)
```
Utilizing normalized mutual information, we can compare the relationships between
different clusterings and the cell type annotations. NMI is a metric which quantifies
the relationships between two clustering methods to determine how much they agree,
with 0 being completely different and 1 being completely identical. By comparing
these values across the clustering methods used, we can determine how much one clustering
method is grouping cell types together.
```{r Run NMI on spatial clusterings vs cell types}
if (!requireNamespace("aricode")) install.packages("aricode")
library(aricode)
library(ggplot2)
library(dplyr)

# Define NMI calculation function for one object
calculate_nmi <- function(seurat_obj, sample_name) {
  # Extract vectors
  cluster_methods <- c("alignment_clusters", "banksy_cluster_0.5", "seurat_cluster.projected")
  cell_type_vector <- as.character(seurat_obj$first_type)
  
  nmi_scores <- sapply(cluster_methods, function(method) {
    cluster_vector <- as.character(seurat_obj[[method]][,1])  # pull as vector
    NMI(cluster_vector, cell_type_vector)
  })
  
  data.frame(
    Sample = sample_name,
    Method = c("10X Graph Clusters", "BANKSY Clusters", "Seurat Clusters"),
    NMI = as.numeric(nmi_scores)
  )
}

# Run NMI calculation for all samples in plot_info
nmi_all <- do.call(rbind, lapply(names(plot_info), function(name) {
  obj <- plot_info[[name]]$object
  calculate_nmi(obj, name)
}))

nmi_all$Sample <- factor(nmi_all$Sample, levels = c("YWV01_UW","YWV04_D4PW","YWV03_D7PW","YWV02_D30PW"))
# Plot
ggplot(nmi_all, aes(x = Sample, y = NMI, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = round(NMI, 3)), 
            position = position_dodge(width = 0.8), vjust = -0.4, size = 3.5) +
  labs(title = "NMI between Clustering Methods and Cell Type Annotations",
       x = "Dataset", y = "Normalized Mutual Information") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10)) +
  scale_fill_brewer(palette = "Set2")
ggsave("ClusteringAnnotationNMI_Barplot.png", width = 8, height = 5)
```

## Cell type quantification
### By Fine cell type
Compare cell type counts across conditions, averaging across regions
```{r Define a function to calculate raw and normalized cell counts and proportions}
calculate_and_plot <- function(object, source_label, annotation_column) {
  
  metadata <- object@meta.data
  annotation_sym <- sym(annotation_column)
  
  # Total cells per region
  total_cells_per_region <- metadata %>%
    group_by(Region) %>%
    summarize(Total_Cells = n(), .groups = "drop")

  avg_total_cells <- mean(total_cells_per_region$Total_Cells)

  # Count, normalize, and compute Per_10k_Cells
  region_counts <- metadata %>%
    group_by(.data[[annotation_column]], Region) %>%
    summarize(Count = n(), .groups = "drop") %>%
    left_join(total_cells_per_region, by = "Region") %>%
    mutate(
      Proportion = (Count / Total_Cells) * 100,
      Normalized_Count = (Count / Total_Cells) * avg_total_cells, # scales to the average size of both regions in an object
      Per_10k_Cells = (Count / Total_Cells) * 10000 # scales to a set value for comparison across objects
    ) %>%
    ungroup()
  
  # Compute mean and SEM per cell type for error bars
  summary_stats <- region_counts %>%
    group_by(.data[[annotation_column]]) %>%
    summarize(
      Avg_Proportion = mean(Proportion),
      Mean_Normalized = mean(Normalized_Count),
      SEM_Normalized = sd(Normalized_Count) / sqrt(n()),
      Mean_Per10k = mean(Per_10k_Cells),
      SEM_Per10k = sd(Per_10k_Cells) / sqrt(n()),
      .groups = "drop"
    ) %>%
    mutate(Source = source_label)

  # Plots
  ## Plot of raw cell count by region
  count_plot <- ggplot(region_counts, aes(x = !!annotation_sym, y = Count, fill = Region)) +
    geom_col(position = position_dodge()) +
    theme_bw() +
    labs(
      title = paste("Cell Counts by Region -", source_label),
      x = "Cell Type", y = "Count"
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  ## Plot of proportion
  proportion_plot <- ggplot(summary_stats, aes(x = Source, y = Avg_Proportion, fill = !!annotation_sym)) +
    geom_col(position = "stack", color = "black") +
    theme_bw() +
    labs(title = paste("Stacked Average Cell Type Proportions -", source_label),
         x = "Source", y = "Average Proportion (%)", fill = "Cell Type")
  
  ## Cell counts normalized to average region count
  normalized_count_plot <- ggplot(region_counts, aes(x = !!annotation_sym, y = Normalized_Count, fill = Region)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    theme_bw() +
    labs(
      title = paste("Normalized Cell Counts by Region -", source_label),
      x = "Cell Type", y = "Normalized Count"
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  ## Average normalized counts with SEM
  mean_normalized_count_plot <- ggplot(summary_stats, aes(x = !!annotation_sym, y = Mean_Normalized, fill = !!annotation_sym)) +
    geom_col(color = "black") +
    geom_errorbar(aes(ymin = Mean_Normalized - SEM_Normalized,
                      ymax = Mean_Normalized + SEM_Normalized),
                  width = 0.3) +
    theme_bw() +
    labs(title = paste("Normalized Cell Counts (Mean ± SEM) -", source_label),
         x = "Cell Type", y = "Mean Normalized Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  ## Normalized to 10k cells with SEM
  mean_per_10k_plot <- ggplot(summary_stats, aes(x = !!annotation_sym, y = Mean_Per10k, fill = !!annotation_sym)) +
    geom_col(color = "black") +
    geom_errorbar(aes(ymin = Mean_Per10k - SEM_Per10k,
                      ymax = Mean_Per10k + SEM_Per10k),
                  width = 0.3) +
    theme_bw() +
    labs(title = paste("Mean Count per 10k Cells (± SEM) -", source_label),
         x = "Cell Type", y = "Cell Count per 10,000 Cells") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  return(list(
    region_counts = region_counts,
    summary_stats = summary_stats,
    avg_total_cells = avg_total_cells,
    plots = list(
      count_plot = count_plot,
      proportion_plot = proportion_plot,
      normalized_count_plot = normalized_count_plot,
      mean_normalized_count_plot = mean_normalized_count_plot,
      mean_per_10k_plot = mean_per_10k_plot
    )
  ))
}
```
With the function defined, we can apply it to the object and save the results as an excel file
```{r Quantify fine cell types using function and save results}
# Use lapply to apply the above function to each Visium object
results_list <- lapply(names(plot_info), function(dataset) {
  info <- plot_info[[dataset]]
  
  # Run the function
  result_object <- calculate_and_plot(
    object = info$object,
    source_label = dataset,
    annotation_column = "first_type"
  )
  
  # Save Excel file
  writexl::write_xlsx(
    list(
      "Region Counts" = result_object$region_counts,
      "Summary Data" = result_object$summary_stats
    ),
    paste0(dataset, "_FineCellTypeQuantification.xlsx")
  )
  
  # Return the result to keep in memory as well
  return(result_object)
})

# Name the list with dataset names
names(results_list) <- names(plot_info)
```

```{r Iterate over the stored plots to save fine cell type quantification by region}
# Define custom plot scales and colors for each dataset
plot_scales <- list(
  "YWV01_UW" = list(
    colors = c("navyblue","lightskyblue"),
    y_break = list(breaks = c(5000,15000), scales = 0.25, ticklabels = c(15000, 20000))
  ),
  "YWV04_D4PW" = list(
    colors = c("firebrick","lightcoral"),
    y_break = list(breaks = c(6500,9000), scales = 0.25, ticklabels = c(10000,12000))
  ),
  "YWV03_D7PW" = list(
    colors = c("darkorange3","orange"),
    y_break = list(breaks = c(10000,20000), scales = 0.25, ticklabels = c(20000, 25000))
  ),
  "YWV02_D30PW" = list(
    colors = c("goldenrod3","khaki"),
    y_break = list(breaks = c(6000,12000), scales = 0.25, ticklabels = c(12000, 15000))
  )
)

# Iterate through results_list and save plots
for(dataset in names(results_list)){
  res <- results_list[[dataset]]
  scales <- plot_scales[[dataset]]
  
  p <- res$plots$normalized_count_plot + 
      scale_fill_manual(values = scales$colors) + 
      scale_y_break(
        breaks = scales$y_break$breaks, 
        scales = scales$y_break$scales, 
        ticklabels = scales$y_break$ticklabels) +
      theme(
      text = element_text(face = "bold"),
      axis.text.x = element_text(size = 11, angle = 90, vjust = 0.5),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 16),
      axis.title.y = element_text(size = 18, hjust = 0.75)
    ) + custom_theme 
  
  ggsave(filename = paste0(dataset, "_NormalizedCountsByRegion.png"), 
         plot = p, width = 15, height = 6, dpi = 300)
}
```
We can repeat the same process for other plots, including the mean cell counts per 10k cells
```{r Save plots of mean cell counts per 10k cells}
# Define custom plot scales and colors for each dataset
plot_scales <- list(
  "YWV01_UW" = list(breaks = c(875, 3000), scales = 0.1, ticklabels = c(3000,3500)),
  "YWV04_D4PW" = list(breaks = c(1250, 1750), scales = 0.1, ticklabels = c(1800,2200)),
  "YWV03_D7PW" = list(breaks = c(1100, 1750), scales = 0.1, ticklabels = c(2000,2600)),
  "YWV02_D30PW" = list(breaks = c(800, 1750), scales = 0.1, ticklabels = c(1800, 2200))
)

# Iterate through results_list and save plots
for(dataset in names(results_list)){
  res <- results_list[[dataset]]
  scales <- plot_scales[[dataset]]
  
  p <- res$plots$mean_per_10k_plot +                          
      scale_fill_manual(values = cell_colors) +
      scale_y_break(
        breaks = scales$breaks, 
        scales = scales$scales, 
        ticklabels = scales$ticklabels,
        expand = expansion(mult = c(0.01, 0.05))) +
      theme(
        text = element_text(face = "bold"),
        plot.title = element_text(size = 18, hjust = 0.4),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 16, vjust = 1.5),
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank()
      ) + NoLegend() + custom_theme 
  
  ggsave_multidevice(filename_base = paste0(dataset, "_CellCountsper10kBarplot"), 
         plot = p, width = 15, height = 6, dpi = 300)
  
 p <- res$plots$mean_per_10k_plot +                          
  scale_fill_manual(values = cell_colors) +
      scale_y_break(
        breaks = scales$breaks, 
        scales = scales$scales, 
        ticklabels = scales$ticklabels,
        expand = expansion(mult = c(0.01, 0.05))) +
   coord_flip() +
 theme(
    text = element_text(face = "bold"),
    plot.title = element_text(size = 18, hjust = 0.4),
    axis.text.x = element_text(hjust = 1, size = 14, angle = 45),
    axis.text.x.top = element_blank(),
    axis.ticks.x.top = element_blank(),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_text(size = 12, angle = 0),
    axis.title.y = element_blank(),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank()
  ) + 
   NoLegend() + custom_theme 
 
ggsave_multidevice(filename_base = paste0(dataset, "_CellCountsper10kBarplot_Vertical"),
                  plot = p, width = 8, height = 18)
}
```

### By Metacluster
```{r Quantify metaclusters and save results}
# Use lapply to apply the above function to each Visium object
results_list <- lapply(names(plot_info), function(dataset) {
  info <- plot_info[[dataset]]
  
  # Run the function
  result_object <- calculate_and_plot(
    object = info$object,
    source_label = dataset,
    annotation_column = "metaclusters"
  )
  
  # Save Excel file
  writexl::write_xlsx(
    list(
      "Region Counts" = result_object$region_counts,
      "Summary Data" = result_object$summary_stats
    ),
    paste0(dataset, "_MetaclusterQuantification.xlsx")
  )
  
  # Return the result to keep in memory as well
  return(result_object)
})

# Name the list with dataset names
names(results_list) <- names(plot_info)
```

```{r Save plots of mean cell counts per 10k cells}
# Define custom plot scales and colors for each dataset
plot_scales <- list(
  "YWV01_UW" = list(breaks = c(2250, 3000), scales = 0.1, ticklabels = c(3500)),
  "YWV04_D4PW" = list(breaks = c(2000, 4750), scales = 0.1, ticklabels = c(5000)),
  "YWV03_D7PW" = list(breaks = c(2000, 4500), scales = 0.1, ticklabels = c(5500)),
  "YWV02_D30PW" = list(breaks = c(800, 1750), scales = 0.1, ticklabels = c(1800, 2200))
)

# Iterate through results_list and save plots
for(dataset in names(results_list)){
  res <- results_list[[dataset]]
  scales <- plot_scales[[dataset]]
  
  p <- res$plots$mean_per_10k_plot +                          
      scale_fill_manual(values = metacluster_colors) +
      scale_y_break(
        breaks = scales$breaks, 
        scales = scales$scales, 
        ticklabels = scales$ticklabels,
        expand = expansion(mult = c(0.01, 0.05))) +
      theme(
        text = element_text(face = "bold"),
        plot.title = element_text(size = 18, hjust = 0.4),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 16, vjust = 1.5),
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank()
      ) + NoLegend() + custom_theme 
  
  ggsave_multidevice(filename_base = paste0(dataset, "_MetaclusterCountsper10kBarplot"), 
         plot = p, width = 15, height = 6, dpi = 300)
  
 p <- res$plots$mean_per_10k_plot +                          
  scale_fill_manual(values = metacluster_colors) +
      scale_y_break(
        breaks = scales$breaks, 
        scales = scales$scales, 
        ticklabels = scales$ticklabels,
        expand = expansion(mult = c(0.01, 0.05))) +
   coord_flip() +
 theme(
    text = element_text(face = "bold"),
    plot.title = element_text(size = 18, hjust = 0.4),
    axis.text.x = element_text(hjust = 1, size = 14, angle = 45),
    axis.text.x.top = element_blank(),
    axis.ticks.x.top = element_blank(),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_text(size = 16, angle = 0),
    axis.title.y = element_blank(),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank()
  ) + 
   NoLegend() + custom_theme 
 
ggsave_multidevice(filename_base = paste0(dataset, "_MetaclusterCountsper10kBarplot_Vertical"),
                  plot = p, width = 8, height = 18)
}
```

```{r Plot averaged normalized metacluster counts per 10k cells}
# Plot the averaged cell count per 10k cells
results_list$YWV02_D30PW$plots$mean_per_10k_plot +                          
  scale_fill_manual(values = metacluster_colors) +
# scale_y_break(c(2000, 4500), scales = 0.1, ticklabels = c(5500), expand = expansion(mult = c(0.01, 0.05))) + # Edit values as needed
#  scale_y_continuous(limits = c(0,3500)) +
  theme(
    text = element_text(face = "bold"),
    plot.title = element_text(size = 18, hjust = 0.4),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 16),
    axis.title.y = element_text(size = 18, vjust = 1.5),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank()
  ) + NoLegend() + custom_theme 

  ggsave_multidevice(filename_base = paste0(dataset, "_MetaclusterCountsper10kBarplot"),
                   width = 12, height = 8)

# Plot the same plot, pivoted to vertical
result_object$plots$per_10k_plot +                          
  scale_fill_manual(values = metacluster_colors) +
#  scale_y_break(c(2000, 3000), scales = 0.1, ticklabels = c(3000, 3500)) + # Edit values as needed
#  scale_y_continuous(limits = c(0,3500)) +
  coord_flip() +
  theme(
    text = element_text(face = "bold"),
    plot.title = element_text(size = 18, hjust = 0.4),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    axis.title.y = element_blank(),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank()
  ) + NoLegend() + custom_theme 

ggsave_multidevice(filename_base = paste0(dataset, "_MetaclusterCounts10kBarplot_Vertical"),
                   width = 8, height = 12)

# Plot the averaged proportions as a stacked barplot
results_list$YWV02_D30PW$plots$proportion_plot +                          
  scale_fill_manual(values = metacluster_colors) +
  scale_x_discrete(expand = expansion(mult = 0.1)) +
  scale_y_continuous(expand = expansion(add = 0.1)) +
  coord_cartesian(clip = "off") +
  theme(
    text = element_text(face = "bold"),
    plot.title = element_text(size = 18, hjust = 0.4),
    axis.text.x = element_text(size = 16),
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 16),
    axis.title.y = element_text(size = 18),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank()
  ) + NoLegend() + custom_theme 
dggsave_multidevice(filename_base = paste0(dataset, "_MetaclusterProportions"),
                   width = 6, height = 12)
```
### Quantify cell type representation across anatomical sites
First we will quantify representation of immune cells in unwounded tissues
```{r Quantify immune cell representation across anatomical sites in UW}
final_df <- object_uw@meta.data %>%
  group_by(Region, AnatomicalSite, first_type) %>%
  summarise(Count = n(), .groups = "drop_last") %>%
  mutate(SiteTotal = sum(Count)) %>%
  ungroup() %>%
  group_by(Region) %>%
  mutate(RegionTotal = sum(Count)) %>%
  ungroup() %>%
  # Add total count across both regions for each cell type in an anatomical site
  group_by(AnatomicalSite, first_type) %>%
  mutate(TotalAcrossRegions = sum(Count)) %>%
  ungroup() %>%
  # Calculate proportions and normalized counts
  mutate(
    PropWithinSite     = Count / SiteTotal,
    PropWithinRegion   = Count / RegionTotal,
    CountPer1000Spots  = (Count / RegionTotal) * 1000
  )

immune_df <- final_df %>%
  filter(first_type %in% immune_types) %>%
  mutate(first_type = droplevels(first_type)) %>%  # remove non-immune levels
  group_by(AnatomicalSite, first_type) %>%
  summarise(CountPer1000Spots = sum(CountPer1000Spots), .groups = "drop") %>%
  complete(AnatomicalSite, first_type, fill = list(CountPer1000Spots = 0))  %>%
  # remove endothelial
  filter(AnatomicalSite != "Endothelial") %>%
  mutate(AnatomicalSite = factor(
    AnatomicalSite, 
    levels = rev(c("Epidermis","Dermis","HF","Adipose","Muscle","Fascia"))
  ))

shortnames <- c("Monocytes I" = "Mono I",
            "Monocytes II" = "Mono II",
            "Monocytes III" = "Mono III",
            "Macrophages I" = "Mac I",
            "Macrophages II" = "Mac II",
            "Macrophages III" = "Mac III",
            "Macrophages IV" = "Mac IV",
            "Cycling Macrophages" = "Cycling Mac",
            "OC-like cells" = "OC-like",
            "moDCs" = "moDCs",
            "cDC1" = "cDC1",
            "cDC2" = "cDC2",
            "Migratory DCs" = "migDCs",
            "Langerhans cells" = "LCs",
            "pDCs" = "pDCs",
            "Neutrophils I" = "Neu I",
            "Neutrophils II" = "Neu II",
            "Neutrophils III" = "Neu III",
            "Basophils" = "Baso",
            "Mast cells" = "Mast",
            "NK cells" = "NK",
            "γδT cells" = "γδT",
            "T cells" = "T cells",
            "Tregs" = "Tregs",
            "ILCs" = "ILCs",
            "B cells" = "B cells")
# Rename immune celltypes to shorter versions

immune_df <- immune_df %>%
  mutate(first_type = recode(first_type, !!!shortnames))

p <- ggplot(immune_df, aes(x = first_type, y = AnatomicalSite, fill = CountPer1000Spots)) +
  geom_tile(color = "white") +
  scale_fill_viridis(option = "E", transform = "log1p", direction = 1) +
  theme_minimal() +
  custom_theme +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8),
    axis.title = element_blank(),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 6),  # text size of legend labels
    legend.title = element_text(size = 8)
  ) +
  labs(
    fill = "Cells per 1000 spots (log1p)"
  ) 
p
ggsave_multidevice(filename_base = "YWV01_UW_AnatomicalSite_ImmuneHeatmap", plot = p,
                   width = 8, height = 2)
```
We can then take the other objects and use UW as a baseline to compare against
```{r}
# Parameters
ref_condition <- "YWV01_UW"  
plot_conditions <- setdiff(names(plot_info), ref_condition)
anatomical_levels <- rev(c("Epidermis","Dermis","HF","Adipose","Muscle","Fascia"))
facet_order <- c("YWV04_D4PW", "YWV03_D7PW", "YWV02_D30PW")

immune_list <- lapply(names(plot_info), function(cond) {
  obj <- plot_info[[cond]]$object
  
  obj@meta.data %>%
    filter(first_type %in% immune_types) %>%
    mutate(first_type = droplevels(first_type)) %>%
    group_by(Region, AnatomicalSite, first_type) %>%
    summarise(Count = n(), .groups = "drop") %>%
    mutate(RegionTotal = sum(Count),
           CountPer1000Spots = (Count / RegionTotal) * 1000) %>%  # keep original scale
    group_by(AnatomicalSite, first_type) %>%
    summarise(CountPer1000Spots = sum(CountPer1000Spots), .groups = "drop") %>%
    complete(AnatomicalSite, first_type, fill = list(CountPer1000Spots = 0)) %>%
    filter(AnatomicalSite %in% anatomical_levels) %>%
    filter(AnatomicalSite != "Endothelial") %>%
    mutate(
      AnatomicalSite = factor(AnatomicalSite, levels = anatomical_levels),
      first_type = recode(first_type, !!!shortnames),
      Condition = cond
    )
})

immune_df_all <- bind_rows(immune_list)

baseline_df <- immune_df_all %>% 
  filter(Condition == ref_condition) %>%
  select(AnatomicalSite, first_type, Baseline = CountPer1000Spots)

immune_fc_A <- immune_df_all %>%
  filter(Condition %in% plot_conditions) %>%
  left_join(baseline_df, by = c("AnatomicalSite", "first_type")) %>%
  mutate(FoldChange = (CountPer1000Spots + 1) / (Baseline + 1),
         FoldChange = pmin(pmax(FoldChange, 0.1), 10),
         Condition = factor(Condition, levels = facet_order))

p_A <- ggplot(immune_fc_A, aes(x = first_type, y = AnatomicalSite, fill = FoldChange)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 1, limits = c(0.1, 10),
                       name = "Fold change vs UW") +
  facet_wrap(~Condition, ncol = 1) +
  theme_minimal() +
  custom_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        legend.key.size = unit(0.5, "cm"),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8))

immune_fc_B <- immune_fc_A %>%
  group_by(Condition, AnatomicalSite) %>%
  mutate(PropWithinSite = CountPer1000Spots / sum(CountPer1000Spots)) %>%
  left_join(
    baseline_df %>%
      group_by(AnatomicalSite) %>%
      mutate(BaselineProp = Baseline / sum(Baseline)) %>%
      select(AnatomicalSite, first_type, BaselineProp),
    by = c("AnatomicalSite", "first_type")
  ) %>%
  mutate(FoldChange = (PropWithinSite + 1e-3) / (BaselineProp + 1e-3),
         FoldChange = pmin(pmax(FoldChange, 0.1), 10))

p_B <- ggplot(immune_fc_B, aes(x = first_type, y = AnatomicalSite, fill = FoldChange)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 1, limits = c(0.1, 10),
                       name = "Fold change vs UW") +
  facet_wrap(~Condition, ncol = 1) +
  theme_minimal() +
  custom_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        legend.key.size = unit(0.5, "cm"),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8))

# Display
p_A  # Absolute abundance fold change
p_B  # Within-anatomical site composition fold change

# Save
ggsave_multidevice("ST_Immune_Abundance_FCHeatmap", plot = p_A, width = 8, height = 6)
ggsave_multidevice("ST_Immune_Composition_FCHeatmap", plot = p_B, width = 8, height = 6)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# ----- Generate summarized data (same as your baseline code) -----
immune_list <- lapply(names(plot_info), function(cond) {
  obj <- plot_info[[cond]]$object
  
  obj@meta.data %>%
    filter(first_type %in% immune_types) %>%
    mutate(first_type = droplevels(first_type)) %>%
    group_by(Region, AnatomicalSite, first_type) %>%
    summarise(Count = n(), .groups = "drop") %>%
    mutate(RegionTotal = sum(Count),
           CountPer1000Spots = (Count / RegionTotal) * 1000) %>%
    group_by(AnatomicalSite, first_type) %>%
    summarise(CountPer1000Spots = sum(CountPer1000Spots), .groups = "drop") %>%
    complete(AnatomicalSite, first_type, fill = list(CountPer1000Spots = 0)) %>%
    filter(AnatomicalSite %in% anatomical_levels, AnatomicalSite != "Endothelial") %>%
    mutate(
      AnatomicalSite = factor(AnatomicalSite, levels = anatomical_levels),
      first_type = recode(first_type, !!!shortnames),
      Condition = cond
    )
})
immune_df_all <- bind_rows(immune_list)

# ----- Compute baseline and fold-change -----
baseline_df <- immune_df_all %>%
  filter(Condition == ref_condition) %>%
  select(AnatomicalSite, first_type, Baseline = CountPer1000Spots)

immune_fc <- immune_df_all %>%
  filter(Condition %in% plot_conditions) %>%
  left_join(baseline_df, by = c("AnatomicalSite", "first_type")) %>%
  mutate(FoldChange = (CountPer1000Spots + 1) / (Baseline + 1),
         FoldChange = pmin(pmax(FoldChange, 0.1), 10),
         Condition = factor(Condition, levels = facet_order))

# ----- Statistical testing (Wilcoxon) -----
test_df <- immune_df_all %>%
  filter(Condition %in% c(ref_condition, plot_conditions)) %>%
  group_by(AnatomicalSite, first_type, Condition) %>%
  summarise(values = CountPer1000Spots, .groups = "drop") %>%  # remove list()
  group_by(AnatomicalSite, first_type) %>%
  summarise(
    stats = list({
      conds <- split(values, Condition)
      lapply(setdiff(names(conds), ref_condition), function(tp) {
        if(length(conds[[tp]]) > 0 && length(conds[[ref_condition]]) > 0){
          p <- wilcox.test(conds[[tp]], conds[[ref_condition]])$p.value
        } else { p <- NA }
        data.frame(Condition=tp, p_val=p)
      }) %>% bind_rows()
    }), .groups="drop"
  ) %>%
  unnest(stats) %>%
  group_by(AnatomicalSite, first_type) %>%
  mutate(
    p_adj = p.adjust(p_val, method = "BH"),
    sig = case_when(
      p_adj <= 0.001 ~ "***",
      p_adj <= 0.01  ~ "**",
      p_adj <= 0.05  ~ "*",
      TRUE ~ ""
    )
  )

# Join p-values with FC data
immune_fc <- immune_fc %>%
  left_join(test_df %>% select(AnatomicalSite, first_type, Condition, sig),
            by = c("AnatomicalSite","first_type","Condition"))

# ----- Option A Plot -----
p_A <- ggplot(immune_fc, aes(x = first_type, y = AnatomicalSite, fill = FoldChange)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sig), color = "black", size = 2.5) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 1, limits = c(0.1, 10),
                       name = "Fold change vs UW") +
  facet_wrap(~Condition, ncol = 1) +
  theme_minimal() + custom_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank())

# ----- Option B -----
immune_fc_B <- immune_fc %>%
  group_by(Condition, AnatomicalSite) %>%
  mutate(PropWithinSite = CountPer1000Spots / sum(CountPer1000Spots)) %>%
  left_join(
    baseline_df %>%
      group_by(AnatomicalSite) %>%
      mutate(BaselineProp = Baseline / sum(Baseline)) %>%
      select(AnatomicalSite, first_type, BaselineProp),
    by = c("AnatomicalSite", "first_type")
  ) %>%
  mutate(FoldChange = (PropWithinSite + 1e-3) / (BaselineProp + 1e-3),
         FoldChange = pmin(pmax(FoldChange, 0.1), 10))

p_B <- ggplot(immune_fc_B, aes(x = first_type, y = AnatomicalSite, fill = FoldChange)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sig), color = "black", size = 2.5) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 1, limits = c(0.1, 10),
                       name = "Fold change vs UW") +
  facet_wrap(~Condition, ncol = 1) +
  theme_minimal() + custom_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank())


p_A
p_B
```

## Highlight cell types of interest
### Highlight non-regenerated cell types
Now with our function defined, we can save plots for a specific cell type of interest.
For figures, we want to show that muscle and adipose tissue are not regenerated, 
so generate highlight plots showing just those populations.
```{r Save highlight plots of muscle and adipose tissue}
# For these plots, we need to add a color for Unselected cells to the palette
metacluster_colors <- c(metacluster_colors, Unselected = "white")

# We can test the function to ensure it looks good
p <- CustomSpatialHighlightPlot(
  seurat_obj = plot_info[["YWV03_D7PW"]]$object,
  region = "Region1",
  annotation_col = "metaclusters",
  cells_highlight = c("Muscle cells","Adipocytes"),
  pt.size.highlight = 1,
  pt.size.unselected = 0.5,
  stroke.size = 0.05,
  pal = metacluster_colors,
  custom_theme = custom_theme
) + NoLegend()
print(p)

# Run the function in a loop to save all regions and timepoints
for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  Idents(seurat_obj) <- "metaclusters"
  
  for (region in regions) {
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      annotation_col = "metaclusters",
      cells_highlight = c("Muscle cells","Adipocytes"),
      pt.size.highlight = 1,
      pt.size.unselected = 0.5,
      stroke.size = 0.05,
      pal = metacluster_colors,
      custom_theme = custom_theme
    ) +
      custom_theme +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MuscleAdiposeHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}
```
### Highlight immune cells
Highlight immune cells in each object using matched color palette to immunome atlas
```{r Define cell types to highlight and the color palette to use}
# Specify immune cell types
immune_types <- c("Monocytes I","Monocytes II","Monocytes III","Macrophages I",
                  "Macrophages II","Macrophages III","Macrophages IV",
                  "Cycling Macrophages","OC-like cells","moDCs","cDC1","cDC2",
                  "Migratory DCs","Langerhans cells","pDCs","Neutrophils I",
                  "Neutrophils II","Neutrophils III","Basophils","Mast cells",
                  "NK cells","γδT cells","T cells","Tregs","ILCs","B cells")

# Define palette for immune types
pal <- setNames(color("smoothrainbow")(length(immune_types), range = c(0.1, 0.9)),
                immune_types) 
# Add value for unselected cells, this is necessary for highlight plots
pal <- c(pal, Unselected = "white")
```

```{r Test custom highlight plot with immune cells}
p <- CustomSpatialHighlightPlot(
  seurat_obj = plot_info[["YWV03_D7PW"]]$object,
  region = "Region1",
  cells_highlight = immune_types,
  pt.size.highlight = 1,
  pt.size.unselected = 0.5,
  stroke.size = 0.05,
  pal = pal,
  custom_theme = custom_theme
) + NoLegend()
print(p)
```

```{r Loop to save immune highlight plots for each object}
for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  Idents(seurat_obj) <- "first_type"
  
  # Filter immune types present in object
  present_idents <- names(table(Idents(seurat_obj)))[table(Idents(seurat_obj)) > 0]
  valid_immune_types <- intersect(immune_types, present_idents)
  if (length(valid_immune_types) == 0) next
  
  for (region in regions) {
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      cells_highlight = valid_immune_types,
      annotation_col = "first_type",
      pt.size.highlight = 1,
      pt.size.unselected = 0.5,
      stroke.size = 0.05,
      pal = pal,
      custom_theme = custom_theme
    ) +
      ggtitle(paste(info_name, region, "Immune Highlight")) +
      custom_theme +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "ImmuneHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}

for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  Idents(seurat_obj) <- "first_type"
  
  # Filter immune types present in object
  present_idents <- names(table(Idents(seurat_obj)))[table(Idents(seurat_obj)) > 0]
  valid_immune_types <- intersect(immune_types, present_idents)
  if (length(valid_immune_types) == 0) next
  
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = ifelse(info_name %in% c("YWV01_UW","YWV02_D30PW"), "ROI","ROI1"),
      cells_highlight = valid_immune_types,
      annotation_col = "first_type",
      pt.size.highlight = 1.5,
      pt.size.unselected = 0,
      stroke.size = 0.05,
      pal = pal,
      custom_theme = custom_theme
    ) +
      ggtitle(paste(info_name, region, "Immune Highlight")) +
      custom_theme +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "ImmuneHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
}
```
```{r Save cell type highlight plot with legend}
Idents(object_d7) <- "first_type"
immune_cells <- CellsByIdentities(object_d7, idents = immune_types)
immune_cells <- immune_cells[immune_types]
SpatialDimPlot(object_d7, images = "Region1", cells.highlight = immune_cells, 
               cols.highlight = pal) & custom_theme

ggsave("YWV03_D7PW_ImmuneHighlightLegend.png", width = 9, height = 6)
```

```{r Plot quantification of immune cells across objects}
plot_data <- bind_rows(
  get_metacluster_counts(object_uw, annotation_col = "first_type", "Unwounded")$region_counts,
  get_metacluster_counts(object_d4, annotation_col = "first_type", "D4PW")$region_counts,
  get_metacluster_counts(object_d7, annotation_col = "first_type", "D7PW")$region_counts,
  get_metacluster_counts(object_d30, annotation_col = "first_type", "D30PW")$region_counts
)

filtered_data <- plot_data %>%
  filter(first_type %in% immune_types)

filtered_data$Object <- factor(filtered_data$Object, levels = c("Unwounded","D4PW","D7PW","D30PW"))
filtered_data$first_type <- factor(filtered_data$first_type,
                                   levels = immune_types)

ggplot(filtered_data, aes(x = Object, y = Normalized, fill = first_type)) +
  stat_summary(fun = mean, geom = "bar", position = position_dodge(width = 0.9)) +
  geom_jitter(
    aes(color = first_type),
    shape = 21,
    color = "black",
    position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0, dodge.width = 0.9),
    size = 2,
    alpha = 0.7
  ) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  labs(
    title = "Proportion of Cell Types by Region Over Time",
    x = "Time Point",
    y = "Normalized Spot Count",
    fill = "Cell Type"
  ) +
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_blank()) + custom_theme

ggsave("ImmuneCells_VisiumHDPropBarplot.png", width = 12, height = 6)
```
```{r Define function to run stats on immune cells compared to unwounded}
# Statistical testing function
run_stats_vs_unwounded <- function(data, value_col) {
  # Value column is either "Normalized" or "Proportion"
  
  # Get the unwounded data
  unwounded_data <- data %>% 
    filter(Object == "Unwounded")
    
  # Get the data for D4PW, D7PW, D30PW
  wounded_timepoints <- c("D4PW", "D7PW", "D30PW")
  
  # Prepare an empty list to store results
  results_list <- list()
  
  for (time_point in wounded_timepoints) {
    # Data for the current time point
    time_point_data <- data %>% 
      filter(Object == time_point)
      
    # Iterate over each cell type
    for (cell_type in unique(data$first_type)) {
      
      # Filter data for the current cell type
      unwounded_subset <- unwounded_data %>% 
        filter(first_type == cell_type) %>% 
        pull(!!sym(value_col))
        
      time_point_subset <- time_point_data %>% 
        filter(first_type == cell_type) %>% 
        pull(!!sym(value_col))
        
      # Check if both groups have enough data to run the test
      if (length(unwounded_subset) > 1 & length(time_point_subset) > 1) {
        # Perform Wilcoxon rank-sum test
        test_result <- wilcox.test(time_point_subset, unwounded_subset, alternative = "two.sided")
        
        # Calculate the log2 fold change (D4PW/Unwounded)
        log2fc <- log2(mean(time_point_subset) / mean(unwounded_subset))
        
        # Store results
        results_list[[paste(time_point, cell_type, value_col, sep = "_")]] <- data.frame(
          Comparison = paste0(time_point, " vs Unwounded"),
          Cell_Type = cell_type,
          Metric = value_col,
          Log2FC = log2fc,
          P_value = test_result$p.value,
          W_statistic = test_result$statistic
        )
      } else {
        # Store results for cases with insufficient data
        results_list[[paste(time_point, cell_type, value_col, sep = "_")]] <- data.frame(
          Comparison = paste0(time_point, " vs Unwounded"),
          Cell_Type = cell_type,
          Metric = value_col,
          Log2FC = NA,
          P_value = NA,
          W_statistic = NA
        )
      }
    }
  }
  
  # Combine all results into a single data frame
  results_df <- bind_rows(results_list)
  
  # Apply multiple testing correction (e.g., Benjamini-Hochberg)
  results_df <- results_df %>%
    mutate(FDR = p.adjust(P_value, method = "BH")) %>%
    arrange(P_value)
    
  return(results_df)
}
```

```{r}
# --- R Code to Aggregate Total Immune Cell Counts ---

# Assuming 'counts_data' (from previous steps) contains all cell types, 
# and 'normalized_data' contains only immune subtypes.
# We will use 'normalized_data' as it already has the 'Normalized' count per region.

# 1. Aggregate the normalized counts across all immune subtypes (first_type) for each Region/Timepoint
total_immune_data <- normalized_data %>%
  # Group by the time point (Object) and the biological replicate (Region)
  group_by(Object, Region) %>%
  summarise(
    # Sum the Normalized counts (spots per 10k)
    Total_Normalized = sum(Normalized), 
    # Sum the Proportion counts (proportion of total spots)
    Total_Proportion = sum(Proportion), 
    .groups = "drop" # Drop the grouping
  ) %>%
  # Add a generic cell type label for plotting consistency
  mutate(Cell_Type = "Total Immune Cells")

# --- Plot A: Total Immune Normalized Count ---

# Calculate mean values for plotting the bar height
plot_means_norm <- total_immune_data %>%
  group_by(Object) %>%
  summarise(Mean = mean(Total_Normalized), .groups = 'drop')

total_norm_plot <- ggplot(total_immune_data, aes(x = Object, y = Total_Normalized)) +
  # Bar represents the mean value for the timepoint
  geom_col(
    data = plot_means_norm, 
    aes(x = Object, y = Mean), 
    fill = "skyblue", 
    color = "black",
    width = 0.7
  ) +
  # Jittered points represent the N=2 individual region values
  geom_jitter(
    width = 0.1, 
    shape = 21, 
    fill = "darkblue", 
    color = "black", 
    size = 3, 
    alpha = 0.8
  ) +
  theme_bw() +
  labs(
    title = "Total Immune Cell Normalized Count Over Time",
    x = "Time Point",
    y = "Total Immune Spots (Normalized to 10k Spots)"
  ) +
  theme(axis.text.x = element_text(size = 12),
        axis.title.x = element_blank()) + custom_theme
        
ggsave("TotalImmuneCells_NormalizedCount.png", plot = total_norm_plot, width = 6, height = 5)

# --- Plot B: Total Immune Proportion Out of All Spots ---

# Calculate mean values for plotting the bar height
plot_means_prop <- total_immune_data %>%
  group_by(Object) %>%
  summarise(Mean = mean(Total_Proportion), .groups = 'drop')

total_prop_plot <- ggplot(total_immune_data, aes(x = Object, y = Total_Proportion)) +
  # Bar represents the mean value for the timepoint
  geom_col(
    data = plot_means_prop, 
    aes(x = Object, y = Mean), 
    fill = "coral", 
    color = "black",
    width = 0.7
  ) +
  # Jittered points represent the N=2 individual region values
  geom_jitter(
    width = 0.1, 
    shape = 21, 
    fill = "darkred", 
    color = "black", 
    size = 3, 
    alpha = 0.8
  ) +
  theme_bw() +
  labs(
    title = "Total Immune Cell Proportion of All Spots Over Time",
    x = "Time Point",
    y = "Total Immune Proportion (%)"
  ) +
  # Ensure the Y-axis starts at 0 for proportion plots
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme(axis.text.x = element_text(size = 12),
        axis.title.x = element_blank()) + custom_theme
        
ggsave("TotalImmuneCells_Proportion.png", plot = total_prop_plot, width = 6, height = 5)
```

```{r Quantify immune cells over time}
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 3 Spatial/Quantification")
# --- Load Libraries ---
library(Seurat)
library(tidyverse)
library(RColorBrewer)
library(ggsignif) 

# --- Define Quantification Functions (Unchanged) ---

# Function 1: Calculates metacluster counts/proportions relative to ALL cells
get_metacluster_counts <- function(seurat_obj, annotation_col, label, scaling_factor = 10000) {
  # Count raw metacluster values by Region
  count_data <- seurat_obj@meta.data %>%
    dplyr::count(Region, !!sym(annotation_col)) %>%
    group_by(Region) %>%
    mutate(
      Normalized = n / sum(n) * scaling_factor, # Normalized to total cells
      Proportion = n / sum(n) * 100             # Proportion relative to total cells (%)
    ) %>%
    ungroup() %>%
    mutate(Object = label)
    
  return(list(region_counts = count_data))
}

# Function 2: Calculates proportions relative to ALL IMMUNE cells
get_immune_proportions <- function(seurat_obj, annotation_col, immune_types, label) {
  
  # 1. Filter Seurat object metadata to include only immune cells
  immune_meta <- seurat_obj@meta.data %>%
    filter(!!sym(annotation_col) %in% immune_types)
  
  # 2. Count immune cell subtypes by Region
  immune_count_data <- immune_meta %>%
    dplyr::count(Region, !!sym(annotation_col)) %>%
    group_by(Region) %>%
    # Calculate proportion relative to the total number of immune cells in that Region
    mutate(
      Proportion_Immune = n / sum(n) * 100
    ) %>%
    ungroup() %>%
    mutate(Object = label)
    
  return(immune_count_data)
}

# --- Define Statistical Functions (Optimized for Wilcoxon Test) ---

# Helper function to convert p-values to significance stars
get_stars <- function(p) {
  stars <- case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ "ns"
  )
  return(stars)
}

# Function 3: Run stats for a specified metric (Normalized, Proportion, or Proportion_Immune) vs. Unwounded
# The Wilcoxon Rank-Sum test is appropriate for comparing independent group means/medians.
run_wilcox_test_vs_unwounded <- function(data, value_col) {
  
  unwounded_data <- data %>% filter(Object == "Unwounded")
  wounded_timepoints <- c("D4PW", "D7PW", "D30PW")
  results_list <- list()
  
  for (time_point in wounded_timepoints) {
    time_point_data <- data %>% filter(Object == time_point)
      
    for (cell_type in unique(data$first_type)) {
      
      unwounded_subset <- unwounded_data %>% 
        filter(first_type == cell_type) %>% pull(!!sym(value_col))
        
      time_point_subset <- time_point_data %>% 
        filter(first_type == cell_type) %>% pull(!!sym(value_col))
        
      if (length(unwounded_subset) > 1 & length(time_point_subset) > 1) {
        # Perform Wilcoxon Rank-Sum test (non-parametric comparison of medians/distributions)
        test_result <- wilcox.test(time_point_subset, unwounded_subset, alternative = "two.sided")
        log2fc <- log2(mean(time_point_subset) / mean(unwounded_subset))
        
        results_list[[paste(time_point, cell_type, value_col, sep = "_")]] <- data.frame(
          Comparison = paste0(time_point, " vs Unwounded"),
          Cell_Type = cell_type,
          Metric = value_col,
          Log2FC = log2fc,
          P_value = test_result$p.value
        )
      }
    }
  }
  
  results_df <- bind_rows(results_list)
  if (nrow(results_df) > 0) {
      results_df <- results_df %>%
        mutate(FDR = p.adjust(P_value, method = "BH")) %>%
        mutate(Stars = get_stars(FDR)) %>%
        arrange(P_value)
  }
  return(results_df)
}

# --- Plotting Function with Stacked/Dodged Logic and Stats Integration ---

# Function 4: Generates the plot and adds significance stars
generate_plot_with_stats <- function(plot_data, stats_data, value_col, y_label, title, filename, plot_type = "dodged") {
  
  # Calculate mean data for plotting stars' Y position
  plot_means <- plot_data %>%
    group_by(Object, first_type) %>%
    summarise(Mean_Value = mean(!!sym(value_col)), .groups = 'drop')
    
  # Prepare stats data for plotting (only significant comparisons)
  star_positions <- stats_data %>%
    filter(Stars != "ns") %>%
    separate(Comparison, into = c("End_Group", "Start_Group"), sep = " vs ", remove = FALSE) %>%
    # Join with mean values for Y position calculation
    left_join(plot_means, by = c("End_Group" = "Object", "Cell_Type" = "first_type")) %>%
    group_by(Cell_Type) %>%
    # Stagger Y positions for multiple stars on the same cell type
    mutate(
      # Use max value of the entire plot data for consistent staggering height
      max_plot_value = max(plot_data[[value_col]]),
      Label_Y = max_plot_value + 0.1 * max_plot_value + (row_number() * 0.05 * max_plot_value) 
    ) %>%
    ungroup()
    
  # Create the base plot
  p <- ggplot(plot_data, aes(x = Object, y = !!sym(value_col), fill = first_type)) +
    theme_bw() +
    scale_fill_manual(values = pal) +
    labs(
      title = title,
      x = "Time Point",
      y = y_label,
      fill = "Cell Type"
    ) +
    theme(axis.text.x = element_text(size = 14),
          axis.text.y = element_text(size = 12),
          axis.title.x = element_blank()) + custom_theme

  # Apply plot type logic
  if (plot_type == "dodged") {
      # Dodged bars for counts (easier to read individual type changes)
      p <- p +
        stat_summary(fun = mean, geom = "bar", position = position_dodge(width = 0.9)) +
        geom_jitter(
          shape = 21, color = "black", 
          position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0, dodge.width = 0.9),
          size = 2, alpha = 0.7
        )
      
      # Add stars for dodged plot (need to dodge them too)
      if (nrow(star_positions) > 0) {
        p <- p +
          geom_text(
            data = star_positions,
            aes(x = End_Group, y = Label_Y, label = Stars, group = Cell_Type),
            position = position_dodge(width = 0.9), 
            size = 5,
            inherit.aes = FALSE 
          )
      }
      
  } else if (plot_type == "stacked") {
      # Stacked bars for proportions (to show 100% composition)
      p <- p +
        stat_summary(fun = mean, geom = "bar", position = "stack", color = "black") +
        # Note: Jittering raw data on a stacked bar plot can be misleading/difficult. 
        # We will use the bar plot only for composition visualization.
        # The stats stars apply to the overall proportions, which is the key comparison.
        
        # Add stars for stacked plot (no dodging needed)
        if (nrow(star_positions) > 0) {
          p <- p +
            geom_text(
              data = star_positions,
              aes(x = End_Group, y = Label_Y, label = Stars, group = Cell_Type),
              position = position_dodge(width = 0.9), # Using dodge(0.9) to prevent overlap on x-axis
              size = 5,
              inherit.aes = FALSE 
            )
        }
  }

  ggsave(filename, plot = p, width = 12, height = 6)
  return(p)
}

# --- EXECUTION: Data Processing & Analysis ---

# --- A. Prepare Data ---

# Normalized Data (relative to total cells)
counts_data <- bind_rows(
  get_metacluster_counts(object_uw, annotation_col = "first_type", "Unwounded")$region_counts,
  get_metacluster_counts(object_d4, annotation_col = "first_type", "D4PW")$region_counts,
  get_metacluster_counts(object_d7, annotation_col = "first_type", "D7PW")$region_counts,
  get_metacluster_counts(object_d30, annotation_col = "first_type", "D30PW")$region_counts
)
normalized_data <- counts_data %>% filter(first_type %in% immune_types)
normalized_data$Object <- factor(normalized_data$Object, levels = c("Unwounded","D4PW","D7PW","D30PW"))
normalized_data$first_type <- factor(normalized_data$first_type, levels = immune_types)

# Immune Proportion Data (relative to total immune cells)
immune_proportion_data <- bind_rows(
  get_immune_proportions(object_uw, annotation_col = "first_type", immune_types, "Unwounded"),
  get_immune_proportions(object_d4, annotation_col = "first_type", immune_types, "D4PW"),
  get_immune_proportions(object_d7, annotation_col = "first_type", immune_types, "D7PW"),
  get_immune_proportions(object_d30, annotation_col = "first_type", immune_types, "D30PW")
)
immune_proportion_data$Object <- factor(immune_proportion_data$Object, levels = c("Unwounded","D4PW","D7PW","D30PW"))
immune_proportion_data$first_type <- factor(immune_proportion_data$first_type, levels = immune_types)

# --- 1. Statistical Testing (Wilcoxon Rank-Sum Test) and CSV Saving ---

# 1.1. Stats for Normalized Counts
normalized_counts_stats <- run_wilcox_test_vs_unwounded(normalized_data, "Normalized")
write.csv(normalized_counts_stats, "ImmuneCell_NormalizedCounts_Stats_vs_Unwounded.csv", row.names = FALSE)
# [File: ImmuneCell_NormalizedCounts_Stats_vs_Unwounded.csv]

# 1.2. Stats for Immune Proportions
immune_proportions_stats <- run_wilcox_test_vs_unwounded(immune_proportion_data, "Proportion_Immune")
write.csv(immune_proportions_stats, "ImmuneCell_ImmuneProportion_Stats_vs_Unwounded.csv", row.names = FALSE)
# [File: ImmuneCell_ImmuneProportion_Stats_vs_Unwounded.csv]


# --- 2. Plot Generation with Stats Stars ---

# 2.1. Plot for Normalized Counts (DODGED)
normalized_counts_plot <- generate_plot_with_stats(
  plot_data = normalized_data,
  stats_data = normalized_counts_stats,
  value_col = "Normalized",
  y_label = "Normalized Spot Count (per 10,000 spots)",
  title = "Immune Cell Normalized Spot Counts Over Time",
  filename = "ImmuneCells_NormalizedCounts_Dodged_WithStats.png",
  plot_type = "dodged"
)

# 2.2. Plot for Immune Proportions (STACKED)
immune_proportions_plot <- generate_plot_with_stats(
  plot_data = immune_proportion_data,
  stats_data = immune_proportions_stats,
  value_col = "Proportion_Immune",
  y_label = "Proportion of Total Immune Cells (%)",
  title = "Immune Cell Subtype Composition (100% Immune)",
  filename = "ImmuneCell_ImmuneProportion_Stacked_WithStats.png",
  plot_type = "stacked"
)

cat("Analysis complete. Both plots generated (one dodged, one stacked).\n")
```
```{r Plot FC Relative to UW}
# --- Load Libraries ---
library(tidyverse)
library(RColorBrewer) 

# Assuming the following functions and objects are already defined and loaded from previous steps:
# - get_metacluster_counts, get_immune_proportions
# - run_wilcox_test_vs_unwounded (calculates Log2FC)
# - normalized_data, normalized_counts_stats (from previous execution)
# - immune_types, pal, custom_theme
# - The Seurat objects (object_uw, etc.)

# --- Dedicated Log2FC Plotting Function ---

plot_log2fc_relative_to_unwounded <- function(stats_data, title, filename, show_stats = TRUE) {
  
  # Ensure the data is filtered and ordered correctly
  plot_data <- stats_data %>%
    filter(!is.na(Log2FC)) %>%
    mutate(
      Cell_Type = factor(Cell_Type, levels = rev(unique(Cell_Type))), # Reverse order for horizontal plot
      Time_Point = gsub(" vs Unwounded", "", Comparison)
    )
    
  # Find the maximum absolute value for symmetric x-axis limits
  max_fc <- max(abs(plot_data$Log2FC)) * 1.1 
  
  log2fc_plot <- ggplot(plot_data, aes(x = Log2FC, y = Cell_Type, fill = Time_Point)) +
    # Use geom_col for bar chart
    geom_col(position = position_dodge(width = 0.8), color = "black", size = 0.3) +
    
    # Add a vertical line at Log2FC = 0 (no change)
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
    
    scale_x_continuous(limits = c(-max_fc, max_fc)) +
    
    labs(
      title = title,
      x = expression(log[2]~"Fold Change vs. Unwounded"),
      y = "Immune Cell Type",
      fill = "Time Point"
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      axis.title.y = element_blank(),
      axis.text.x = element_text(size = 10),
      legend.position = "bottom"
    ) + custom_theme
    
    # --- NEW LOGIC TO PLOT SIGNIFICANCE LABELS CONDITIONALLY ---
    if (show_stats) {
      log2fc_plot <- log2fc_plot +
        geom_text(
          aes(label = Stars),
          # Position the stars to the right of the bar, avoiding overlap with the bar itself
          # The sign(Log2FC) ensures the star is placed outward from the zero line
          hjust = ifelse(plot_data$Log2FC < 0, 1.1, -0.1), 
          position = position_dodge(width = 0.8),
          size = 3.5,
          color = "black"
        )
    }
    # --- END NEW LOGIC ---
    
  ggsave(filename, plot = log2fc_plot, width = 10, height = 8)
  
  return(log2fc_plot)
}

# --- EXECUTION: Generate Log2FC Plot ---

# 1. Ensure normalized_counts_stats has been run (from previous code chunk)
# If running this independently, you must run the statistical analysis first:
# normalized_counts_stats <- run_wilcox_test_vs_unwounded(normalized_data, "Normalized")

# 2. Generate the plot
log2fc_plot <- plot_log2fc_relative_to_unwounded(
  stats_data = normalized_counts_stats,
  title = "Log2 Fold Change in Normalized Cell Counts Relative to Unwounded",
  filename = "ImmuneCells_NormalizedCounts_Log2FC_Plot.png",
  show_stats = FALSE
)

cat("Generated Log2FC plot: ImmuneCells_NormalizedCounts_Log2FC_Plot.png\n")
```

Save plots of specific immune subclusters
```{r Save plots of specific immune populations}
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 3 Spatial/Immune cell highlight plots")

immune_cells_highlight <- c("Macrophages I","Macrophages II","Macrophages IV","OC-like cells")

for (cond in names(plot_info)) {
  plot_celltypes(
    seurat_obj = plot_info[[cond]]$object,
    cells_highlight = immune_cells_highlight,
    annotation_col = "first_type",
    regions = plot_info[[cond]]$regions,
    pt.size.factor = plot_info[[cond]]$pt.size.factor,
    palette = pal
  )
}
```
Save plots of multiple cell types simultaneously
```{r Colabel immune cells - Mac I and III}
immune_types <- c("Macrophages I","Macrophages III")

# # Define palette for immune types
# pal_fun <- color("smoothrainbow")
# pal <- pal_fun(length(immune_types), range = c(0.1, 0.9))
# names(pal) <- immune_types
# pal <- c(pal, Unselected = "white")

# Loop through each condition/sample
for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  pt.size <- info$pt.size.factor

  Idents(seurat_obj) <- "first_type"

  # Filter immune types to those actually present in the object
  present_idents <- names(table(Idents(seurat_obj)))[table(Idents(seurat_obj)) > 0]
  valid_immune_types <- intersect(immune_types, present_idents)

  # Skip if no immune types are present
  if (length(valid_immune_types) == 0) next

  # Identify immune cells
  immune_cells <- CellsByIdentities(seurat_obj, idents = valid_immune_types)

  # Skip if no cells were found
  if (length(immune_cells) == 0) next

  for (region in regions) {
   # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      cells_highlight = valid_immune_types,
      annotation_col = "first_type",
      pt.size.highlight = 1.2,
      pt.size.unselected = 0.5,
      stroke.size = 0.05,
      pal = pal,
      custom_theme = custom_theme
    ) +
      custom_theme +
      theme(
        panel.grid = element_blank(),
        plot.title = element_blank()
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MacI-III-Highlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}
```
```{r Colabel immune cells - Mac II and IV}
immune_types <- c("Macrophages II","Macrophages IV")

# # Define palette for immune types
# pal_fun <- color("smoothrainbow")
# pal <- pal_fun(length(immune_types), range = c(0.1, 0.9))
# names(pal) <- immune_types
# pal <- c(pal, Unselected = "white")

# Loop through each condition/sample
for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  pt.size <- info$pt.size.factor

  Idents(seurat_obj) <- "first_type"

  # Filter immune types to those actually present in the object
  present_idents <- names(table(Idents(seurat_obj)))[table(Idents(seurat_obj)) > 0]
  valid_immune_types <- intersect(immune_types, present_idents)

  # Skip if no immune types are present
  if (length(valid_immune_types) == 0) next

  # Identify immune cells
  immune_cells <- CellsByIdentities(seurat_obj, idents = valid_immune_types)

  # Skip if no cells were found
  if (length(immune_cells) == 0) next

  for (region in regions) {
   # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      cells_highlight = valid_immune_types,
      annotation_col = "first_type",
      pt.size.highlight = 1,
      pt.size.unselected = 0.5,
      stroke.size = 0.05,
      pal = pal,
      custom_theme = custom_theme
    ) +
      custom_theme +
      theme(
        panel.grid = element_blank(),
        plot.title = element_blank()
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MacII-IV-Highlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}
```
```{r Save cell type highlight plot with legend}
Idents(object_d7) <- "first_type"
immune_cells <- CellsByIdentities(object_d7, idents = immune_types)
immune_cells <- immune_cells[immune_types]
SpatialDimPlot(object_d7, images = "Region1", cells.highlight = immune_cells, 
               cols.highlight = pal) & custom_theme

ggsave("YWV03_D7PW_MacHighlightLegend.png", width = 9, height = 6)
```
```{r Define a function to plot individual cell types for a seurat object}
# Define function, using custom plotting function from above, to save plots
plot_celltypes <- function(seurat_obj, 
                           cells_highlight,
                           annotation_col,
                           regions, 
                           pt.size.factor = 1, 
                           palette, 
                           stroke.size = 0.05,
                           plot_height = 7,
                           plot_width = 7) {
  # Set identities to cell type annotation
  Idents(seurat_obj) <- annotation_col
  
  for (region in regions) {
    # Loop through each cell type in cells_highlight
    for (cell_type in cells_highlight) {

      # Generate plot with the custom function
      p <- CustomSpatialHighlightPlot(
        seurat_obj = seurat_obj,
        region = region,
        annotation_col = annotation_col,
        cells_highlight = cell_type,  # highlight one cell type at a time
        pt.size.highlight = pt.size.factor * 1,
        pt.size.unselected = pt.size.factor * 0.25,
        stroke.size = stroke.size,
        pal = palette,
        custom_theme = custom_theme
      ) +
        custom_theme +
        theme(panel.grid = element_blank()) +
        NoLegend()
      
      # Save plot
      ggsave_multidevice(
        filename_base = paste(unique(seurat_obj$orig.ident), region, cell_type, "HighlightPlot", sep = "_"),
        plot = p,
        width = plot_width,
        height = plot_height
      )
    }
  }
}
```
Save plots of each cell type individually 
```{r Save individual highlights of macrophages I-IV}
mac_types <- c("Macrophages I","Macrophages II","Macrophages III","Macrophages IV")

setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Users/Simon/Spatial Data/Immune Cell highlight plots")

for (cond in names(plot_info)) {
  plot_celltypes(
    seurat_obj = plot_info[[cond]]$object,
    cells_highlight = mac_types,
    annotation_col = "first_type",
    regions = plot_info[[cond]]$regions,
    pt.size.factor = plot_info[[cond]]$pt.size.factor,
    palette = pal
  )
}
```
```{r Save plot of all macrophage populations}
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Users/Simon/Spatial Data/Immune Cell highlight plots")
for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  Idents(seurat_obj) <- "first_type"
  
  for (region in regions) {
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      cells_highlight = mac_types,
      annotation_col = "first_type",
      pt.size.highlight = 1,
      pt.size.unselected = 0.25,
      stroke.size = 0.05,
      pal = pal,
      custom_theme = custom_theme
    ) +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MacrophageHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}

```

Quantify immune cell representation across wound sites in the data
```{r Quantify immune representation across wound sites}
# Initialize an empty list to store results
all_results <- list()

for (cond in names(plot_info)) {
  obj <- plot_info[[cond]]$object
  
  # Quantify counts per region for immune cells
  results_df <- obj@meta.data %>%
    filter(first_type %in% immune_types) %>%
    count(Region, WoundSite, first_type, name = "Count") %>%
    mutate(Condition = cond)  # add condition info
  
  # Compute mean counts, SEM, and proportions
  mean_df <- results_df %>%
    group_by(WoundSite, first_type, Condition) %>%
    summarise(
      MeanCount = mean(Count, na.rm = TRUE),
      SEM       = sd(Count, na.rm = TRUE) / sqrt(n()),   # standard error
      TotalCount = sum(Count, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    group_by(WoundSite, Condition) %>%
    mutate(
      Prop = TotalCount / sum(TotalCount)
    ) %>%
    ungroup()
  
  all_results[[cond]] <- mean_df
}

# Combine into one dataframe
immune_quant_df <- bind_rows(all_results)

# Set levels for plotting
## Timepoint
immune_quant_df$Condition <- factor(immune_quant_df$Condition, 
                                    levels = c("YWV01_UW","YWV04_D4PW","YWV03_D7PW","YWV02_D30PW"))
## Wound site
immune_quant_df$WoundSite <- factor(immune_quant_df$WoundSite, 
                                    levels = c("Periphery","WoundEdge","Wound"))

for(site in unique(immune_quant_df$WoundSite)){
    # Filter to wound site of interest
    periphery_df <- immune_quant_df %>%
      filter(WoundSite == site)
    
    # --- Counts over time ---
    ggplot(periphery_df, aes(x = Condition, y = MeanCount, fill = first_type)) +
      geom_bar(stat = "identity", position = "dodge", color = "grey25") +
      scale_fill_manual(values = pal) +
      labs(title = paste0("Immune Cell Counts in ",site," Over Time"),
           y = "Cell Count",
           fill = "Cell Type") +
      ggprism::theme_prism() +
      theme(axis.text.x = element_text(),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(size = 14),
            plot.title = element_text(hjust = 0.5)) + 
      custom_theme + NoLegend()
    ggsave_multidevice(filename_base = paste0("ImmuneCounts_",site,"_BarPlot"), width = 10, height = 6)
    
    # --- Proportions over time ---
    ggplot(periphery_df, aes(x = Condition, y = Prop, fill = first_type)) +
      geom_bar(stat = "identity", position = "stack") +
      scale_fill_manual(values = pal) +
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
      labs(title =  paste0("Immune Cell Proportions in ",site," Over Time"),
           x = "Condition",
           y = "Proportion of Immune Cells",
           fill = "Cell Type") +
      ggprism::theme_prism() +
      theme(axis.text.x = element_text(),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(size = 14),
            plot.title = element_text(hjust = 0.5)) + 
      custom_theme + NoLegend()
    ggsave_multidevice(filename_base = paste0("ImmuneProps_",site,"_BarPlot"), width = 10, height = 6)
}

for(selected_timepoint in names(plot_info)){

    timepoint_df <- immune_quant_df %>%
      filter(Condition == selected_timepoint)
    
    # --- Counts across wound sites ---
    ggplot(timepoint_df, aes(x = WoundSite, y = MeanCount, fill = first_type)) +
      geom_bar(stat = "identity", position = "dodge", color = "grey25") +
      scale_fill_manual(values = pal) +
      labs(title = paste("Immune Cell Counts -", selected_timepoint),
           x = "Wound Site",
           y = "Cell Count",
           fill = "Cell Type") +
      ggprism::theme_prism() +
      theme(axis.text.x = element_text(),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(size = 14),
            plot.title = element_text(hjust = 0.5)) + 
      custom_theme + NoLegend()
    ggsave_multidevice(filename_base = paste0("ImmuneWoundSiteCounts_",selected_timepoint,"_BarPlot"), width = 10, height = 6)
    
    # --- Proportions across wound sites ---
    ggplot(timepoint_df, aes(x = WoundSite, y = Prop, fill = first_type)) +
      geom_bar(stat = "identity", position = "stack") +
      scale_fill_manual(values = pal) +
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
      labs(title = paste("Immune Cell Proportions -", selected_timepoint),
           x = "Wound Site",
           y = "Proportion of Immune Cells",
           fill = "Cell Type") +
      ggprism::theme_prism() + 
      theme(axis.text.x = element_text(),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(size = 14),
            plot.title = element_text(hjust = 0.5)) + 
      custom_theme + NoLegend()
    ggsave_multidevice(filename_base = paste0("ImmuneWoundSiteProps_",selected_timepoint,"_BarPlot"), width = 10, height = 6)
}
```
Also save a plot for just unwounded which is formatted to match the scRNAseq
```{r Save UW plot with matched formatting to scRNAseq}
timepoint_df <- immune_quant_df %>%
      filter(Condition == "YWV01_UW")
    
    # --- Counts across wound sites ---
    ggplot(timepoint_df, aes(x = first_type, y = MeanCount, fill = first_type)) +
      geom_bar(stat = "identity", position = "dodge", width = 0.6) +
      geom_errorbar(
        aes(ymin = MeanCount - SEM, ymax = MeanCount + SEM),
        width = 0.2,
        position = position_dodge(width = 0.6)
        ) +
      scale_fill_manual(values = pal) +
      labs(title = paste("Unwounded Immune Cell Counts from ST"),
           x = "Wound Site",
           y = "Cell Count",
           fill = "Cell Type") +
      ggprism::theme_prism(axis_text_angle = 45) +
      theme(
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 12),
            axis.title.y = element_text(size = 14),
            plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank()) + 
      custom_theme + NoLegend()
    
ggsave_multidevice(filename_base = paste0("YWV01_UW_ImmuneCounts_BarPlot"), width = 9, height = 6)
```
Lastly, in conjunction with the immune analysis, we can examine the localization
of various populations associated with the three modes of immune recruitment
First, save vectors of immune cells for each dynamic
```{r Save immune cell lists by recruitment mode}
recruitment <- c("Monocytes I","Monocytes II","Monocytes III","Macrophages I","Macrophages III",
                 "OC-like cells","moDCs","Neutrophils I","Neutrophils II","Basophils")

restoration <-c("cDC1","cDC2","pDCs","Neutrophils III","Mast cells","NK cells","T cells","B cells")

neostasis <- c("Macrophages II","Macrophages IV","Cycling Macrophages",
               "Migratory DCs","Langerhans cells","γδT cells","Tregs","ILCs")

# Save each vector to a list
recruitment_modes <- list("Recruitment" = recruitment,
                          "Restoration" = restoration,
                          "Neostasis" = neostasis)
```

```{r Save plots of immune cell types grouped by temporal dynamics}
# Iterate over the list of modes to save plots
for(mode in names(recruitment_modes)){
    # Create a new folder for output files
    new.dir <- paste0(getwd(),"/",mode," Highlight plots")
    dir.create(new.dir)
    setwd(new.dir)
    
    for (info_name in names(plot_info)) {
      info <- plot_info[[info_name]]
      seurat_obj <- info$object
      regions <- info$regions
      
      Idents(seurat_obj) <- "first_type"
      
      for (region in regions) {
        # Generate the plot using the custom function
        p <- CustomSpatialHighlightPlot(
          seurat_obj = seurat_obj,
          region = region,
          cells_highlight = recruitment_modes[[mode]],
          annotation_col = "first_type",
          pt.size.highlight = 1,
          pt.size.unselected = 0.5,
          stroke.size = 0.05,
          pal = pal,
          custom_theme = custom_theme
        ) +
          theme(
            panel.grid = element_blank(),
            plot.title = element_text(hjust = 0.5)
          ) +
          NoLegend()
        
        # Save plot
        ggsave_multidevice(
          filename_base = paste(info_name, region, mode, "Highlight", sep = "_"),
          plot = p,
          width = 7,
          height = 7
        )
      }
      rm(seurat_obj)
    }
    
    # Save one plot with the legend
    p <- CustomSpatialHighlightPlot(
      seurat_obj = plot_info[["YWV04_D4PW"]]$object,
      region = "Region1",
      cells_highlight = recruitment_modes[[mode]],
      annotation_col = "first_type",
      pt.size.highlight = 1.25,
      pt.size.unselected = 0.5,
      stroke.size = 0.05,
      pal = pal,
      custom_theme = custom_theme
    ) 
    ggsave(paste0("YWV04_D4PW_",mode,"_Legend.png"), plot = p)
}
```

```{r Save immune modes with single color}
# Phase palette you provided
phase.pal <- c(
  "Recruitment" = "#F8766D",
  "Restoration" = "#00BA38",
  "Neostasis"   = "#619CFF"
)

# Loop over all modes
for (mode in names(recruitment_modes)) {

  # Make output folder
  out.dir <- file.path(getwd(), paste0(mode, " Highlight plots"))
  dir.create(out.dir, showWarnings = FALSE)

  # Build color mapping: every highlighted cell type gets SAME color
  highlight_colors <- setNames(
    rep(phase.pal[mode], length(recruitment_modes[[mode]])),
    recruitment_modes[[mode]]
  )

  # Loop over datasets
  for (info_name in names(plot_info)) {

    info <- plot_info[[info_name]]
    seurat_obj <- info$object
    regions <- info$regions

    Idents(seurat_obj) <- "first_type"

    # Loop over regions
    for (region in regions) {

      p <- CustomSpatialHighlightPlot(
        seurat_obj = seurat_obj,
        region = region,
        cells_highlight = recruitment_modes[[mode]],
        annotation_col = "first_type",
        pt.size.highlight = 1,
        pt.size.unselected = 0.5,
        stroke.size = 0.05,
        pal = c(highlight_colors, Unselected = "white"),    
        custom_theme = custom_theme
      ) +
        theme(
          panel.grid = element_blank(),
          plot.title = element_text(hjust = 0.5)
        ) +
        NoLegend()

      # Save per-region plot without legend
      ggsave_multidevice(
        filename_base = paste(info_name, region, mode, "Highlight", sep = "_"),
        plot = p,
        width = 7,
        height = 7
      )
    }

    rm(seurat_obj)
  }

  # One plot WITH legend (fixed dataset + Region1)
  legend_plot <- CustomSpatialHighlightPlot(
    seurat_obj = plot_info[["YWV04_D4PW"]]$object,
    region = "Region1",
    cells_highlight = recruitment_modes[[mode]],
    annotation_col = "first_type",
    pt.size.highlight = 1.25,
    pt.size.unselected = 0.5,
    stroke.size = 0.05,
    pal =  c(highlight_colors, Unselected = "white"),
    custom_theme = custom_theme
  )

  ggsave(
    filename = file.path(out.dir, paste0("YWV04_D4PW_", mode, "_Legend.png")),
    plot = legend_plot,
    width = 7,
    height = 7
  )
}

```

```{r Save highlight plots of all immune modes colored by mode}
# Phase palette you provided
# Combine all highlighted cell types
all_highlight <- unlist(recruitment_modes)

# Build per-cell color map
highlight_colors <- unlist(lapply(names(recruitment_modes), function(mode){
  setNames(
    rep(phase.pal[mode], length(recruitment_modes[[mode]])),
    recruitment_modes[[mode]]
  )
}))

# Save list of images
dataset_regions <- list(
  YWV01_UW    = c("uw_Region1",  "uw_Region2"),
  YWV02_D30PW = c("d30_Region1", "d30_Region2"),
  YWV03_D7PW  = c("d7_Region1",  "d7_Region2"),
  YWV04_D4PW  = c("d4_Region1",  "d4_Region2")
)

# Loop to save combined-mode plots
for(dataset in names(dataset_regions)){

  regions <- dataset_regions[[dataset]]

  for(region in regions){

    p <- CustomSpatialHighlightPlot(
      seurat_obj = merged_st,
      dataset = dataset,
      region = region,
      annotation_col = "first_type",
      cells_highlight = all_highlight,
      pal = c(highlight_colors, Unselected = "white"),
      stroke.size = 0.01,
      pt.size.highlight = 1.5,
      pt.size.unselected = 0.25,
      custom_theme = custom_theme,
      show_scalebar = F
    ) +
      NoLegend()

    ggsave_multidevice(
      filename_base = paste0("ST_", dataset, "_", region, "_TemporalModes"),
      plot = p,
      height = 8,
      width = 8
    )
  }
}

```

With the groupings, we can then ask where these populations primarily localize in
relation to wound zones
```{r Quantify cell representation by temporal dynamics and wound zone}
# Build lookup table from recruitment_modes list
mode_lookup <- enframe(recruitment_modes, name = "Mode", value = "CellTypes") %>%
  unnest(CellTypes) %>%
  rename("CellTypes" = "first_type")

# Initialize an empty list to store results
all_results <- list()

for (cond in names(plot_info)) {
  obj <- plot_info[[cond]]$object

  # Quantify counts per region for immune cells
  results_df <- obj@meta.data %>%
    filter(first_type %in% immune_types) %>%
    count(Region, WoundSite, first_type, name = "Count") %>%
    mutate(Condition = cond)

  # Compute mean + SEM across Regions
  mean_df <- results_df %>%
    group_by(Condition, WoundSite, first_type) %>%
    summarise(
      MeanCount = mean(Count, na.rm = TRUE),
      SEM       = sd(Count, na.rm = TRUE) / sqrt(n()),
      .groups   = "drop"
    )

  all_results[[cond]] <- mean_df
}

# Combine into one dataframe
immune_quant_df <- bind_rows(all_results)

# Add mode annotations
immune_quant_modes <- immune_quant_df %>%
  left_join(mode_lookup, by = "first_type") %>%
  filter(!is.na(Mode))

# Collapse to Mode level: sum of mean counts across cell types in that Mode
mode_summary <- immune_quant_modes %>%
  group_by(Condition, WoundSite, Mode) %>%
  summarise(
    MeanCount = sum(MeanCount, na.rm = TRUE),
    SEM       = sqrt(sum(SEM^2, na.rm = TRUE)),  # combine SEMs conservatively
    .groups = "drop"
  )

# Normalize to proportions within each site if desired
mode_summary <- mode_summary %>%
  group_by(Condition, WoundSite) %>%
  mutate(Prop = MeanCount / sum(MeanCount)) %>%
  ungroup()

# Set factor levels for plotting
mode_summary$Condition <- factor(
  mode_summary$Condition,
  levels = c("YWV01_UW","YWV04_D4PW","YWV03_D7PW","YWV02_D30PW")
)
mode_summary$Mode <- factor(
  mode_summary$Mode,
  levels = c("Recruitment","Restoration","Neostasis")
)
mode_summary$WoundSite <- factor(
  mode_summary$WoundSite,
  levels = c("Wound","WoundEdge","Periphery")
)

# --- Plot using mean counts (bars or alluvial) ---
ggplot(mode_summary,
  aes(x = Condition, stratum = Mode, alluvium = Mode, y = MeanCount, fill = Mode)) +
  ggalluvial::geom_flow(stat = "alluvium") +
  ggalluvial::geom_stratum() +
  facet_wrap(~WoundSite) +
  ggprism::theme_prism() +
  labs(x = "Timepoint",
       y = "Mean count across regions")

ggplot(mode_summary,
  aes(x = Condition, stratum = Mode, alluvium = Mode, y = Prop, fill = Mode)) +
  ggalluvial::geom_flow(stat = "alluvium") +
  ggalluvial::geom_stratum(width = 2/3) +
  facet_wrap(~WoundSite) +
  ggprism::theme_prism() +
  labs(x = "Timepoint",
       y = "Proportion")
ggsave_multidevice(filename_base = "ST_ImmuneModes_WoundSitePropBarplot", width = 8, height = 4)

ggplot(mode_summary, aes(x = as.numeric(factor(Condition)), y = Prop, color = Mode)) +
  geom_smooth(se = FALSE, method = "loess") +
  facet_wrap(~WoundSite) +
  ggprism::theme_prism() +
  labs(x = "Timepoint",
       y = "Proportion") +
  custom_theme
ggsave_multidevice(filename_base = "ST_ImmuneModes_WoundSitePropLinegraph", width = 8, height = 4)
```
### Emergent cell types
Now we can highlight cell types that we identified in OWHA and wish to validate
```{r Define cell types of interest}
human_celltypes <- c("Outer Bulge III","Myofibroblasts IV", "Lox+ Fibroblasts", 
               "Vein ECs")

major_celltypes <- c("Reticular III","Papillary II","Myofibroblasts II","Basal IV","Spinous I","HF I")

minor_celltypes <- c("Fascia","Proliferating ECs","Muscle Progenitors","Repair SC II","Pericyte I","Adipocyte II")

cells_to_plot <- c("Spinous I","Proliferating ECs","Pericyte I","HF I")


cells_to_plot <- c("Basal IV")
```
We can quantify the proportions of cell types of interest across multiple objects
```{r Quantify specific cell type representation}
plot_data <- bind_rows(
  get_metacluster_counts(object_uw, annotation_col = "first_type", "Unwounded")$region_counts,
  get_metacluster_counts(object_d4, annotation_col = "first_type", "D4PW")$region_counts,
  get_metacluster_counts(object_d7, annotation_col = "first_type", "D7PW")$region_counts,
  get_metacluster_counts(object_d30, annotation_col = "first_type", "D30PW")$region_counts
)

filtered_data <- plot_data %>%
  filter(first_type %in% human_celltypes)

filtered_data$Object <- factor(filtered_data$Object, levels = c("Unwounded","D4PW","D7PW","D30PW"))
filtered_data$first_type <- factor(filtered_data$first_type,
                                   levels = c("Lox+ Fibroblasts", "Myofibroblasts IV","Outer Bulge III", "Vein ECs"))

ggplot(filtered_data, aes(x = Object, y = Proportion, fill = first_type)) +
  stat_summary(fun = mean, geom = "bar", position = position_dodge(width = 0.9)) +
  geom_jitter(
    aes(color = first_type),
    shape = 21,
    color = "black",
    position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0, dodge.width = 0.9),
    size = 2,
    alpha = 0.7
  ) +
  scale_fill_manual(values = cell_colors) +
  theme_bw() +
  labs(
    title = "Proportion of Cell Types by Region Over Time",
    x = "Time Point",
    y = "Proportion of Spots (%)",
    fill = "Cell Type"
  ) +
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_blank()) + custom_theme

ggsave("HumanCellTypes_VisiumHDPropBarplot.png", width = 9, height = 6)
```
Save plots of emergent cell types using the custom highlighting function
```{r Save plots of major cell types}
# Add color for unselected cells to palette
cell_colors <- c(cell_colors, Unselected = "white")

for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  Idents(seurat_obj) <- "first_type"
  
  for (region in regions) {
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      cells_highlight = major_celltypes,
      annotation_col = "first_type",
      pt.size.highlight = 1,
      pt.size.unselected = 0.25,
      stroke.size = 0.05,
      pal = cell_colors,
      custom_theme = custom_theme
    ) +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MajorCelltypesHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}

# Save one plot with the legend
p <- CustomSpatialHighlightPlot(
  seurat_obj = plot_info[["YWV04_D4PW"]]$object,
  region = "Region1",
  cells_highlight = major_celltypes,
  annotation_col = "first_type",
  pt.size.highlight = 1.25,
  pt.size.unselected = 0.5,
  stroke.size = 0.05,
  pal = cell_colors,
  custom_theme = custom_theme
) 
ggsave_multidevice("YWV04_D4PW_MajorCellTypes_Legend", plot = p)
```

```{r Save plots of minor cell types}
# Add color for unselected cells to palette
cell_colors <- c(cell_colors, Unselected = "white")

for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  Idents(seurat_obj) <- "first_type"
  
  for (region in regions) {
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      cells_highlight = minor_celltypes,
      annotation_col = "first_type",
      pt.size.highlight = 1,
      pt.size.unselected = 0.25,
      stroke.size = 0.05,
      pal = cell_colors,
      custom_theme = custom_theme
    ) +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MinorCelltypesHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}

# Save one plot with the legend
p <- CustomSpatialHighlightPlot(
  seurat_obj = plot_info[["YWV04_D4PW"]]$object,
  region = "Region1",
  cells_highlight = minor_celltypes,
  annotation_col = "first_type",
  pt.size.highlight = 1.25,
  pt.size.unselected = 0.5,
  stroke.size = 0.05,
  pal = cell_colors,
  custom_theme = custom_theme
) 
ggsave_multidevice("YWV04_D4PW_MinorCellTypes_Legend", plot = p)
```
We can also combine the lists and save them together
```{r Save major and minor celltypes in one plot}
# Add color for unselected cells to palette
cell_colors <- c(cell_colors, Unselected = "white")

for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  Idents(seurat_obj) <- "first_type"
  
  for (region in regions) {
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      region = region,
      cells_highlight = c(major_celltypes, minor_celltypes),
      annotation_col = "first_type",
      pt.size.highlight = 1,
      pt.size.unselected = 0.25,
      stroke.size = 0.05,
      pal = cell_colors,
      custom_theme = custom_theme
    ) +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MajorMinorCelltypesHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}

# Save one plot with the legend
p <- CustomSpatialHighlightPlot(
  seurat_obj = plot_info[["YWV04_D4PW"]]$object,
  region = "Region1",
  cells_highlight = c(major_celltypes, minor_celltypes),
  annotation_col = "first_type",
  pt.size.highlight = 1.25,
  pt.size.unselected = 0.5,
  stroke.size = 0.05,
  pal = cell_colors,
  custom_theme = custom_theme
) 
ggsave_multidevice("YWV04_D4PW_MajorMinorCellTypes_Legend", plot = p)
```

We can also supply each of our cell type lists to save plots for each.
```{r Iterate over objects to save individual highlight plots of major and minor cell types}
# Major cell types
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Shared Data/Spatial Data/Figures/Major Cell Types")
for (cond in names(plot_info)) {
  plot_celltypes(
    seurat_obj = plot_info[[cond]]$object,
    cells_highlight = major_celltypes,
    annotation_col = "first_type",
    regions = plot_info[[cond]]$regions,
    pt.size.factor = plot_info[[cond]]$pt.size.factor,
  palette = cell_colors
  )
}
# Minor cell types
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Shared Data/Spatial Data/Figures/Minor Cell Types")
for (cond in names(plot_info)) {
  plot_celltypes(
    seurat_obj = plot_info[[cond]]$object,
    cells_highlight = minor_celltypes,
    annotation_col = "first_type",
    regions = plot_info[[cond]]$regions,
    pt.size.factor = plot_info[[cond]]$pt.size.factor,
  palette = cell_colors
  )
}
```
```{r Save D7PW WoundEdge ROI plots for specific cell types}
plot_celltypes(
    seurat_obj =object_d7,
    cells_highlight = cells_to_plot,
    annotation_col = "first_type",
    regions = "WoundEdgeROI",
    pt.size.factor = 6,
    stroke.size = 0.1,
    palette = cell_colors,
    plot_height = 7.5,
    plot_width = 7
  )

p <- CustomSpatialHighlightPlot(
  seurat_obj = object_d7,
  region = "WoundEdgeROI",
  cells_highlight =  c(major_celltypes, minor_celltypes),
  annotation_col = "first_type",
  pt.size.highlight = 6,
  pt.size.unselected = 1.5,
  stroke.size = 0.1,
  pal = cell_colors,
  custom_theme = custom_theme
) 
ggsave_multidevice(filename_base = "YWV03_D7PW_MajorMinorCellTypes_WoundEdgeROI_Legend", plot = p)

p + NoLegend()
ggsave_multidevice(filename_base = "YWV03_D7PW_MajorMinorCellTypes_WoundEdgeROI", height = 7.5)

cluster_pal <- setNames(viridis(length(unique(object_d7$banksy_cluster_0.5)), option = "H"),levels(object_d7$banksy_cluster_0.5))

p <- CustomSpatialHighlightPlot(
  seurat_obj = object_d7,
  region = "WoundEdgeROI",
  cells_highlight =  unique(object_d7$banksy_cluster_0.5),
  annotation_col = "banksy_cluster_0.5",
  pt.size.highlight = 2.5,
  pt.size.unselected = 1.5,
  stroke.size = 0.05,
  pal = cluster_pal,
  custom_theme = custom_theme
) + NoLegend()
p

ggsave_multidevice(filename_base = "YWV03_D7PW_BanksyClusters_WoundEdgeROI", height = 7.5)

SpatialDimPlot(
  object = object_d7,
  images = "WoundEdgeROI",
  group.by = "banksy_cluster_0.5",
  pt.size.factor = 6,
  cols = cluster_pal
) & custom_theme + NoLegend()
```

```{r}
source("~/Documents/Shared-Utils/Useful Functions/CustomSpatialHighlightPlot2.R")
```

```{r Save colabel plots of fascia and macrophages}
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Users/Simon/Spatial Data/Immune Cell highlight plots/Macrophage Fascia Colabeling")

cols <- c(cell_colors["Fascia"], pal)

cells_to_plot <- c("Fascia","Macrophages I","Macrophages II","Macrophages III",
                   "Macrophages IV")

for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  dataset <- unique(seurat_obj$orig.ident)
  
  Idents(seurat_obj) <- "first_type"
  
  for (region in regions) {
    # Generate the plot using the custom function
    p <- CustomSpatialHighlightPlot(
      seurat_obj = seurat_obj,
      dataset = dataset,
      region = region,
      cells_highlight = cells_to_plot,
      annotation_col = "first_type",
      pt.size.highlight = 1,
      pt.size.unselected = 0.25,
      stroke.size = 0.05,
      pal = cols,
      custom_theme = custom_theme
    ) +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) +
      NoLegend()
    
    # Save plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "MacFasciaHighlight", sep = "_"),
      plot = p,
      width = 7,
      height = 7
    )
  }
}

# # Save one plot with the legend
# p <- CustomSpatialHighlightPlot(
#   seurat_obj = plot_info[["YWV04_D4PW"]]$object,
#   region = "Region1",
#   cells_highlight = major_celltypes,
#   annotation_col = "first_type",
#   pt.size.highlight = 1.25,
#   pt.size.unselected = 0.5,
#   stroke.size = 0.05,
#   pal = cell_colors,
#   custom_theme = custom_theme
# ) 
# ggsave_multidevice("YWV04_D4PW_MajorCellTypes_Legend", plot = p)
```

We can also use the function to save individual plots of metaclusters for each object
```{r Save individual plots of each metacluster}
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Hair Follicle Paper/Shared Data/Spatial Data/Figures/Cell Type Annotation/Individual Metacluster highlight plots")

for (cond in names(plot_info)) {
  plot_celltypes(
    seurat_obj = plot_info[[cond]]$object,
    cells_highlight = names(metacluster_colors),
    annotation_col = "metaclusters",
    regions = plot_info[[cond]]$regions,
    pt.size.factor = plot_info[[cond]]$pt.size.factor,
    palette = metacluster_colors
  )
}
```

```{r Load custom feature plot for coexpression}
source("~/Documents/Shared-Utils/Useful Functions/CustomSpatialFeatureCoexpPlot.R")
```

```{r Use custom feature plot to plot multiple features at once for OC-like inducing genes}
feats <- c("Tnfsf11","Tnfrsf11a","Tnfrsf11b","Mmp3")
feat_colors <- setNames(c("#D53E4F","#3288BD","#66C2A5","#FDAE61"),feats) 

p_uw <- CustomSpatialFeatureCoexpPlot(
  object = object_uw,
  image = "Region2",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 0.75,
  pt.size.zero = 0.25,
  bg.col       = "grey95"
) & custom_theme 
p_uw

p_d4 <- CustomSpatialFeatureCoexpPlot(
  object = object_d4,
  image = "Region1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 0.75,
  pt.size.zero = 0.25,
  bg.col       = "grey95"
) & custom_theme 
p_d4

p_d7 <- CustomSpatialFeatureCoexpPlot(
  object = object_d7,
  image = "Region1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 0.75,
  pt.size.zero = 0.25,
  bg.col       = "grey95"
) & custom_theme 
p_d7

p_d30 <- CustomSpatialFeatureCoexpPlot(
  object = object_d30,
  image = "Region1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 0.75,
  pt.size.zero = 0.25,
  bg.col       = "grey95"
)
p_d30

ggsave("YWV01_UW_OClikeInducersFeatPlot.png", plot = p_uw, width = 7, height = 8)
ggsave("YWV02_D30PW_OClikeInducersFeatPlot.png", plot = p_d30, width = 7, height = 8)
ggsave("YWV04_D4PW_OClikeInducersFeatPlot.png", plot = p_d4, width = 7, height = 8)
ggsave("YWV03_D7PW_OClikeInducersFeatPlot.png", plot = p_d7, width = 7, height = 8)
```
We can then loop through the objects to save more plots, this time of a region of
interest
```{r Loop to save OC-inducing genes for ROI}
for(cond in names(plot_info)){
  p <- CustomSpatialFeatureCoexpPlot(
    object = plot_info[[cond]]$object,
    image = ifelse(cond %in% c("YWV01_UW","YWV02_D30PW"),"ROI","ROI1"),
    genes = feats,
    gene_colors = feat_colors,
    assay = "Spatial",
    pt.size.expr = 1.5,
    pt.size.zero = 0,
    bg.col       = "grey99"
  ) & custom_theme 
  ggsave_multidevice(filename_base = paste0(cond,"_OCInducerFeatplot_ROI"),
                     width = 7, height = 8)
}

# We can also loop over the features to save each one individually
for(cond in names(plot_info)){
  for(feat in feats){
  p <- CustomSpatialFeatureCoexpPlot(
    object = plot_info[[cond]]$object,
    image = ifelse(cond %in% c("YWV01_UW","YWV02_D30PW"),"ROI","ROI1"),
    genes = feat,
    gene_colors = feat_colors[feat],
    assay = "Spatial",
    pt.size.expr = 1.5,
    pt.size.zero = 0.5,
    bg.col       = "grey99"
  ) & custom_theme 
  ggsave_multidevice(filename_base = paste0(cond,"_",feat,"Featplot_ROI"),
                     width = 7, height = 8)
  }
}
```

```{r Plot expression of OC-inducing genes on dotplot}
feats <- c("Tnfsf11","Tnfrsf11a","Tnfrsf11b","Mmp3")

DotPlot(object_uw, feats, assay = "Spatial", group.by = "first_type") + 
  theme(axis.title = element_blank()) + custom_theme
ggsave("YWV01_UW_OCInducerDotPlot.png", width = 8,height = 18)

DotPlot(object_d4, feats, assay = "Spatial", group.by = "first_type")+ 
  theme(axis.title = element_blank()) + custom_theme
ggsave("YWV04_D4PW_OCInducerDotPlot.png", width = 8,height = 18)

DotPlot(object_d7, feats, assay = "Spatial", group.by = "first_type")+ 
  theme(axis.title = element_blank()) + custom_theme
ggsave("YWV03_D7PW_OCInducerDotPlot.png", width = 8,height = 18)

DotPlot(object_d30, feats, assay = "Spatial", group.by = "first_type")+ 
  theme(axis.title = element_blank()) + custom_theme
ggsave("YWV02_D30PW_OCInducerDotPlot.png", width = 8,height = 18)
```

```{r Use custom feature plot to plot Mac II markers}

feats <- c("Igf1")

feat_colors <- c("purple")
p_uw <- CustomSpatialFeatureCoexpPlot(
  object = object_uw,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_uw

SpatialFeaturePlot(object_uw, images = "Region2",features = feats, blend = T)

p_d4 <- CustomSpatialFeatureCoexpPlot(
  object = object_d4,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_d4

p_d7 <- CustomSpatialFeatureCoexpPlot(
  object = object_d7,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_d7

p_d30 <- CustomSpatialFeatureCoexpPlot(
  object = object_d30,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_d30

ggsave("YWV01_UW_OClikeInducersFeatPlot.png", plot = p_uw, width = 7, height = 8)
ggsave("YWV02_D30PW_OClikeInducersFeatPlot.png", plot = p_d30, width = 7, height = 8)
ggsave("YWV04_D4PW_OClikeInducersFeatPlot.png", plot = p_d4, width = 7, height = 8)
ggsave("YWV03_D7PW_OClikeInducersFeatPlot.png", plot = p_d7, width = 7, height = 8)
```

```{r Plot markers on FeaturePlot}
# "Ctsk","Mmp12",

feats <- c("Ccl8")

# feats <- c("Itgav","Itgb3")

DefaultAssay(object_uw) <- "Spatial"
DefaultAssay(object_d4) <- "Spatial"
DefaultAssay(object_d7) <- "Spatial"
DefaultAssay(object_d30) <- "Spatial"

low_color = "gray90"
high_color = "red3"

SpatialFeaturePlot(object_uw, images = "slice1", feats) & scale_fill_gradient(low = low_color, high = high_color)
SpatialFeaturePlot(object_d4, images = "slice1", feats) & scale_fill_gradient(low = low_color, high = high_color)
SpatialFeaturePlot(object_d7, images = "slice1", feats) & scale_fill_gradient(low = low_color, high = high_color)
SpatialFeaturePlot(object_d30, images = "slice1", feats) & scale_fill_gradient(low = low_color, high = high_color)
```
```{r}
DotPlot(object_uw, "Igf1", group.by = "first_type")
DotPlot(object_d4, "Igf1", group.by = "first_type")
DotPlot(object_d7, "Ccl8", group.by = "first_type")
DotPlot(object_d30, "Igf1", group.by = "first_type")
```


```{r Loop to save individual feature plots of various markers}
markers <- c("Arg1","Nos2","Vegfa","Acp5","Mrc1","C1qa","Cd163","Igf1","Retnla","Spp1")

# Set default assay for all objects
plot_info <- lapply(plot_info, function(obj) {
  DefaultAssay(obj$object) <- "Spatial"
  return(obj)
})

# Loop through conditions, regions, and markers
for (cond in names(plot_info)) {
  obj <- plot_info[[cond]]$object
  regions <- plot_info[[cond]]$regions
  pt.size <- plot_info[[cond]]$pt.size.factor
  
  for (reg in regions) {
    for (marker in markers) {
      p <- SpatialFeaturePlot(
        object = obj,
        features = marker,
        images = reg,
        pt.size.factor = pt.size * 1.1
      ) +
        ggtitle(paste(cond, reg, marker, sep = " - ")) +
        scale_fill_gradient(low = "gray90", high = "red3")
      
      print(p)
      
      # Optionally save
      ggsave(paste0("plots/", cond, "_", reg, "_", marker,".png"), plot = p, width = 7, height = 7, dpi = 300)
    }
  }
}
```

```{r Save function to plot feature expression with cell type highlight}
source("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Custom Functions/plot_spatial_overlay.R")
```

```{r Use custom feature/dimplot to plot expression in a given celltype}
# Define variables
genep <- "Igf1"
slice <- "Macrophages II"

# Generate spatial overlay plots
p1 <- plot_spatial_overlay(object_d7,  gene = genep, highlight_group = c(slice, "D7"))
p2 <- plot_spatial_overlay(object_d4,  gene = genep, highlight_group = c(slice, "D4"))
p3 <- plot_spatial_overlay(object_uw,  gene = genep, highlight_group = c(slice, "UW"))
p4 <- plot_spatial_overlay(object_d30, gene = genep, highlight_group = c(slice, "D30"))

p1
# Save individual plots
ggsave(filename = paste0("spatial_plot_", genep, "_D7_", slice, ".png"),
       plot = p1, width = 5, height = 5, dpi = 300, units = "in")

ggsave(filename = paste0("spatial_plot_", genep, "_D4_", slice, ".png"),
       plot = p2, width = 5, height = 5, dpi = 300, units = "in")

ggsave(filename = paste0("spatial_plot_", genep, "_UW_", slice, ".png"),
       plot = p3, width = 5, height = 5, dpi = 300, units = "in")

ggsave(filename = paste0("spatial_plot_", genep, "_D30_", slice, ".png"),
       plot = p4, width = 5, height = 5, dpi = 300, units = "in")

# Combine plots using patchwork
library(patchwork)
wrap_plot <- (p3 | p1) / (p2 | p4)

# Save combined plot
ggsave(filename = paste0("Spatial_", slice, "_plot_", genep, "_combined.png"),
       plot = wrap_plot, width = 10, height = 10, dpi = 300, units = "in")
```

```{r Plot coexpression of IGF1 in Macrophage II}
# Define variables
genep <- "Igf1"
slice <- "Macrophages II"

p1 <- plot_spatial_overlay(object_uw,  
                           gene = genep, 
                           highlight_group = c(slice, "UW"),
                           image = "Region2",
                           highlight_color = pal[slice],
                           high_color = "goldenrod1")

p1

ggsave_multidevice(filename_base = "YWV01_UW_MacII_Igf1_CoexpressionPlot", width = 8, height = 8)

p1 <- plot_spatial_overlay(object_uw,  
                           gene = genep, 
                           highlight_group = c(slice, "UW"),
                           image = "ROI",
                           pt.size.highlight = 1.5,
                           pt.size.unselected = 0.5,
                           highlight_color = pal[slice],
                           high_color = "goldenrod1")

p1 + NoLegend() + theme(plot.title = element_blank())
ggsave_multidevice(filename_base = "YWV01_UW_MacII_Igf1_CoexpressionPlot_ROI", width = 7, height = 7)

```
```{r}
# Define variables
genep <- "Thbs1"
slice <- "Macrophages I"

p1 <- plot_spatial_overlay(object_d4,  
                           gene = genep, 
                           highlight_group = c(slice, "D4"),
                           image = "Region1",
                           highlight_color = pal[slice],
                           high_color = "red3")

p1
```

```{r Add OxPhos module score}
celltype_kegg <- read_xlsx("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 1/IntegratedImmune_CellMarkers_KEGGResults.xlsx")

pathways <- geneInCategory(celltype_kegg)
# Pull gene IDs associated with OxPhos pathway
genes_mmu00190 <- unlist(lapply(pathways, function(cell) cell[["mmu00190"]]))
genes_mmu00190 <- unique(genes_mmu00190)  # remove duplicates if needed

# Convert IDs back to symbols
genes_mmu00190 <- bitr(genes_mmu00190,
                    fromType = "ENTREZID",
                    toType = "SYMBOL",
                    OrgDb = org.Mm.eg.db)

# Use the gene symbols to calculate a module score
for(cond in names(plot_info)){
  obj <- plot_info[[cond]]$object
  
  DefaultAssay(obj) <- "Spatial"
  
  obj <- AddModuleScore(obj, list(genes_mmu00190$SYMBOL), name = "OxPhos Score")
  
  plot_info[[cond]]$object <- obj
  
  rm(obj)
}

SpatialFeaturePlot(plot_info$YWV01_UW$object, "OxPhos Score1",images = "Region2",
                        min.cutoff = c("q5"))  & custom_theme
SpatialFeaturePlot(plot_info$YWV04_D4PW$object, "OxPhos Score1",images = "Region1",
                        min.cutoff = c("q5"))  & custom_theme
SpatialFeaturePlot(plot_info$YWV03_D7PW$object, "OxPhos Score1",images = "Region1",
                        min.cutoff = c("q5"))  & custom_theme
SpatialFeaturePlot(plot_info$YWV02_D30PW$object, "OxPhos Score1",images = "Region1",
                        min.cutoff = c("q5"))  & custom_theme

plot_info <- lapply(plot_info, function(info) {
  obj <- info$object
  Idents(obj) <- "first_type"
  info$object <- obj
  info
})

# Plot expression of OxPhos in immune cells
DotPlot(plot_info$YWV01_UW$object, "OxPhos Score1", idents = immune_types)
DotPlot(plot_info$YWV04_D4PW$object, "OxPhos Score1", idents = immune_types)
DotPlot(plot_info$YWV03_D7PW$object, "OxPhos Score1", idents = immune_types)
DotPlot(plot_info$YWV02_D30PW$object, "OxPhos Score1", idents = immune_types)
```
We can then use the spatial overlay plot to show different cell types
```{r}
genep <- "OxPhos Score1"
slice <- "OC-like cells"

p1 <- plot_spatial_overlay(plot_info$YWV03_D7PW$object,  
                           gene = genep, 
                           highlight_group = c(slice, "D4PW"),
                           image = "Region1",
                           highlight_color = pal[slice],
                           high_color = "goldenrod1",
                           min_expr = 0.25)

p1

```

## Wound Site specific analysis
```{r Add ROI Metadata}
for (cond in names(plot_info)) {
  obj <- plot_info[[cond]]$object
  
  roi_file <- file.path(
    "~/SRSP Laboratory Dropbox/SRSP Lab/Team/Immunodynamics/Simon Van Deursen/DATA/Spatial Transcriptomics",
    cond, paste0(cond, "_Realignment/ROIs.csv")
  )
  
  # 1. Skip if ROI file doesn't exist
  if (!file.exists(roi_file)) {
    message("Skipping ", cond, " — no ROI file.")
    next
  }
  
  roi.coordinates <- read.csv(roi_file)
  
  # 2. Skip if no barcodes in file
  if (nrow(roi.coordinates) == 0) {
    message("Skipping ", cond, " — ROI file empty.")
    next
  }
  
  roi.assignments <- setNames(roi.coordinates$ROIs, roi.coordinates$Barcode)
  
  # 3. Add metadata only for barcodes that exist in obj
  roi.assignments <- roi.assignments[names(roi.assignments) %in% colnames(obj)]
  if (length(roi.assignments) == 0) {
    message("Skipping ", cond, " — no matching barcodes in object.")
    next
  }
  
  obj <- AddMetaData(obj, roi.assignments, col.name = "ROI")
  
  # 4. Extract coordinates
  coords <- GetTissueCoordinates(obj)
  coords$ROI <- obj$ROI[rownames(coords)]
  
  # 5. Split by ROI — keep even if only 1 ROI present
  region_coords_list <- split(coords, coords$ROI)
  
  # 6. Loop over ROIs and make convex hull overlays
  for (region_name in names(region_coords_list)) {
    region_coords <- region_coords_list[[region_name]]
    
    # Need ≥3 points for convex hull
    if (nrow(region_coords) < 3) {
      message("Skipping ROI ", region_name, " — not enough points.")
      next
    }
    
    hull_indices <- chull(region_coords$x, region_coords$y)
    hull_coords <- region_coords[hull_indices, c("x", "y")]
    hull_coords <- rbind(hull_coords, hull_coords[1, ]) # close polygon
    hull_coords$cell <- region_name
    
    print(hull_coords)  # sanity check
    
    region_segmentation <- CreateSegmentation(hull_coords)
    
    # Show overlay — this won't overwrite the object
    obj[[region_name]] <- Overlay(obj[["slice1"]], region_segmentation)
  }
  
  # Save modified object back into plot_info
  plot_info[[cond]]$object <- obj
}

SpatialDimPlot(object = plot_info$YWV01_UW$object, 
               group.by = "first_type", pt.size.factor = 3, images = "ROI")

SpatialDimPlot(object = plot_info$YWV03_D7PW$object, 
               group.by = "banksy_cluster_0.5", pt.size.factor = 3, images = "ROI1")
```
```{r Add Wound Edge ROI to D7PW dataset}
cond <- "YWV03_D7PW"

roi_file <- "~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Data/Spatial Data/YWV03_D7PW/YWV03_D7PW_Realignment/WoundEdgeROI.csv"

roi.coordinates <- read.csv(roi_file)

roi.assignments <- setNames(roi.coordinates$WoundEdgeROI, roi.coordinates$Barcode)
  
# 3. Add metadata only for barcodes that exist in obj
  roi.assignments <- roi.assignments[names(roi.assignments) %in% colnames(object_d7)]
  if (length(roi.assignments) == 0) {
    message("Skipping ", cond, " — no matching barcodes in object.")
    next
  }
  
  object_d7 <- AddMetaData(object_d7, roi.assignments, col.name = "WoundEdgeROI")
  
  # 4. Extract coordinates
  coords <- GetTissueCoordinates(object_d7)
  coords$ROI <- object_d7$WoundEdgeROI[rownames(coords)]
  
  # 5. Split by ROI — keep even if only 1 ROI present
  region_coords_list <- split(coords, coords$ROI)
  
  # 6. Loop over ROIs and make convex hull overlays
  for (region_name in names(region_coords_list)) {
    region_coords <- region_coords_list[[region_name]]
    
    # Need ≥3 points for convex hull
    if (nrow(region_coords) < 3) {
      message("Skipping ROI ", region_name, " — not enough points.")
      next
    }
    
    hull_indices <- chull(region_coords$x, region_coords$y)
    hull_coords <- region_coords[hull_indices, c("x", "y")]
    hull_coords <- rbind(hull_coords, hull_coords[1, ]) # close polygon
    hull_coords$cell <- region_name
    
    print(hull_coords)  # sanity check
    
    region_segmentation <- CreateSegmentation(hull_coords)
    
    # Show overlay — this won't overwrite the object
    object_d7[[region_name]] <- Overlay(object_d7[["slice1"]], region_segmentation)
  }

SpatialDimPlot(object = object_d7, 
               group.by = "banksy_cluster_0.5", pt.size.factor = 3, images = "WoundEdgeROI")
```

Save plots of ROI clusters
```{r Loop to save plots of Wound site ROI}
# Define images and point sizes for each sample
plot_settings <- list(
  "YWV01_UW" = list(image = "ROI", pt_size = 1.5),
    "YWV04_D4PW" = list(image = "ROI1", pt_size = 1.5),
  "YWV03_D7PW" = list(image = "ROI1", pt_size = 1.5),
  "YWV02_D30PW" = list(image = "ROI", pt_size = 1.5)
)

for (cond in names(plot_info)) {
  if (!cond %in% names(plot_settings)) {
    message("Skipping ", cond, " (not in plot list).")
    next
  }
  
  obj <- plot_info[[cond]]$object
  cluster_col <- "banksy_cluster_0.5"
  
  # Get cluster levels
  if (is.factor(obj@meta.data[[cluster_col]])) {
    cluster_levels <- levels(obj@meta.data[[cluster_col]])
  } else {
    cluster_levels <- sort(unique(as.character(obj@meta.data[[cluster_col]])))
  }
  
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), cluster_levels)
  
  image_to_plot <- plot_settings[[cond]]$image
  pt_size <- plot_settings[[cond]]$pt_size
  
  p <- SpatialDimPlot(obj,
                      group.by = cluster_col,
                      pt.size.factor = pt_size,
                      images = image_to_plot,
                      cols = cluster_colors) & custom_theme +
      theme(panel.grid = element_blank())
  
  ggsave_multidevice(filename_base = paste0(cond, "_ROI_ClusterPlot"),
         plot = p,
         width = 8, height = 7)
  
   p <- SpatialDimPlot(obj,
                      group.by = cluster_col,
                      pt.size.factor = pt_size,
                      images = image_to_plot,
                      label = T,repel = T,label.size = 2,label.box = T,
                      cols = cluster_colors) & custom_theme +
      theme(panel.grid = element_blank())
  
  ggsave_multidevice(filename_base = paste0(cond, "_ROI_ClusterPlot_Labeled"),
         plot = p,
         width = 8, height = 7)
  
    p <- SpatialDimPlot(obj,
                      group.by = "metaclusters",
                      pt.size.factor = pt_size,
                      images = image_to_plot,
                      cols = metacluster_colors) + NoLegend()  & custom_theme +
      theme(panel.grid = element_blank())
  
  ggsave_multidevice(filename_base = paste0(cond, "_ROI_MetaclusterPlot"),
         plot = p,
         width = 7, height = 7)
}
```
Save plots of unwounded immune cells in the ROI for figure 1
```{r Unwounded ROI Immune highlight for Fig 1}
p <- CustomSpatialHighlightPlot(
  seurat_obj = plot_info[["YWV01_UW"]]$object,
  region = "ROI",
  annotation_col = "first_type",
  cells_highlight = immune_types,
  pt.size.highlight = 1.5,
  pt.size.unselected = 0.5,
  stroke.size = 0.05,
  pal = pal,
  custom_theme = custom_theme
) + NoLegend()
print(p)

ggsave_multidevice(filename_base = "YWV01_UW_ROI_ImmuneHighlight", width = 7, height = 7)

SpatialDimPlot( plot_info[["YWV01_UW"]]$object, group.by = "AnatomicalSite",
                images = "ROI", pt.size.factor = 2) & custom_theme +
      theme(panel.grid = element_blank()) + NoLegend()

ggsave_multidevice(filename_base = "YWV01_UW_ROI_AnatomicalRegions", width = 7, height = 7)

```
Double check the wound site annotations for each object
```{r Check Wound Site annotations}
unique(object_uw$WoundSite)
unique(object_d4$WoundSite)
unique(object_d7$WoundSite)
unique(object_d30$WoundSite)
```
Examining the unique annotations, we can see the D7 object has some spots missing 
annotations. We can plot where they are in the object and then add an annotation
```{r Examine spots in D7 missing wound site and add annotation}
Idents(object_d7) <- "WoundSite"
missing_cells <-  WhichCells(object_d7, idents = "")
# Plot wound sites
SpatialDimPlot(
      object_d7,
      images = "slice1"
    )
# Plot unannotated cells
SpatialDimPlot(
      object_d7,
      images = "slice1",
      cells.highlight = missing_cells
    )

# Based on the plots, we can see that the annotation for spots missing one depends on the region
object_d7$WoundSite[colnames(object_d7) %in% missing_cells & object_d7$Region == "Region1"] <- "Periphery"
object_d7$WoundSite[colnames(object_d7) %in% missing_cells & object_d7$Region == "Region2"] <- "WoundEdge"

Idents(object_d7) <- "WoundSite"
levels(object_d7) <- c("Wound","WoundEdge","Periphery")
object_d7$WoundSite <- Idents(object_d7)

SpatialDimPlot(
      object_d7,
      images = "slice1"
    )
```
Save DimPlots of woundsite for each object
```{r Save WoundSite annotation Dimplots}
woundsite_pal <- c("Wound" = "#7E1900",
                   "WoundEdge" = "#E5E598",
                   "Periphery" = "#3F85BB")

for (info_name in names(plot_info)) {
  info <- plot_info[[info_name]]
  seurat_obj <- info$object
  regions <- info$regions
  
  for (region in regions) {
    # Generate plot
    p <- SpatialDimPlot(
      seurat_obj,
      images = region,
      group.by = "WoundSite",
      cols = woundsite_pal
    ) +
      ggtitle(paste(info_name, region, "WoundSite")) &
      custom_theme +
      theme(
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)
      ) + NoLegend()
    
    # Save the plot
    ggsave_multidevice(
      filename_base = paste(info_name, region, "WoundSite", sep = "_"),
      plot = p,
      width = 6,
      height = 7
    )
  }
}

```

```{r}
p <- CustomSpatialHighlightPlot(
  seurat_obj = object_d7,
  dataset = "YWV03_D7PW",
  region = "ROI1",
  annotation_col = "first_type",
  cells_highlight = c("OC-like cells", "Tregs"),
  pt.size.highlight = 1,
  pt.size.unselected = 0.5,
  stroke.size = 0.05,
  pal = pal,
  custom_theme = custom_theme,
  scalebar_length_um = 100,
  scalebar_position = "bottomright"
) + NoLegend()

print(p)

```

We can also manually annotate wound center in loupe browser, and take that info to 
add to the respective datasets which need it
```{r Add wound center image to D4 and D7PW}
datasets <- c("YWV02_D30PW","YWV03_D7PW","YWV04_D4PW")
for (cond in datasets) {
  
  obj <- plot_info[[cond]]$object
  
  roi_file <- file.path(
    "~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Sequencing Repository/Data/Spatial Data/",
    cond, paste0(cond, "_Realignment/WoundSiteROI.csv")
  )

  # 1. Skip if ROI file doesn't exist
  if (!file.exists(roi_file)) {
    message("Skipping ", cond, " — no ROI file.")
    next
  }
  
  roi.coordinates <- read.csv(roi_file)
  
  # 2. Skip if no barcodes in file
  if (nrow(roi.coordinates) == 0) {
    message("Skipping ", cond, " — ROI file empty.")
    next
  }
  
  roi.assignments <- setNames(roi.coordinates$WoundSiteROI, roi.coordinates$Barcode)
  
  # 3. Add metadata only for barcodes that exist in obj
  roi.assignments <- roi.assignments[names(roi.assignments) %in% colnames(obj)]
  if (length(roi.assignments) == 0) {
    message("Skipping ", cond, " — no matching barcodes in object.")
    next
  }
  
  obj <- AddMetaData(obj, roi.assignments, col.name = "WoundSiteROI")
  
  # 4. Extract coordinates
  coords <- GetTissueCoordinates(obj)
  coords$WoundSiteROI <- obj$WoundSiteROI[rownames(coords)]
  
  # 5. Split by ROI — keep even if only 1 ROI present
  region_coords_list <- split(coords, coords$WoundSiteROI)
  
  # 6. Loop over ROIs and make convex hull overlays
  for (region_name in names(region_coords_list)) {
    region_coords <- region_coords_list[[region_name]]
    
    # Need ≥3 points for convex hull
    if (nrow(region_coords) < 3) {
      message("Skipping ROI ", region_name, " — not enough points.")
      next
    }
    
    hull_indices <- chull(region_coords$x, region_coords$y)
    hull_coords <- region_coords[hull_indices, c("x", "y")]
    hull_coords <- rbind(hull_coords, hull_coords[1, ]) # close polygon
    hull_coords$cell <- region_name
    
    print(hull_coords)  # sanity check
    
    region_segmentation <- CreateSegmentation(hull_coords)
    
    # Show overlay — this won't overwrite the object
    obj[[region_name]] <- Overlay(obj[["slice1"]], region_segmentation)
  }
  
  # Save modified object back into plot_info
  plot_info[[cond]]$object <- obj
}

SpatialDimPlot(object = plot_info$YWV04_D4PW$object, 
               group.by = "first_type", pt.size.factor = 3, images = "WoundCenter1") + NoLegend()

SpatialDimPlot(object = plot_info$YWV03_D7PW$object, 
               group.by = "first_type", pt.size.factor = 3, images = "WoundEdge1") + NoLegend()
```

With wound center added, we can save plots highlighting immune cells 
```{r Loop to save ROI plots for wounded timepoints}
setwd("~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 3 Spatial/Anatomical Sites")

# Define datasets and ROIs
datasets <- c("YWV02_D30PW", "YWV03_D7PW", "YWV04_D4PW")
rois     <- c("WoundCenter1", "WoundEdge1", "Periphery1")

# Collect all possible levels across all datasets
all_sites_global <- unique(unlist(
  lapply(datasets, function(cond) {
    unique(plot_info[[cond]]$object$AnatomicalSite)
  })
))

all_sites_global <- factor(all_sites_global, 
                           levels = sort(all_sites_global))

# Build a single consistent palette across datasets
site_pal_global <- setNames(
  hue_pal()(length(all_sites_global)),
  levels(all_sites_global)
)

# Loop through datasets and ROIs
for (cond in datasets) {
  
  obj <- plot_info[[cond]]$object
  
  # Subset palette to only sites present in this object
  site_pal <- site_pal_global[names(site_pal_global) %in% unique(obj$AnatomicalSite)]
  
  for (roi in rois) {
    # Skip if ROI overlay not present
    if (!(roi %in% names(obj))) {
      message("Skipping ", cond, " — no overlay for ", roi)
      next
    }
    
    # Make plot highlighting immune cells 
    # p <- CustomSpatialHighlightPlot(
    #   seurat_obj        = obj,
    #   region            = roi,
    #   annotation_col    = "first_type",
    #   cells_highlight   = immune_types,
    #   pt.size.highlight = 2,
    #   pt.size.unselected= 1,
    #   stroke.size       = 0.05,
    #   pal               = pal,
    #   custom_theme      = custom_theme
    # ) + NoLegend()
    # 
    # print(p)
    # 
    # # Save with dataset and ROI in filename
    # outname <- paste0(cond, "_", roi, "_ImmuneHighlight")
    # ggsave_multidevice(filename_base = outname, width = 7, height = 7)
    
    # Make plot showing anatomical sites
    p <- SpatialDimPlot(
      object = obj,
      images = roi,
      group.by = "AnatomicalSite", 
      pt.size.factor = 3.75,
      cols = site_pal) & 
      custom_theme +
      theme(panel.grid = element_blank()) 
    
    print(p) + NoLegend()
    
    # Save with dataset and ROI in filename
    outname <- paste0(cond, "_", roi, "_AnatomicalSites")
    ggsave_multidevice(filename_base = outname, width = 7, height = 7)
    
    p <- SpatialDimPlot(
      object = obj,
      images = "slice1",
      group.by = "AnatomicalSite", 
      pt.size.factor = 2.5,
      cols = site_pal) & 
      custom_theme +
      theme(panel.grid = element_blank()) 
    
    print(p)
    
    # Save with dataset and ROI in filename
    outname <- paste0(cond, "_", roi, "_AnatomicalSitesLegend")
    ggsave_multidevice(filename_base = outname, width = 7, height = 7)
  }
}

```
We can then take some of these highly expressed genes and plot feature plots
```{r Save feature plots}
# Save wound site associated features
feats <- c("Timp1","Srgn","Tnc","Spp1","Cfd","Car3","Ckm")

# Set default assay for all objects
plot_info <- lapply(plot_info, function(obj) {
  DefaultAssay(obj$object) <- "Spatial"
  return(obj)
})

# Loop through conditions, regions, and markers
for (cond in names(plot_info)) {
  obj <- plot_info[[cond]]$object
  regions <- plot_info[[cond]]$regions
  pt.size <- plot_info[[cond]]$pt.size.factor
  
  for (reg in regions) {
    for (marker in feats) {
      p <- SpatialFeaturePlot(
        object = obj,
        features = marker,
        images = reg,
        pt.size.factor = pt.size * 1.1
      ) +
        ggtitle(paste(cond, reg, marker, sep = " - ")) +
        scale_fill_gradient(low = "gray90", high = "red3") &
        custom_theme
      
      print(p)
      
      # Optionally save
      ggsave(paste0("WoundSite Marker FeatPlots/", cond, "_", reg, "_", marker,".png"), plot = p, width = 7, height = 7, dpi = 300)
    }
  }
}
```
We can also use the coexpression function to plot multiple genes at once
```{r Use coexpression featureplot to show wound site markers}
feats <- c("Timp1","Car3")
feat_colors <-  c("#7E1900", "#3F85BB")

p_uw <- CustomSpatialFeatureCoexpPlot(
  object = object_uw,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_uw

p_d4 <- CustomSpatialFeatureCoexpPlot(
  object = object_d4,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_d4

p_d7 <- CustomSpatialFeatureCoexpPlot(
  object = object_d7,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_d7

p_d30 <- CustomSpatialFeatureCoexpPlot(
  object = object_d30,
  image = "slice1",
  genes = feats,
  gene_colors = feat_colors,
  assay = "Spatial",
  pt.size.expr = 1,
  pt.size.zero = 0.5,
  bg.col       = "lightgrey"
)
p_d30

ggsave("YWV01_UW_WoundPeripheryFeatPlot.png", plot = p_uw, width = 7, height = 8)
ggsave("YWV02_D30PW_WoundPeripheryFeatPlot.png", plot = p_d30, width = 7, height = 8)
ggsave("YWV04_D4PW_WoundPeripheryFeatPlot.png", plot = p_d4, width = 7, height = 8)
ggsave("YWV03_D7PW_WoundPeripheryFeatPlot.png", plot = p_d7, width = 7, height = 8)
```
### Marker identification 
To characterize clusters which are in regions of interest, we will first find 
markers of each cluster vs the rest
```{r Find markers for BANKSY clusters}
# UW
uw_cluster_marks <- FindAllMarkers(object_uw, assay = "Spatial", only.pos = T,
                                   group.by = "banksy_cluster_0.5", verbose = F)
# D4PW
d4_cluster_marks <- FindAllMarkers(object_d4, assay = "Spatial", only.pos = T,
                                   group.by = "banksy_cluster_0.5", verbose = F)
# D7PW
d7_cluster_marks <- FindAllMarkers(object_d7, assay = "Spatial", only.pos = T,
                                   group.by = "banksy_cluster_0.5", verbose = F)
# D30PW
d30_cluster_marks <- FindAllMarkers(object_d30, assay = "Spatial",  only.pos = T,
                                   group.by = "banksy_cluster_0.5", verbose = F)
```
Examine BANKSY cluster markers in dotplot
```{r Plot DotPlots of BANKSY markers}
top_marks <- uw_cluster_marks %>%
  filter(avg_log2FC > 1 & p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice_head(n=5) %>%
  pull(gene)
  
DotPlot(object_uw, unique(top_marks), group.by = "banksy_cluster_0.5", cols = "RdYlBu",
        assay = "Spatial") +
  geom_point(
    aes(size = pct.exp),
    shape = 21,         # filled circle
    color = "black",    # black outline
    stroke = 0.5
  ) + 
  labs(y = "BANKSY Cluster")+
  guides(
    fill = guide_colorbar(title = "Average\nExpression"),
    size = guide_legend(title = "Percent\nExpressed")
  ) + 
 # theme_bw() +
  theme(
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1.1),
    axis.title.x = element_blank()) + custom_theme
```

```{r Loop to save BANKSY cluster marker dotplots}
# Save DE results as a list
de_results <- list("YWV01_UW" = uw_cluster_marks,
                   "YWV02_D30PW" = d30_cluster_marks,
                   "YWV03_D7PW" = d7_cluster_marks,
                   "YWV04_D4PW" = d4_cluster_marks)

# Loop over datasets to save dotplots
for(dataset in names(plot_info)){
  info <- plot_info[[dataset]]
  seurat_obj <- info$object
  cluster_marks <- de_results[[dataset]]
  
  top_marks <- cluster_marks %>%
  filter(avg_log2FC > 1 & p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice_head(n=5) %>%
  pull(gene)
  
  p <- DotPlot(seurat_obj, unique(top_marks), group.by = "banksy_cluster_0.5", 
             cols = "RdYlBu", assay = "Spatial") +
      geom_point(
        aes(size = pct.exp),
        shape = 21,
        color = "black",
        stroke = 0.5) +
        labs(y = "BANKSY Cluster") + 
    guides(
      fill = guide_colorbar(title = "Average\nExpression"),
      size = guide_legend(title = "Percent\nExpressed")
      ) +
      theme(
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1, vjust = 1.1),
        axis.title.x = element_blank()
      ) +
      custom_theme
  
  ggsave(paste0(dataset,"_BANKSYClusters_DotPlot.png"),plot = p,
         width = 15, height = 6)
}
```
## Find Wound-associated clusters
We are also interested in leveraging the spatial clusterings to identify wound-associated cellular niches.
To do so, we can quantify the representation of each banksy cluster across different
wound sites to determine which clusters are associated with various regions
```{r Determine BANKSY clusters found primarily in Wound site}
# Save function to determine proportion of cells found in wound site
get_wound_prop <- function(obj, cluster_col = "banksy_cluster_0.5", wound_col = "WoundSite") {
  tab <- table(obj@meta.data[[cluster_col]], obj@meta.data[[wound_col]])
  wound_levels <- colnames(tab)
  
  if ("Wound" %in% wound_levels) {
    counts_wound <- tab[, "Wound"]
  } else {
    counts_wound <- rep(0, nrow(tab))
    names(counts_wound) <- rownames(tab)
  }
  
  prop <- counts_wound / rowSums(tab)
  prop[is.na(prop)] <- 0
  data.frame(cluster = names(prop), prop_in_wound = prop)
}

results <- lapply(names(plot_info), function(name) {
  obj <- plot_info[[name]]$object
  df <- get_wound_prop(obj)
  df$sample <- name
  df
})

results_df <- do.call(rbind, results)

# Filter clusters with majority of spots in Wound
wound_clusters <- subset(results_df, prop_in_wound > 0.5)

# Split wound clusters by sample for downstream applications
wound_clusters <- split(wound_clusters$cluster, wound_clusters$sample)

# For D30PW, add clusters 0 and 1
wound_clusters$YWV02_D30PW <- c("0","1",wound_clusters$YWV02_D30PW)
```
We can also be less stringent in our criteria
```{r Determine BANKSY clusters with any number of cells in wound}
# Function to get clusters with ANY cells in "Wound"
get_clusters_with_wound <- function(obj, cluster_col = "banksy_cluster_0.5", wound_col = "WoundSite") {
  # Skip objects with no "Wound" at all
  if (!"Wound" %in% obj@meta.data[[wound_col]]) {
    return(NULL)
  }
  
  tab <- table(obj@meta.data[[cluster_col]], obj@meta.data[[wound_col]])
  
  # Get clusters where Wound count > 0
  counts_wound <- tab[, "Wound"]
  clusters_in_wound <- names(counts_wound[counts_wound > 0])
  
  data.frame(cluster = clusters_in_wound)
}

# Apply to all samples
results <- lapply(names(plot_info), function(name) {
  obj <- plot_info[[name]]$object
  clusters_df <- get_clusters_with_wound(obj)
  
  if (!is.null(clusters_df)) {
    clusters_df$sample <- name
    return(clusters_df)
  }
  
  NULL
})

# Remove NULLs and combine
results_df <- do.call(rbind, results)

# Split by sample
wound_clusters <- split(results_df$cluster, results_df$sample)
```
And we can find clusters which are found in wound and wound edge
```{r Find clusters primarily in Wound OR WoundEdge}
get_wound_and_edge_prop <- function(obj, cluster_col = "banksy_cluster_0.5", wound_col = "WoundSite",
                                   top_n = 3) {
  tab <- table(obj@meta.data[[cluster_col]], obj@meta.data[[wound_col]])
  wound_levels <- colnames(tab)
  clusters <- rownames(tab)

  # Safely get counts for Wound and WoundEdge (0 if missing)
  counts_wound <- if ("Wound" %in% wound_levels) tab[, "Wound"] else rep(0, nrow(tab))
  counts_edge  <- if ("WoundEdge" %in% wound_levels) tab[, "WoundEdge"] else rep(0, nrow(tab))
  names(counts_wound) <- clusters
  names(counts_edge) <- clusters

  # Calculate proportions per cluster
  prop_wound <- counts_wound / rowSums(tab)
  prop_edge <- counts_edge / rowSums(tab)
  prop_wound[is.na(prop_wound)] <- 0
  prop_edge[is.na(prop_edge)] <- 0

  # Identify top N clusters in Wound and WoundEdge by absolute counts
  top_wound_clusters <- names(sort(counts_wound, decreasing = TRUE))[1:top_n]
  top_edge_clusters <- names(sort(counts_edge, decreasing = TRUE))[1:top_n]

  # Logical vector: TRUE if cluster is in top N of both Wound and WoundEdge
  top_both <- clusters %in% intersect(top_wound_clusters, top_edge_clusters)

  data.frame(
    cluster = clusters,
    count_wound = counts_wound,
    prop_in_wound = prop_wound,
    count_edge = counts_edge,
    prop_in_edge = prop_edge,
    top_in_both = top_both
  )
}

# Usage example:
results <- lapply(names(plot_info), function(name) {
  obj <- plot_info[[name]]$object
  df <- get_wound_and_edge_prop(obj)
  df$sample <- name
  df
})

results_df <- do.call(rbind, results)

# Filter clusters with majority of spots in Wound
wound_edge_clusters <- subset(results_df, prop_in_wound > 0.5 | prop_in_edge > 0.5)

# Split wound clusters by sample for downstream applications
wound_edge_clusters <- split(wound_edge_clusters$cluster, wound_edge_clusters$sample)
```
We can examine where these clusters are in each object
```{r Save plots of wound site clusters}
# Loop to save plots of wound-site clusters for each object which has them
for(cond in names(wound_clusters)){
  # Isolate objects and clusters for each condition
  obj <- plot_info[[cond]]$object
  clusters <- wound_clusters[[cond]]
  regions <- plot_info[[cond]]$regions
  pt.size <- plot_info[[cond]]$pt.size.factor
  
  Idents(obj) <- "banksy_cluster_0.5"
  # Save color palette for each condition
  cluster_levels <- levels(obj)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), cluster_levels)
  cluster_colors <- cluster_colors[clusters]
  cluster_colors <- c(cluster_colors, Unselected = "grey90")
  
  cells <- CellsByIdentities(obj, idents = clusters)
  
   for (region in regions) {
      p <- SpatialDimPlot(obj,
                          images = region,
                          cells.highlight = cells, 
                          cols.highlight = cluster_colors) &
      custom_theme +
      theme(panel.grid = element_blank())
      
      ggsave(paste0(cond, "_WoundClusters_", region,".png"), width = 8, height = 7)
   }
}
```

We can then take the list of wound clusters and determine the cell types composing them
Quantify fine cell type representation within each specified cluster
```{r Quantify fine cell types in clusters of interest}
# Loop through each relevant condition
for (cond in names(wound_clusters)) {
  seurat_obj <- plot_info[[cond]]$object
  clusters_of_interest <- wound_clusters[[cond]]

  # Set identity class to clusters
  Idents(seurat_obj) <- "banksy_cluster_0.5"

  # Subset cells from the desired clusters
  selected_cells <- WhichCells(seurat_obj, idents = clusters_of_interest)

  # Get metadata and count first_type representation
  meta_sub <- seurat_obj@meta.data[selected_cells, ]

  summary_df <- meta_sub %>%
    dplyr::count(banksy_cluster_0.5, first_type) %>%
    group_by(banksy_cluster_0.5) %>%
    mutate(proportion = n / sum(n)) %>%
    ungroup() %>%
    filter(proportion >= 0.05)  # Filter out cell types < 5% in their cluster

p <-  ggplot(summary_df, aes(x = reorder(first_type, -proportion), y = proportion, fill = first_type)) +
    geom_bar(stat = "identity", color = "black") +
    facet_wrap(~banksy_cluster_0.5, scales = "free_y") +
    coord_flip() +
    scale_fill_manual(values = cell_colors) +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(
      y = "Proportion of Cluster"
    ) + theme(
      strip.text = element_text(size = 18), color = "black",
      axis.text.x = element_text(size = 12, color = "black"),
      axis.text.y = element_text(size = 12, color = "black"),
      axis.title.x = element_text(size = 16, color = "black"),
      axis.title.y = element_blank(),
      panel.border = element_rect(fill = NA, color = "black")
    ) + custom_theme

print(p)

  # Save plot
  ggsave(filename = paste0(cond,"ROICluster_CellTypeQuant.png"),
         plot = p, width = 16, height = 8)
}
```
We can also try to manipulate plotting features to make plots more consistent
```{r}
ncol_fixed <- 4  # Same as before
base_facet_height <- 2.5  # Height per facet row in inches (adjust as you like)

for (cond in names(wound_clusters)) {
  seurat_obj <- plot_info[[cond]]$object
  clusters_of_interest <- wound_clusters[[cond]]

  Idents(seurat_obj) <- "banksy_cluster_0.5"
  selected_cells <- WhichCells(seurat_obj, idents = clusters_of_interest)

  meta_sub <- seurat_obj@meta.data[selected_cells, ]

  summary_df <- meta_sub %>%
    dplyr::count(banksy_cluster_0.5, first_type) %>%
    group_by(banksy_cluster_0.5) %>%
    mutate(proportion = n / sum(n)) %>%
    ungroup() %>%
    filter(proportion >= 0.05, first_type != "Unknown")

  p <- ggplot(summary_df, aes(x = reorder(first_type, -proportion),
                              y = proportion,
                              fill = first_type)) +
    geom_bar(stat = "identity", color = "black") +
    facet_wrap(~banksy_cluster_0.5, scales = "free_y", ncol = ncol_fixed) +
    coord_flip() +
    scale_fill_manual(values = cell_colors) +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(y = "Proportion of Cluster") +
    theme(
      strip.text = element_text(size = 18, color = "black"),
      axis.text.x = element_text(size = 12, color = "black"),
      axis.text.y = element_text(size = 14, color = "black"),
      axis.title.x = element_text(size = 16, color = "black"),
      axis.title.y = element_blank(),
      panel.border = element_rect(fill = NA, color = "black")
    ) +
    custom_theme

  print(p)

  # Calculate how many rows of facets this plot will have
  n_clusters <- length(clusters_of_interest)
  n_rows <- ceiling(n_clusters / ncol_fixed)

  # Set height proportional to number of rows
  plot_height <- n_rows * base_facet_height

  ggsave(filename = paste0(cond, "ROICluster_CellTypeQuant.png"),
         plot = p,
         width = 16,  # your fixed width
         height = plot_height)
}

```
Next, find markers of these different populations and save plots
```{r Find markers for wound-specific clusters}
# Loop to find markers for clusters
for(cond in names(wound_clusters)){
  
  obj <- plot_info[[cond]]$object
  clusters <- wound_clusters[[cond]]
  
  Idents(obj) <- "banksy_cluster_0.5"
  # Save color palette for each condition
  cluster_levels <- levels(obj)
  cluster_colors <- setNames(viridis(length(cluster_levels), option = "H"), cluster_levels)
  cluster_colors <- cluster_colors[clusters]
  
  # Subset to only keep woundsite clusters
  obj <- subset(obj, idents = clusters)
  
  # Find markers for said clusters
  cluster_marks <- FindAllMarkers(obj, assay = "Spatial", 
                                  only.pos = T, 
                                  subset.ident = clusters,
                                  group.by = "banksy_cluster_0.5", 
                                  verbose = F)
  
  # Select top 5 genes per cluster
  top_marks <- cluster_marks %>%
  filter(avg_log2FC > 1 & p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice_head(n=10) %>%
  pull(gene)
  
  obj <- ScaleData(obj, assay = "Spatial", features = top_marks)
  
  p <- DoHeatmap(obj,
          features = top_marks,
          group.by = "banksy_cluster_0.5",
          group.colors = cluster_colors, 
          assay = "Spatial") +
    scale_fill_gradient2(
      low = "blue",
      mid = "white",
      high = "red",
      midpoint = 0
    ) + theme(
      axis.text.y = element_text(size = 10)
    ) & custom_theme

    ggsave(paste0(cond,"_WoundSiteClusters_Heatmap.png"),plot = p,
         width = 14, height = 8)
    
  # p <- DotPlot(obj, unique(top_marks), group.by = "banksy_cluster_0.5", 
  #            cols = "RdYlBu", assay = "Spatial") +
  #     geom_point(
  #       aes(size = pct.exp),
  #       shape = 21,
  #       color = "black",
  #       stroke = 0.5) +
  #       labs(y = "BANKSY Cluster") + 
  #   guides(
  #     fill = guide_colorbar(title = "Average\nExpression"),
  #     size = guide_legend(title = "Percent\nExpressed")
  #     ) +
  #     theme(
  #       axis.text.x = element_text(size = 8, angle = 45, hjust = 1, vjust = 1.1),
  #       axis.title.x = element_blank()
  #     ) +
  #     custom_theme
  
  # ggsave(paste0(cond,"_WoundSiteClusters_DotPlot.png"),plot = p,
  #        width = 15, height = 6)
  
}

```

With markers for each cluster identified, we can select only the relevant clusters
for region-specific analysis
```{r Select clusters in regions of interest and plot markers}
# Save cluster lists
d4_clusters <- c("1","2","5","6","8","12","13","14","15","20","21","22")

d7_clusters <- c("1","2","3","10","11","13","17","18","21")

d30_clusters <- c("1","2","6","11","16","17","18","19","22")

# D4PW
d4_cluster_marks1 <- FindAllMarkers(object_d4, assay = "Spatial", only.pos = T, subset.ident = d4_clusters,
                                   group.by = "alignment_clusters", verbose = F)

# Filter marker lists 
d4_cluster_marks_filtered <- d4_cluster_marks %>%
  filter(p_val_adj < 0.05 & cluster %in% d4_clusters) %>%
  group_by(cluster) %>%
  slice_head(n = 5)
  
# d4_cluster_marks_filtered <- d4_cluster_marks %>%
#   filter(p_val_adj < 0.05, cluster %in% d4_clusters) %>%
#   group_by(gene) %>%
#   filter(n() == 1) %>%
#   group_by(cluster) %>%
#   slice_head(n = 5) %>%
#   ungroup()

# Plot marker genes on DotPlots
Idents(object_d4) <- "alignment_clusters"
DotPlot(object_d4, unique(d4_cluster_marks_filtered$gene), idents = d4_clusters) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
We can also subset the objects and find markers across each of the wound-associated clusters
```{r Subset D4 and find ROI markers}
Idents(object_d4) <- "alignment_clusters"
object_d4_subset <- subset(object_d4, idents = d4_clusters)

d4_cluster_marks <- FindAllMarkers(object_d4_subset, assay = "Spatial", only.pos = T,
                                   group.by = "alignment_clusters", verbose = F)

d4_cluster_marks_filtered <- d4_cluster_marks %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice_head(n = 5)

DotPlot(object_d4, unique(d4_cluster_marks_filtered$gene), idents = d4_clusters,
        cols = "RdYlBu") +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.25) +
  guides(color = guide_colorbar(title = "Average\nExpression"),
         size=guide_legend(override.aes=list(shape=21, colour="black", fill="white"))) +
  theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1, face = "italic"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 18),
        axis.title.y = element_blank()) + custom_theme
ggsave_multidevice(filename_base = "YWV04_D4PW_ROI_MarkerDotPlot", width = 12, height = 6)
```
```{r Subset D7 and find ROI markers}
Idents(object_d7) <- "alignment_clusters"
object_d7_subset <- subset(object_d7, idents = d7_clusters)

d7_cluster_marks <- FindAllMarkers(object_d7_subset, assay = "Spatial", only.pos = T,
                                   group.by = "alignment_clusters", verbose = F)

d7_cluster_marks_filtered <- d7_cluster_marks %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice_head(n = 5)

DotPlot(object_d7, unique(d7_cluster_marks_filtered$gene), idents = d7_clusters,
        cols = "RdYlBu", assay = "Spatial") +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.25) +
  guides(color = guide_colorbar(title = "Average\nExpression"),
         size=guide_legend(override.aes=list(shape=21, colour="black", fill="white"))) +
  theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1, face = "italic"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 18),
        axis.title.y = element_blank()) + custom_theme
ggsave_multidevice(filename_base = "YWV03_D7PW_ROI_MarkerDotPlot", width = 12, height = 6)
```
```{r Subset D30 and find ROI markers}
Idents(object_d30) <- "alignment_clusters"
object_d30_subset <- subset(object_d30, idents = d30_clusters)

d30_cluster_marks <- FindAllMarkers(object_d30_subset, assay = "Spatial", only.pos = T,
                                   group.by = "alignment_clusters", verbose = F)

d30_cluster_marks_filtered <- d30_cluster_marks %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice_head(n = 5)

DotPlot(object_d30, unique(d30_cluster_marks_filtered$gene), idents = d30_clusters,
        cols = "RdYlBu", assay = "Spatial") +
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.25) +
  guides(color = guide_colorbar(title = "Average\nExpression"),
         size=guide_legend(override.aes=list(shape=21, colour="black", fill="white"))) +
  theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1, face = "italic"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 18),
        axis.title.y = element_blank()) + custom_theme
ggsave_multidevice(filename_base = "YWV02_D30PW_ROI_MarkerDotPlot", width = 12, height = 6)
```

### Analysis of peripheral cells
We are also interested in quantifying changes in the representation of immune cells
in the wound periphery
```{r Quantify changes in cells in peripheral tissue}

```


### OC-like analysis
Load in the spatial quantification data calculated elsewhere
```{r Loop to plot histograms of genes of interest}
library(Seurat)
library(tidyverse)
library(ggbeeswarm)
library(scCustomize)
library(readxl)

# -----------------------------
# Setup
# -----------------------------
target_population <- "OC-like cells"
genes_of_interest <- c("Tnfsf11", "Tnfrsf11a", "Tnfrsf11b", "Mmp3")
genes_of_interest <- c("Spp1","Ccl3","Ccl8","Ccl9")
gene_colors <- setNames(c("#D53E4F","#3288BD","#66C2A5","#FDAE61"),genes_of_interest) 

gene_colors <- setNames(c("#D53E4F","#FDAE61","#66C2A5","#3288BD"),genes_of_interest) 
# Full projection files (all cells)
full_files <- c(
  "YWV01_UW_MetaclusterDistributionFromWoundSite.csv",
  "YWV04_D4PW_MetaclusterDistributionFromWoundSite.csv",
  "YWV03_D7PW_MetaclusterDistributionFromWoundSite.csv",
  "YWV02_D30PW_MetaclusterDistributionFromWoundSite.csv"
)

# Immune-only files with barcodes
immune_files <- c(
  "YWV01_UW_DistributionFromWoundSite_ImmuneOnly.xlsx",
  "YWV04_D4PW_DistributionFromWoundSite_ImmuneOnly.xlsx",
  "YWV03_D7PW_DistributionFromWoundSite_ImmuneOnly.xlsx",
  "YWV02_D30PW_DistributionFromWoundSite_ImmuneOnly.xlsx"
)

# Seurat objects
objects <- list(
  "YWV01_UW" = object_uw,
  "YWV04_D4PW" = object_d4,
  "YWV03_D7PW" = object_d7,
  "YWV02_D30PW" = object_d30
)

base_path <- "~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 3 Spatial/SpaceFold"

df_list <- list()

# Define a palette for target population
pal <- c("OC-like cells" = "#4D8DC3")

# --------- Loop through datasets--------
for (i in seq_along(full_files)) {
  
  full_file <- full_files[i]
  immune_file <- immune_files[i]
  
  # Extract sample name as e.g., "YWV01_UW"
  sample_name <- paste(strsplit(full_file, "_")[[1]][1:2], collapse = "_")
  
  message("Processing ", sample_name)
  
  # --- Read full projection (all cells) ---
  df_full <- read.csv(file.path(base_path, full_file), row.names = 1)
  
  if (!"barcode" %in% colnames(df_full)) df_full$barcode <- rownames(df_full)
  
  # --- Read immune-only file with barcodes ---
  df_immune <- read_xlsx(file.path(base_path, immune_file))
  
  # --- Filter for target population ---
  df_target <- df_immune %>%
    filter(celltype == target_population) %>%
    dplyr::select(barcode)
  
  # --- Pull gene expression directly from Seurat ---
  seurat_obj <- objects[[sample_name]]
  DefaultAssay(seurat_obj) <- "Spatial"
  genes_present <- intersect(genes_of_interest, rownames(seurat_obj))
  
  expr_mat <- as.data.frame(
    t(GetAssayData(seurat_obj, assay = "Spatial", slot = "data")[genes_present, ])
  )
  expr_mat$barcode <- rownames(expr_mat)
  
  # --- Merge expression with full projection ---
  df_full <- left_join(df_full, expr_mat, by = "barcode")
  
  # --- Timepoint label ---
  tp <- case_when(
    grepl("UW", sample_name) ~ "UW",
    grepl("D4PW", sample_name) ~ "D4PW",
    grepl("D7PW", sample_name) ~ "D7PW",
    grepl("D30PW", sample_name) ~ "D30PW",
    TRUE ~ "Other"
  )
  df_full$timepoint <- tp
  
  # --- Plot 1: target population distribution ---
  # Filter immune-only data for the target population
  df_plot1 <- df_immune %>% filter(celltype == target_population)
  # --- Bin OC-like cell counts along the tissue axis ---
  df_plot1_binned <- df_plot1 %>%
    mutate(proj_bin = cut(projection, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)) %>%
    group_by(proj_bin, celltype) %>%
    summarise(
      cell_count = n(),
      proj_mid = mean(projection, na.rm = TRUE),
      .groups = "drop"
    )

  # Plot localization as a histogram
  p0 <- ggplot(df_plot1_binned, aes(x = proj_mid, y = cell_count, fill = celltype)) +
  geom_col(position = "dodge", color = "black", alpha = 0.9) +
  scale_fill_manual(values = pal) +
  labs(
    x = "Position along tissue axis",
    y = "Cell Count",
    title = paste0(tp, " — ", target_population, " distribution")
  ) +
  coord_cartesian(ylim = c(0, 80)) +
  theme_ggprism_mod(base_size = 14) +
  theme(legend.position = "none") +
  custom_theme
  
  # Plot localization as a beeswarm plot
  p1 <- ggplot(df_plot1, aes(x = projection, y = celltype, fill = celltype)) +
    geom_quasirandom(
      shape = 21,
      size = 1.8,
      stroke = 0.25,
      color = "black",
      alpha = 0.9
    ) +
    scale_fill_manual(values = pal) +
    labs(x = "Position along tissue axis",
         title = paste0(tp, " — ", target_population, " distribution")) +
    theme_ggprism_mod(base_size = 14) +
    theme(legend.position = "none",
          axis.title.y = element_blank()) + 
    custom_theme
  
  # --- Plot 2: gene expression heatmap ---
  df_long <- df_full %>%
    pivot_longer(cols = all_of(genes_present),
                 names_to = "gene", values_to = "expression") %>%
    mutate(proj_bin = cut(projection, breaks = seq(0,1,by=0.05), include.lowest=TRUE)) %>%
    group_by(gene, proj_bin) %>%
    summarise(mean_expr = mean(expression, na.rm=TRUE),
              proj_mid = mean(projection, na.rm=TRUE),
              .groups="drop")
  
  p2 <- ggplot(df_long, aes(x = proj_mid, y = gene, fill = mean_expr)) +
    geom_tile() +
    scale_fill_viridis_c(option = "plasma", direction = -1) +
    labs(x = "Position along tissue axis",
         y = NULL,
         fill = "Mean expression",
         title = paste0(tp, " — Expression of genes of interest")) +
    theme_ggprism_mod(base_size = 14) +
    theme(axis.text.y = element_text(face="italic"))
  
  # --- Plot 3: histogram of positive expression ---
  df_hist <- df_full %>%
    pivot_longer(cols = all_of(genes_present),
                 names_to = "gene", values_to = "expression") %>%
    mutate(expr_positive = expression > 0,
           proj_bin = cut(projection, breaks=seq(0,1,by=0.05), include.lowest=TRUE)) %>%
    group_by(gene, proj_bin) %>%
    summarise(positive_count = sum(expr_positive, na.rm=TRUE),
              proj_mid = mean(projection, na.rm=TRUE),
              .groups="drop")
  
  p3 <- ggplot(df_hist, aes(x = proj_mid, y = positive_count, fill = gene)) +
    geom_col(position = "dodge", color="black", alpha=0.9) +
    scale_fill_manual(values = gene_colors) +
   # scale_fill_viridis_d(option="plasma") +
    labs(x = "Position along tissue axis",
         y = "Number of cells (expression > 0)",
         fill = "Gene",
         title = paste0(tp, " — Localization of gene expression")) +
    theme_ggprism_mod(base_size = 14)
  
  # Make histogram faceted by each gene
  p4 <- p3 +
    facet_grid(rows = vars(gene), scales = "free_y") + 
    NoLegend()
  
  # --- Plot 5: mean expression per bin (faceted by gene) ---
  df_mean <- df_full %>%
    pivot_longer(
      cols = all_of(genes_present),
      names_to = "gene",
      values_to = "expression"
    ) %>%
    mutate(proj_bin = cut(projection, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)) %>%
    group_by(gene, proj_bin) %>%
    summarise(
      mean_expr = mean(expression, na.rm = TRUE),
      proj_mid = mean(projection, na.rm = TRUE),
      .groups = "drop"
    )
  
  p5 <- ggplot(df_mean, aes(x = proj_mid, y = mean_expr, fill = gene)) +
    geom_col(position = "dodge", color = "black", alpha = 0.9) +
    scale_fill_manual(values = gene_colors) +
   # scale_fill_viridis_d(option = "plasma") +
    facet_wrap(~gene, ncol = 1, scales = "free_y") +
    labs(
      x = "Position along tissue axis",
      y = "Mean expression",
      fill = "Gene",
      title = paste0(tp, " — Mean expression of genes across tissue axis")
    ) +
    theme_ggprism_mod(base_size = 14) + 
    NoLegend()

  # --- Save plots ---
  base_out <- tools::file_path_sans_ext(full_file)
  
  ggsave_multidevice(
    filename_base = paste0(base_out, "_", gsub(" ", "_", target_population), "_CountHistogram"),
    plot = p0, width = 5, height = 3, dpi = 300
  )
  
  ggsave_multidevice(
    filename_base = paste0(base_out, "_", gsub(" ", "_", target_population), "_Distribution"),
    plot = p1, width = 5, height = 3, dpi = 300
  )
  
  ggsave_multidevice(
    filename_base = paste0(base_out, "_GenesOfInterest_Heatmap"),
    plot = p2, width = 6, height = 5, dpi = 300
  )
  
  ggsave_multidevice(
    filename_base = paste0(base_out, "_GenesOfInterest_Histogram"),
    plot = p3, width = 6, height = 5, dpi = 300
  )
  
  ggsave_multidevice(
    filename_base = paste0(base_out, "_GenesOfInterest_FacetedHistogram"),
    plot = p4, width = 5, height = 7, dpi = 300
  )
  
  ggsave_multidevice(
    filename_base = paste0(base_out, "_GenesOfInterest_MeanExpressionHist"),
    plot = p5, width = 5, height = 7, dpi = 300
  )
  
  # --- Store for combined plotting if needed ---
  df_list[[sample_name]] <- df_full
}

# Optional: combine all timepoints
combined_df <- bind_rows(df_list)

```
We can do the same, this time asking is the expression specific to an immune mode
```{r}
library(Seurat)
library(tidyverse)
library(ggbeeswarm)
library(scCustomize)
library(readxl)

# -----------------------------
# Setup
# -----------------------------
genes_of_interest <- c("Spp1","Ccl8")
gene_colors <- setNames(c("#D53E4F","#66C2A5"), genes_of_interest)

# Full projection files (all cells)
full_files <- c(
  "YWV01_UW_MetaclusterDistributionFromWoundSite.csv",
  "YWV04_D4PW_MetaclusterDistributionFromWoundSite.csv",
  "YWV03_D7PW_MetaclusterDistributionFromWoundSite.csv",
  "YWV02_D30PW_MetaclusterDistributionFromWoundSite.csv"
)

# Immune-only files with barcodes
immune_files <- c(
  "YWV01_UW_DistributionFromWoundSite_ImmuneOnly.xlsx",
  "YWV04_D4PW_DistributionFromWoundSite_ImmuneOnly.xlsx",
  "YWV03_D7PW_DistributionFromWoundSite_ImmuneOnly.xlsx",
  "YWV02_D30PW_DistributionFromWoundSite_ImmuneOnly.xlsx"
)

# Seurat objects
objects <- list(
  "YWV01_UW" = object_uw,
  "YWV04_D4PW" = object_d4,
  "YWV03_D7PW" = object_d7,
  "YWV02_D30PW" = object_d30
)

base_path <- "~/SRSP Laboratory Dropbox/SRSP Lab/Resources/Town Square - Data Share/Manuscripts/Wound Healing Atlas Paper/Immunome Paper/Figure 3 Spatial/SpaceFold"

# Define immune mode groups
recruitment <- c("Monocytes I","Monocytes II","Monocytes III","Macrophages I","Macrophages III",
                 "OC-like cells","moDCs","Neutrophils I","Neutrophils II","Basophils")
restoration <-c("cDC1","cDC2","pDCs","Neutrophils III","Mast cells","NK cells","T cells","B cells")
neostasis <- c("Macrophages II","Macrophages IV","Cycling Macrophages",
               "Migratory DCs","Langerhans cells","γδT cells","Tregs","ILCs")

recruitment_modes <- list(
  "Recruitment" = recruitment,
  "Restoration" = restoration,
  "Neostasis" = neostasis
)

df_list <- list() # store combined data

# -----------------------------
# Loop through datasets
# -----------------------------
for (i in seq_along(full_files)) {

  full_file <- full_files[i]
  immune_file <- immune_files[i]

  # Extract sample name
  sample_name <- paste(strsplit(full_file, "_")[[1]][1:2], collapse = "_")
  message("Processing ", sample_name)

  # --- Read full projection (all cells) ---
  df_full <- read.csv(file.path(base_path, full_file), row.names = 1)
  if (!"barcode" %in% colnames(df_full)) df_full$barcode <- rownames(df_full)

  # --- Read immune-only file with barcodes ---
  df_immune <- read_xlsx(file.path(base_path, immune_file))

  # --- Pull gene expression directly from Seurat ---
  seurat_obj <- objects[[sample_name]]
  DefaultAssay(seurat_obj) <- "Spatial"
  genes_present <- intersect(genes_of_interest, rownames(seurat_obj))

  expr_mat <- as.data.frame(
    t(GetAssayData(seurat_obj, assay = "Spatial", slot = "data")[genes_present, ])
  )
  expr_mat$barcode <- rownames(expr_mat)

  # --- Merge expression with full projection ---
  df_full <- left_join(df_full, expr_mat, by = "barcode")

  # --- Timepoint label ---
  tp <- case_when(
    grepl("UW", sample_name) ~ "UW",
    grepl("D4PW", sample_name) ~ "D4PW",
    grepl("D7PW", sample_name) ~ "D7PW",
    grepl("D30PW", sample_name) ~ "D30PW",
    TRUE ~ "Other"
  )
  df_full$timepoint <- tp

  # --- Loop over immune mode groups ---
  for (mode_name in names(recruitment_modes)) {

    group_cells <- recruitment_modes[[mode_name]]

    # Filter immune data for this group
    df_group <- df_immune %>% filter(celltype %in% group_cells)
    df_group_expr <- df_full %>% filter(barcode %in% df_group$barcode)

    # Define color palette for this group (optional: one color per cell type)
    pal <- setNames(
      RColorBrewer::brewer.pal(n = length(group_cells), name = "Set2")[seq_along(group_cells)],
      group_cells
    )

    # --- Plot 4: histogram of positive expression ---
    df_hist <- df_group_expr %>%
      pivot_longer(cols = all_of(genes_present),
                   names_to = "gene", values_to = "expression") %>%
      mutate(expr_positive = expression > 0,
             proj_bin = cut(projection, breaks=seq(0,1,by=0.05), include.lowest=TRUE)) %>%
      group_by(gene, proj_bin) %>%
      summarise(positive_count = sum(expr_positive, na.rm=TRUE),
                proj_mid = mean(projection, na.rm=TRUE),
                .groups="drop")
      
    p3 <- ggplot(df_hist, aes(x = proj_mid, y = positive_count, fill = gene)) +
      geom_col(position = "dodge", color="black", alpha=0.9) +
      scale_fill_manual(values = gene_colors) +
      labs(x = "Position along tissue axis",
           y = "Number of cells (expression > 0)",
           fill = "Gene",
           title = paste0(tp, " — ", mode_name, " positive gene expression")) +
      theme_classic() +
      theme(text = element_text(size = 14, family = "Arial"))

    p4 <- p3 +
      facet_wrap(~gene, ncol = 1, scales = "free_y") + 
      labs(title = paste0(tp, " — ", mode_name, " gene expression")) +
      theme(strip.text = element_text(face = "italic", size = 16),
       strip.background = element_blank() ) +
      NoLegend()

    # --- Plot 5: mean expression per bin (faceted) ---
    df_mean <- df_group_expr %>%
      pivot_longer(cols = all_of(genes_present),
                   names_to = "gene", values_to = "expression") %>%
      mutate(proj_bin = cut(projection, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)) %>%
      group_by(gene, proj_bin) %>%
      summarise(mean_expr = mean(expression, na.rm = TRUE),
                proj_mid = mean(projection, na.rm = TRUE),
                .groups = "drop")

    p5 <- ggplot(df_mean, aes(x = proj_mid, y = mean_expr, fill = gene)) +
      geom_col(position = "dodge", color = "black", alpha = 0.9) +
      scale_fill_manual(values = gene_colors) +
      facet_wrap(~gene, ncol = 1, scales = "free_y") +
      labs(x = "Position along tissue axis",
           y = "Mean expression",
           fill = "Gene",
           title = paste0(tp, " — ", mode_name, " mean gene expression")) +
      theme(text = element_text(size = 14, family = "Arial"),
            strip.text = element_text(face = "italic", size = 16),
       strip.background = element_blank() ) +
      NoLegend()

    # --- Save plots ---
    base_out <- tools::file_path_sans_ext(full_file)

    ggsave_multidevice(paste0(base_out, "_", mode_name, "_Histogram"), plot = p3, width = 6, height = 5, dpi = 300)
    ggsave_multidevice(paste0(base_out, "_", mode_name, "_FacetedHistogram"), plot = p4, width = 5, height = 4, dpi = 300)
    ggsave_multidevice(paste0(base_out, "_", mode_name, "_MeanExpressionHist"), plot = p5, width = 5, height = 4, dpi = 300)

  } # end group loop

  # Store combined expression for optional downstream analyses
  df_list[[sample_name]] <- df_full

} # end sample loop

# Combine all samples if needed
combined_df <- bind_rows(df_list)

```

```{r}

```

